### 2012-10-03 latex/paragraphs.tex
10/3 weekly report done

\documentclass{article}

\begin{document}
\title{Weekly Paragraphs}
\author{Kui Tang}
\date{\today}
\maketitle
\section*{October 3, 2012}
To appear in: Related Work

All known connectomes reply on manual labels. White et al. published the first connectome of \emph{C. elegans} in 1986, marking axons, dendrites, and cell bodies in each frame and tracing the labels through frames \cite{White12111986}. More recent work uses computer assistance but fundamentally rely on a skilled human's identification of neural bodies \cite{Jarrell2012,Bock2011Roberts2011}.

Koshevoy et al. \cite{Koshevoy2006} use a variant of SIFT \cite{Lowe2004} to register a stack of images into a 3D volume. According to Koshevoy, images between images exhibit warping due to the physical slicing, physical structural changes, and a distinct rotation and displacement on each frame. From the displacements of individual keypoints, a global transformation vector is calculated, allowing slices to be properly aligned. However, the SIFT features do not capture neural processes well, and thus is unsuited for identifying neurons across images.

\bibliographystyle{plain}
\bibliography{refs.bib}
\end{document}

### 2012-10-03 latex/slides.tex
10/3 weekly report done

\documentclass{beamer}
\usetheme{Berkeley}
\usecolortheme{dolphin}
\title{Weekly Slides}
\author{K.~Tang\inst{1}}
\institute{
  \inst{1}
  Columbia University
}
\date{2012}

\begin{document}
\frame{\titlepage}

\section{October 3, 2012}
\begin{frame}
  \frametitle{SIFT fails to identify neural processes}
  \includegraphics[width=\textwidth]{fig/sift_bock11}\\
  Two adjacent-depth images from the bock11 dataset \cite{Bock2011}. While large SIFT features are consistently transformed, small features move wildly. Thus SIFT is not appropriate for identifying neural processes.
\end{frame}

\section{References}
  \begin{frame}[allowframebreaks]
  \frametitle{References}
  \bibliographystyle{plain}
  \bibliography{refs.bib}
\end{frame}

\end{document}

### 2012-10-15 latex/paragraphs.tex
paragraphs for this week

@@
 
+\section*{October 15, 2012}
+Introduction:
+
+Related Work:
+
+EM images present several challenges for classical segmentation algorithm.
+First, structures or interest are small and densely packed \cite{Jain2007}.
+Second, due to the physical cutting process, images may be warped from slice to slice \cite{Koshevoy2006}.
+
+Semi-automated methods take small amounts of user input and then use the morphology of the image segment entire neural processes.
+Marker-Controlled watersheds \cite{Gonzales2008} is a classic technique, requiring the user to mark background and foreground.
+Roberts et al.
+\cite{Roberts2011} require a "scribble" inside a neuron on the first and last slices.
+At each slice, their algorithm segments the neuron by minimizing the anisotropic weighted total variation \cite{Unger2009} and propagating the inferred boundary as soft constraints for the next layer.
+
+Unfortunately, current baseline speeds of computer-assisted manual tracing would take 10,000 person-years to construct a mouse cortical column, suggesting that fully automatic methods are necessary \cite{Briggman2006}.
+
+Fully-automated methods use manual labels as input for supervised learning.
+Segmentation can be done in voxel space, after which the learned segmentation can be directly used as a 3D reconstruction.
+Because cells are usually disconnected, segmentation can be reduced to binary classification of whether voxels are inside or outside a cell.
+Jain et al.
+\cite{Jain2007} achieved an 8.0\% test error with a 34,000 convolutional neural network, a statistically significant improvement other the 18.95\% error of thresholding, while MRFs and CRFs do no better than thresholding.
+Andres et al.
+obtain similar performance, at 8\% misclassification, using the watershed algorithm over a probability map of a random forest with hand-engineered features \cite{Andres2008, Gonzales2008}.
+However, the dataset used by both used a special intracellular stain, making the segmentation job easier.
+
+Alternatively, one can segment in pixel space and track the identity of segmented bodies in the vertical dimension.
+This is motivated by the poor vertical resolution of EM images: the thinnest slices are 40nm think \cite{Kaynig2010b, Briggman2006}.
+Jurrus et al.
+\cite{Jurrus2008} first obtain 2D segmentations with watershed after preprocessing linear diffusion filtering (NB: learn what this is!), and construct a weighted graph of these segments, with edges between adjacent sections with edge costs favoring high cross-correlation of segments low lateral displacement.
+Each neural process is thus a shortest path through this graph, computed by Dijkstra's algorithm.
+This cost function is robust to merging and loss of sections, but because Dijkstra's algorithm finds a single path, it cannot model branching.
+Kaynig et al.
+\cite{Kaynig2010a} presuppose segmentation and represent the image as a graph of regions, like Jurrus.
+Pairwise features are represented as edge weights, and a random forest classifier is trained to output a scalar similarity score for each pair.
+Then a probabilistic model encoding geometric constraints with cross-section correspondences as hidden variables is trained with EM.
+Finally, the maximum-likelihood result is the 3D reconstruction.
+These pairwise correspondences are fed to an agglomerative clustering algorithm, constructing the neural processes from ground-up in the same manner as a human.
+Unlike in Jurrus, processes may merge.
+However, each region is explicitly limited to two correspondences: one above and below, so processes still cannot branch.
+
+Feature engineering has improved 2D segmentation.
+Venkataraju et al.
+\cite{Venkataraju2009} generated 100 Hessian neighborhood features at each pixel and classified Jurrus's dataset with AdaBoost.
+Lucchi et al.
+\cite{Lucchi2010} develop ray descriptors, which capture shape features without a predefined shape model.
+These features are rotation, scale, and translation invariant, but strongly discriminate the types of objects observed in EM images.
+There features were used in SVMs to classify interior and boundary points of mitochondria, which are considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy.
+Kaynig et al.
+encode perceptual grouping constraints---rules of thumb used by human experts to group neuron structures---as pairwise interaction terms in an energy function solved by graph cut.
+Minimizing the energy function gives a segmentation, and identifying maximum overlapping regions between slices is sufficient to obtain a 3D reconstruction
+
 \bibliographystyle{plain}

### 2012-11-01 src/beta_intensity_like.m
beta likelihood sampler works; experiment doesn't

@@ -12,6 +12,7 @@ function [ like ] = beta_intensity_like( mm, k, x )
     % Parameters
     Nsamps = 50;
     burn_in = 100;
+    MAX_LIKE = 1e10;
     
     % Update step: compute the posterior     
     like = zeros(Nk, 1);
@@ -25,10 +26,11 @@ function [ like ] = beta_intensity_like( mm, k, x )
     %
     % Actually, make this a TODO. Just evaluate each sample. We should get
     % the probabilistic behavior by drawing many, many distribution, and
-    % the general smoothness of the posterior.       
-    [a b] = beta_sm_to_ab(mm.beta_posterior_samps{k}(:,1), ...
-                          mm.beta_posterior_samps{k}(:,2));
-
+    % the general smoothness of the posterior.
+    s = mm.beta_posterior_samps{k}(:,1);                          
+    m = mm.beta_posterior_samps{k}(:,2);
+    [a b] = beta_sm_to_ab(s, m);        
+    
     for ns = 1:Nsamps
         like = like + betapdf(x, a(ns), b(ns));
     end
@@ -36,7 +38,19 @@ function [ like ] = beta_intensity_like( mm, k, x )
     like = like / Nsamps;
     
     disp(['Accept rate: ', num2str(accept_rate)]);
+    figure(10);
+    subplot(2,1,1);    
+    hist(s);
+    title('s samples');
     
-    assert(all(like >= 0));    
+    subplot(2,1,2);
+    hist(m);
+    title('m samples');
+    drawnow;
+    
+    % Clamp the top
+    like(like == Inf) = MAX_LIKE;
+    
+    assert(all(like >= 0 & like ~= Inf));    
 end
 

### 2012-11-01 src/experiment/test_beta_mixture.m

@@ -4,7 +4,7 @@ ss_sigmoid = @(gmm_, X_) (1 / gmm_.N) * (1 / 0.21265926) * (1 - (1 ./ (1 + exp( 
 
 %% Prior
 mm = make_spatial_mm_prior(PHI1, 6);
-mm.prior_beta_params = prior_params = [ 1 0.1 0.0005 ];
+mm.prior_beta_params = [ 0.1 0.1 0.0005 ];
 mm.background_like = ss_sigmoid;
 mm.beta_posterior_samps = cell(6, 1);
 mm.feature_likes = { @beta_intensity_like };
@@ -13,4 +13,4 @@ mm.feature_likes = { @beta_intensity_like };
 %% Samples!
 figure;
 beta_samps = run_mm(@imm_iter, PHI1, f1, mm, 400, 100, 10);
-title('GMM triangular background, k = 6');
+title('IMM beta [0.1 0.1 0.005] 1-avoiding, k =6 ');

### 2012-11-01 src/mcmc/beta_posterior_mh.m

@@ -27,8 +27,21 @@ function [ s, m_star, accept ] = beta_posterior_mh( prior_params, s_old, m_star_
               lognlike([log(m_star_old) stdev], m_star) - ...
               lognlike([log(s_old)      stdev], s);                  
     
-    r = exp(log_top - log_bot);
-    assert(~isnan(r));    
+    % Special cases
+    if log_top == -Inf && log_bot == -Inf
+        % This point has zero probability.
+        r = 0;
+    elseif log_top == Inf && log_bot == Inf
+        % This point has huge probability, so it is certainly > 1
+        r = 2;
+    elseif isinf(log_top) || isinf(log_bot)
+        % Infinities point in opposite directions. Oh no!
+        error('rOppInf', 'opposite infinities in acceptance! check your math.');
+    elseif isnan(log_top) || isnan(log_bot)
+        error('rNan', 'nan in acceptance! check your math.');
+    else                    
+        r = exp(log_top - log_bot);
+    end    
     
     if rand > r % M-H reject        
         s = s_old;

### 2012-11-01 src/mcmc/beta_posterior_samples.m

@@ -12,8 +12,17 @@ function [ samps, accept_rate ] = beta_posterior_samples( prior_params, x, Nsamp
     % Sample!
     samps = zeros(Nsamps, 2);
     
-    mle_params = betafit(x);
-    [s m_star] = beta_ab_to_sm(mle_params(1), mle_params(2));
+    % MLE init is stupid and underestimates mean. Do method of moments
+    % instead.
+    m = mean(x);
+    m_star = interval_to_positive(m);
+    % Notebook pp 29
+    s = var(x) / (m * (1 - m));
+            
+    % Our prior is designed for s > 2.
+    % Push away for a low-likelihood region
+    s = max(4, s);
+        
     for n = 1:Nburn_in
         [s m_star] = beta_posterior_mh(prior_params, s, m_star, x, stdev);
     end

### 2012-11-01 src/mcmc/beta_posterior_updf.m

@@ -1,6 +1,6 @@
 function [ up ] = beta_posterior_updf( prior_params, s, m, x, logp )
 % Posterior of the beta_prior_updf distribution.
-% Bouguila (19) and (20)
+% Bouguila (19) and (20). Modified to always return s > 1.
 %
 % Used as a subroutine in Beta mixture models. Class labels don't exist
 % here: handle them (through indexing the big X) before calling this.
@@ -23,15 +23,22 @@ function [ up ] = beta_posterior_updf( prior_params, s, m, x, logp )
 
     log_prior = beta_prior_updf(prior_params, s, m, true);
     
+    % It's an indicator for (1, Inf)
+%     if s < 2
+%         
+%         up = -Inf;
+%         
+%     else    
     sm  = s * m;
     s1m = s * (1 - m);
     log_Z     = N * (gammaln(s) - gammaln(sm) - gammaln(s1m));        
     log_like  = sm * sum(log(x)) + s1m * sum(log1p(-x));
     up = log_prior + log_Z + log_like;
-    
+
     if nargin > 4 && logp
         return
     end
+%    end
     
     up = exp(up);
     assert(all(~isinf(up) & ~isnan(up) & up >= 0));

### 2012-11-01 src/mcmc/test_beta_posterior_rnd.m

@@ -1,31 +1,17 @@
-%% Look at the data
+%% Plot the prior
 %
 % Want: s > 1 (unimodal)
 prior_params = [ 0.1 0.1 0.0005 ];
-figure(1);
-ezcontourf(@(s, m) beta_prior_updf(prior_params, s, m), [0 120 0, 1]);
-colorbar;
-
-% % Evaluate sample statistics
-% % Stupidest integral ever
-% samps = zeros(100*100, 1);
-% ns = 1;
-% for s = linspace(0, 120, 100);
-%     for m = linspace(0, 1, 100);        
-%         samps(ns) = beta_prior_updf(prior_params, s, m);
-%         ns = ns + 1;
-%     end
-% end
+ezcontourf(@(s, m) beta_prior_updf(prior_params, s, m), [0 50 0.01 0.99]);
 
 
-% Focus on the 0-3 region
-figure(2);
-ezcontourf(@(s, m) beta_prior_updf(prior_params, s, m), [0 3 0, 1]);
-colorbar;
+%% Focus on small variance
+ezcontourf(@(s, m) beta_prior_updf(prior_params, s, m), [0 3 0.01 0.99]);
 
-%% Generate data and initialize with MLE
+%% Generate data
 true_m = 0.8;
-true_s = 10;
+true_s = 0.9;
+
 [a b] = beta_sm_to_ab(true_s, true_m);
 a = true_s * true_m;
 b = true_s * (1 - true_m);
@@ -34,6 +20,7 @@ SN = 1000;
 
 xs = betarnd(a * ones(SN,1), b * ones(SN, 1));
 
+%% Initialize with MLE
 mle_params = betafit(xs);
 [s m] = beta_ab_to_sm(mle_params(1), mle_params(2))
 
@@ -45,7 +32,7 @@ Nburn_in = 100;
 samps = zeros(Nsamps, 2);
 
 for n = 1:Nburn_in
-    [s m_star] = beta_posterior_mh(prior_params, s, m_star, xs, 0.01);
+    [s m_star] = beta_posterior_mh(prior_params, s, m_star, xs, 0.1);
 end
 
 samps(1,:) = [s m_star];
@@ -55,7 +42,7 @@ Nreject = 0;
 for n = 2:Nsamps
     s_old      = samps(n - 1,1);
     m_star_old = samps(n - 1,2);
-    [s_new m_star_new acc] = beta_posterior_mh(prior_params, s_old, m_star_old, xs, 0.01);
[…]

### 2012-11-01 src/run_mm.m

@@ -24,9 +24,12 @@ function [ samples ] = run_mm(iter_f, X, img, mm, niters, burn_in, sample_freq, 
         fprintf(1, 'Iter %d: time = %0.2f; speed = %.2f; change = %0.3f; loglike = %06.2f\n', ...
                 i, itertime, rate, change, mm.loglike);     
         
+        figure(1);
+        plot_point_overlay(img, mm, X);
+        drawnow;
+        
         if mod(i, sample_freq) == 0
-            plot_point_overlay(img, mm, X);
-            drawnow;
+
             
             if i > burn_in
                 samples = [samples mm];

### 2012-11-01 src/experiment/test_background_models.m
Gaussian diagnostic plots.

@@ -17,9 +17,11 @@ title('GMM beta(1,3) background, k = 6');
 figure;
 beta_loose_samps = gmm_gibbs(PHI1, f1, make_gmm_prior(PHI1, 6, beta_tight), 500, 200, 10);
 title('GMM beta(1,1.5) background, k = 6');
+
+%% Sigmoid only!
 figure;
 ss_sigmoid_samps = gmm_gibbs(PHI1, f1, make_gmm_prior(PHI1, 6, ss_sigmoid), 500, 200, 10);
-title('GMM background sigmoid centered at 0.2, compressed 10x, k =6');
+title('GMM background sigmoid centered at 0.2, compressed 10x, k=6');
 
 %% The shebang
 figure;

### 2012-11-01 src/experiment/test_beta_posterior_rnd.m

@@ -23,6 +23,11 @@ figure(2);
 ezcontourf(@(s, m) beta_prior_updf(prior_params, s, m), [0 3 0, 1]);
 colorbar;
 
+% Focus on the 2-5 region
+figure(2);
+ezcontourf(@(s, m) beta_prior_updf(prior_params, s, m), [2 5 0, 1]);
+colorbar;
+
 %% Generate data and initialize with MLE
 true_m = 0.8;
 true_s = 10;

### 2012-11-01 src/gmm_gibbs.m

@@ -1,0 +1,4 @@
+function [ samples ] = gmm_gibbs( varargin )   
+    samples = run_mm(@gmm_gibbs_iter, varargin{:});
+end
+

### 2012-11-01 src/make_gmm_prior.m

@@ -13,7 +13,7 @@ function [ gmm ] = make_gmm_prior( X, K, background_like )
                  'prior_mean', [100 100 0.75], ...
                  'prior_cov', cov(X), ...
                  'prior_dof', 4, ...
-                 'prior_scale', 1, ...
+                 'prior_scale', 0.1, ... % Contribute more precision??? UNDERSTAND THIS PARAMETER!
                  's_z', randsample(K, N, true), ...
                  'prior_mix', []);        
     
@@ -26,7 +26,12 @@ function [ gmm ] = make_gmm_prior( X, K, background_like )
     % determined by a very sophisticated method called "taking thresholds
     % and looking." For now, fit a method-of-moments Beta (notebook pp 23)    
         
-    gmm.background_like = background_like;
+    gmm.background_like = background_like;        
+    
+    gmm.h_diagnostic = figure;
+    set(gmm.h_diagnostic, 'Units', 'normalized', 'Position', ...
+        [0 0.7 0.9 0.4]);
+    gmm.plot_diagnostic = @plot_intensity_hists;
     
     % Initialize with k-means
     % Only valid if each feature is Gaussian!

### 2012-11-01 src/plot/plot_intensity_hists.m

@@ -1,0 +1,21 @@
+function [ h ] = plot_intensity_hists( mm, X )
+% Plot intensity histogram for each class.
+    h = mm.h_diagnostic;    
+    
+    for k = 1:mm.K        
+        subplot(2,mm.K,k);
+        idxs = mm.s_z == k;
+        hist(X(idxs,3));
+        title(['Class ' num2str(k) ' intensities']);   
+        
+        % GMM only. TODO: Generalize the parametric plotter.
+        if k > 1
+            subplot(2,mm.K,mm.K + k);        
+            ezplot(@(x) normpdf(x, mm.mean(k,3), sqrt(mm.cov(3,3,k))), linspace(0,1));
+            title(['Class ' num2str(k) ' fitted']);
+        end
+        
+    end
+        
+end
+

### 2012-11-01 src/plot/plot_point_overlay.m

@@ -14,10 +14,22 @@ function [ g ] = plot_point_overlay( img, gmm, PHI )
     end
     
     %h = imshow(img);
-    %hold on
-    g = image(overlay);    
+%    hold on
+    g = image(overlay);
+    hold on
+    for k = 1:gmm.K
+        text(gmm.mean(k,1), gmm.mean(k,2), num2str(k), ...
+             'FontSize', 30, 'FontWeight', 'bold');
+    end
+    hold off
+    
+    % Plot the 95% (2*sigma) confidence ellipse    
+%     for k = 1:mm.K
+%         [eig_vec eig_val] = eig(mm.cov(1:2,1:2,k));
+%     end
+%     
     %set(g, 'AlphaData', 1);    
-    hold off    
+%    hold off    
 
 end
 

### 2012-11-01 src/plot/plot_posterior_intensity_gaussians.m

@@ -1,0 +1,13 @@
+function [ h ] = plot_posterior_intensity_gaussians( mm, varargin )
+            
+    h = figure;
+    set(h, 'Units', 'normalized', 'Position', ...
+        [0 0.7 0.9 0.2]);    
+    
+    for k = 1:mm.K
+        
+    end
+    
+
+end
+

### 2012-11-01 src/run_mm.m

@@ -1,21 +1,32 @@
 function [ samples ] = run_mm(iter_f, X, img, mm, niters, burn_in, sample_freq, recovery_file )    
     
     [N, D] = size(X);
+    
+    % Initialize
     samples = [];
+    loglike = zeros(niters, 1);
     mm.n = accumarray(mm.s_z, ones(1,N));    
     
+    % Setup the plotter
+    h = figure;
+    set(h, 'Units', 'normalized', 'position', [0.25 0.4 0.6 0.4]);
+    
+    % Handle crashes
     if nargin >=8 && ~isempty(recovery_file)
         load(recovery_file);
         first = iter + 1;        
     else
         first = 1;
-    end
-        
+    end                
+    
     for i = first:niters
         tic
         z_old = mm.s_z;
-                
-        mm = iter_f(mm, X);
+        
+        % Iterate
+        mm = iter_f(mm, X);        
+        
+        % Timing and diagnostics
         change = mean(abs(z_old ~= mm.s_z));
         
         itertime = toc;
@@ -24,12 +35,30 @@ function [ samples ] = run_mm(iter_f, X, img, mm, niters, burn_in, sample_freq, 
         fprintf(1, 'Iter %d: time = %0.2f; speed = %.2f; change = %0.3f; loglike = %06.2f\n', ...
                 i, itertime, rate, change, mm.loglike);     
         
-        figure(1);
-        plot_point_overlay(img, mm, X);
-        drawnow;
+        % Save the log-likelihood
+        loglike(i) = mm.loglike;                
         
-        if mod(i, sample_freq) == 0
-
+        if mod(i, sample_freq) == 0                      
+            % Draw the figure
+            figure(h);
+            subplot(1,2,1);
+            plot(loglike(loglike ~= 0));
+            title('Log-likelihood vs iteration');
+            xlabel('Iteration');
+            ylabel('Log-likelihood');
+            
+            subplot(1,2,2);
+            plot_point_overlay(img, mm, X);
+            title(['K = ' num2str(mm.K) ' iter ' num2str(i) ' ll = ' num2str(mm.loglike)]);
+            
+            % Draw diagnostic images
+            if isfield(mm, 'plot_diagnostic')
+                % TODO: Proper handles
+                figure(mm.h_diagnostic);
+                mm.plot_diagnostic(mm, X);
+            end            
+            
+            drawnow;
             
             if i > burn_in
                 samples = [samples mm];

### 2012-11-01 src/util/struct_mean.m

@@ -6,11 +6,13 @@ function [ m ] = struct_mean( st )
     [~, N] = size(st);
     m = st(1);
     
-    fields = fieldnames(st);    
+    fields = fieldnames(st);   
     for i = 1:numel(fields)
+        f = fields{i};
         % Delete non-numeric fields
         if ~isnumeric(st(1).(f))
             fields(i) = []
+            i = i - 1;
         end
     end
     

### 2012-11-02 latex/paragraphs.tex
shitty paragraphs

@@
 
+\section{November 2, 2012}
+Initial Approaches
+
+We modelled a single slice of neural tissue with a finite Gaussian mixture model with full covariances.
+The feature vector $(x, y, i)$ consisted of the positional coordinates and pixel intensity.
+Each cluster represented the pixels belonging to a single cell, with a special background cluster to contain cell membrance and interstitial space.
+For this background, we experimented with both a triangular probability density $p(i) = 2 - 2*i$ and a sigmoid density of the form $p(i) = \frac{1}{Z} \left[ 1 - \left( 1 + \exp(-p(i-c)) \right)^-1 \right]$ with $p$ and $c$ denoting the precision (inverse scale) and location (the preimage of the midpoint of the range).
+Inference was performed with Gibbs sampling.
+
+Visually, our segmentation seems reasonable, although cells whose interior intensity varies significantly tend to be oversegmented, and the cell boundary model is insufficient because the boundary intensity varies locally (see slides).
+A more accurate background model must be adaptive.
+
+A more serious problem is that our Gaussian intensity model is insufficienltly compact: the standard deviations are far too high to accurately model the bounded $[0, 1]$ intensity range.
+To rectify this, we experimented with modelling the third dimension as a Beta distribution independently of the first two dimensions, but still dependent on class labels.
+We implemented a Metropolis-Hastings sampler using a lognormal proposal and a prior based on \cite{Bouguila2006}.
+We modified the prior to take the form $$ \left[ 1 - exp(-( (s - 0.5)^2)) (1 - exp(-( (a - 1)^2)))(1 - exp(-( (b - 1).^2))) (1 - exp(-d * ((s - 2).^2 + (m - 0.5).^2))) exp(-r / (s.^2 .* m .* (1 - m)) - (k * s.^2) / 2) \right]$$.
+This avoids distributions with $a, b < 1$, which lead to unbounded U-shaped distributions, as well as $a = b = 1$, the uniform distribution.
+Unfortunately, our experiment with this model reverted to axis-aligned decision boundaries.
+It appears that spatio-intensity correlations are essential.
+
 \bibliographystyle{plain}

### 2012-11-02 src/make_gmm_prior.m

@@ -12,15 +12,20 @@ function [ gmm ] = make_gmm_prior( X, K, background_like )
                  'nY', nY, ...
                  'prior_mean', [100 100 0.75], ...
                  'prior_cov', cov(X), ...
-                 'prior_dof', 4, ...
-                 'prior_scale', 0.1, ... % Contribute more precision??? UNDERSTAND THIS PARAMETER!
+                 'prior_dof', 1000, ...
+                 'prior_scale', 1e-2, ... % Contribute more precision??? UNDERSTAND THIS PARAMETER!
                  's_z', randsample(K, N, true), ...
-                 'prior_mix', []);        
+                 'prior_mix', []);
     
+    % Tighter intensities
+    %gmm.prior_cov(3,3) = 0.01 * gmm.prior_cov(3,3);             
+             
     % Weaker Dirichlet prior (\sum \alpha_k = 1)
     gmm.prior_mix = zeros(K, 1);
     gmm.prior_mix(1) = 0.4;
-    gmm.prior_mix(2:end) = 0.6 / (K - 1);        
+    gmm.prior_mix(2:end) = 0.6 / (K - 1);
+    
+    gmm.prior_mix = 0.1 * N * gmm.prior_mix;
     
     % All points with intensity < 0.3 are considered candidates for
     % determined by a very sophisticated method called "taking thresholds
@@ -30,11 +35,12 @@ function [ gmm ] = make_gmm_prior( X, K, background_like )
     
     gmm.h_diagnostic = figure;
     set(gmm.h_diagnostic, 'Units', 'normalized', 'Position', ...
-        [0 0.7 0.9 0.4]);
+        [0 0.7 1 0.4]);
     gmm.plot_diagnostic = @plot_intensity_hists;
     
-    % Initialize with k-means
-    % Only valid if each feature is Gaussian!
-    % [gmm.z, gmm.mu_mean] = kmeans(X, K);
+    % Initialize with k-means, but don't set background
+    % Only valid if each feature is Gaussian!    
+    %[gmm.s_z, ~] = kmeans(X, K - 1);
+    %gmm.s_z = gmm.s_z + 1;
         
 end

### 2012-11-02 src/make_sigmoid_pdf.m

@@ -1,0 +1,6 @@
+function [ ss ] = make_sigmoid_pdf( center, precision )
+    pre_ss = @(x) 1-(1./(1+exp(-precision * (x - center))));
+    Z = quad(pre_ss, 0, 1);
+    ss = @(x) 1/Z * pre_ss(x);
+end
+

### 2012-11-02 src/plot/plot_intensity_hists.m

@@ -5,15 +5,21 @@ function [ h ] = plot_intensity_hists( mm, X )
     for k = 1:mm.K        
         subplot(2,mm.K,k);
         idxs = mm.s_z == k;
-        hist(X(idxs,3));
-        title(['Class ' num2str(k) ' intensities']);   
+        x = X(idxs,3);
+        hist(x);        
+        title([' N = ' num2str(sum(idxs)) ...
+               ' m = ' num2str(mean(x), 2) ...
+               ' sd = ' num2str(std(x, 1), 2)]);   
         
         % GMM only. TODO: Generalize the parametric plotter.
         if k > 1
             subplot(2,mm.K,mm.K + k);        
             ezplot(@(x) normpdf(x, mm.mean(k,3), sqrt(mm.cov(3,3,k))), linspace(0,1));
-            title(['Class ' num2str(k) ' fitted']);
+            title(['fitted m = ' num2str(mm.mean(k,3), 2) ...
+                   ' sd = ' num2str(sqrt(mm.cov(3,3,k)), 2)]);
         end
+        
+        mm.cov
         
     end
         

### 2012-11-02 src/plot/plot_point_overlay.m

@@ -17,7 +17,7 @@ function [ g ] = plot_point_overlay( img, gmm, PHI )
 %    hold on
     g = image(overlay);
     hold on
-    for k = 1:gmm.K
+    for k = 2:gmm.K
         text(gmm.mean(k,1), gmm.mean(k,2), num2str(k), ...
              'FontSize', 30, 'FontWeight', 'bold');
     end

### 2012-12-07 latex/submission.tex
started submission paper

--- 2012-11-02 latex/paragraphs.tex
@@
 \documentclass{article}
+\usepackage{natbib}
 
 \begin{document}
-\title{Weekly Paragraphs}
-\author{Kui Tang}
+
+\title{A Bayesian Nonparametric Approach to Neural Image Segmentation}
+
+\author{Kui Tang and Frank Wood}
 \date{\today}
 \maketitle
+\begin{abstract}
+We develop a semi-supervised method to produce a probabilistic segmentation of neurons in electron micrograph images of brain tissue.
+Unlike previous approaches, which generate , BLAH BLAH.
+To do so, we model the distribution of neurons in an image as an infinite Gaussian mixture model.
+Each neuron's pixels are characterized by likelihood function that is a combination of multivariate Gaussian and naive Bayes features, while the total and location of neurons are inferred automatically, sidestepping many model-selection or hypothesis-combination difficulties in current approaches.
+Importantly, our model can incorporate both features derived directly from the data and features derived from the output of a supervised learner.
+Using data collected by Kasthuri et al., we show that our model can obtain plausible segmentations with no training data, while receiving a significant boost in its presence.
+
+\section{Introduction}
+A new branch of computational neuroscience, \emph{connectomics}, aims to reconstruct the physical location and circuit connectivity of the brain.
+Recent advances in electron microscopy have produced images of 10 nm in the lateral plane and 50 nm in the vertical plane, sufficient detail to capture most of even the smallest neural processes~\citep{Briggman2006}.
+The scale of the data is profound; the human brain is estimated to contain 100 billion neurons connected by up to a quadrillion synapses~\citep{Kasthuri2010}.
+Like the human genome project before, it is clear that automated methods are needed to map the connectome.
+
+All published connectomes have been manually labelled.
+White et al.
+published the first connectome of \emph{C.
+elegans} in 1986, marking axons, dendrites, and cell bodies in each frame and tracing the labels through frames~\cite{White12111986}.
+More recent work uses computer assistance but fundamentally rely on a skilled human's identification of neural bodies~\cite{Jarrell2012,Bock2011Roberts2011}.
+Several authors have developed efficient semi-automated methods \cite{Roberts2011,Unger2009}.
+Unfortunately, current baseline speeds of computer-assisted manual tracing would still take 10,000 person-years to construct a mouse cortical column~\cite{Briggman2006}.
+Therefore, in the rest of the paper, we focus on automated techniques.
+
+\section{Related Work}
+Automated techniques generally begin by segmenting a 2D slice based on cell membrane boundaries---a highly distinctive feature.
+
+\section{Methods}
+We model an image as an infinite Gaussian mixture model.
+First proposed by Rasmussen~\citep{Rasmussen2000}, this model extends the classical finite Gaussian mixture model with $K$ classes to estimate $K$ automatically from the data.
+
+\section{Results}
+
+\section{Discussion}
+
+\end{abstract}
 \section*{October 3, 2012}
 To appear in: Related Work
-
-All known connectomes reply on manual labels.
-White et al.
-published the first connectome of \emph{C.
-elegans} in 1986, marking axons, dendrites, and cell bodies in each frame and tracing the labels through frames \cite{White12111986}.
-More recent work uses computer assistance but fundamentally rely on a skilled human's identification of neural bodies \cite{Jarrell2012,Bock2011Roberts2011}.
 
@@
 Second, due to the physical cutting process, images may be warped from slice to slice \cite{Koshevoy2006}.
-
-Semi-automated methods take small amounts of user input and then use the morphology of the image segment entire neural processes.
-Marker-Controlled watersheds \cite{Gonzales2008} is a classic technique, requiring the user to mark background and foreground.
-Roberts et al.
-\cite{Roberts2011} require a "scribble" inside a neuron on the first and last slices.
-At each slice, their algorithm segments the neuron by minimizing the anisotropic weighted total variation \cite{Unger2009} and propagating the inferred boundary as soft constraints for the next layer.
-
-Unfortunately, current baseline speeds of computer-assisted manual tracing would take 10,000 person-years to construct a mouse cortical column, suggesting that fully automatic methods are necessary \cite{Briggman2006}.
 

### 2012-12-07 src/feature_map.m

@@ -2,18 +2,30 @@ function [ PHI ] = feature_map( img )
     [X, Y] = size(img);
     N = X * Y;
     
-    D = 3;
+    % Number of features
+    D = 4;
+    
+    res = 72;
     
     PHI = zeros(N, D);
+    
+    bg_intx = radonLikeFeatures(img, true, res, @(scanseg) min(scanseg));
+    m_bg_intx = mean(bg_intx,3);
+    
+    % Scale to 0-1
+    range = max(m_bg_intx(:)) - min(m_bg_intx(:));
+    m_bg_intx = (m_bg_intx - min(m_bg_intx(:))) ./ range;
+    
     n = 1;
     for y = 1:Y
         for x = 1:X
-            PHI(n, :) = [x y img(x,y)];
+            PHI(n, :) = [x y img(x,y) m_bg_intx(x,y)];
             n = n+1;
         end
     end
     
     % Censor zero intensities (causes some problems for Gamma)
     PHI(:,3) = max(eps, PHI(:,3));
+    PHI(:,4) = max(eps, PHI(:,4));
+  
 end
-

### 2012-12-07 src/make_mdp_prior.m

@@ -4,13 +4,18 @@ function [ mdp ] = make_mdp_prior( X )
     nY = max(X(:,2));
     
     %% Parameter priors
-    prior_mean = [100 100 0.75];
+    prior_mean = [100 100 0.75 0.4]; % mean of the nonzero entries
     
     prior_cov  = cov(X);
+    
     v_intensity = prior_cov(3,3);
-    prior_cov(3,:) = zeros(3, 1);
-    prior_cov(:,3) = zeros(1, 3);   
-    prior_cov(3,3) = 0.1 * v_intensity;
+    v_bg_radon  = prior_cov(4,4);
+    
+    prior_cov(3:4,:) = zeros(2, 4);
+    prior_cov(:,3:4) = zeros(4, 2);    
+        
+    prior_cov(3,3) = 0.5 * v_intensity;
+    prior_cov(4,4) = 0.5 * v_bg_radon;
     
     background = GammaGamma(1, 6, 2);
     
@@ -21,13 +26,18 @@ function [ mdp ] = make_mdp_prior( X )
     
     %% Initialize cluster assignments
     % Wood: assign them uniformly at random
-    cluster_assigns = initialize_one_cluster_and_background(0.99, 0.4, X);
+    cluster_assigns = initialize_one_cluster_and_background(0.99, 0.5, X);
     
     %% Construct objects
     nw_prior = NormalWishart(prior_mean, prior_cov, prior_dof, prior_n);
     mdp = MDP(concentration, X, cluster_assigns, nw_prior);
     
-    mdp.cluster_likes{1} = ProductDistribution(1, 2, Uniform(size(X, 1)), 3, 3, background);
+    
+    % Hack: one uniform for (x,y) and another uniform for dimension 4 (the
+    % background Radon response) ... ; NO IT'S BORKED!
+    mdp.cluster_likes{1} = ProductDistribution(1, 2, Uniform(size(X, 1)), ...
+                                               3, 3, background, ...
+                                               4, 4, Uniform(1)); % Because domain is [0,1]
     mdp.refit(1);
     %[particles, weights] = smc_init(mdp, ProductDistribution(1, 2, Uniform(size(X, 1)), 3, 3, background));
     

### 2012-12-07 src/neal3_iter.m

@@ -12,7 +12,15 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
     prior_pred_like = prior.pred_like(mdp.X); % col vector
     
     ll = 0;
+    
+    actual_n = 0;
+    
     for n = randperm(mdp.N)
+        actual_n = actual_n + 1;
+        if mod(actual_n, 1000) == 0
+            disp(['actual_n = ' num2str(actual_n)]);
+        end
+        
         k_old = mdp.remove_point(n);
         % Removed a point, so we must refit the posterior
         %

### 2012-12-07 src/plot/plot_img.m

@@ -8,18 +8,23 @@ function [ h ] = plot_img( h, n, loglike_vec, X, img, mdp )
     
     figure(h);
     
-    subplot(1,3,1);
+    subplot(1,4,1);
     nz_loglike = loglike_vec(loglike_vec ~= 0);
     plot(nz_loglike);
     title('Log-likelihood vs iteration');
     xlabel('Iteration');
     ylabel('Log-likelihood');
 
-    subplot(1,3,2);
+    subplot(1,4,2);
     plot_point_overlay(img, mdp, X);
     title(['K = ' num2str(mdp.n_clusters) ' iter ' num2str(n) ' ll = ' num2str(nz_loglike(end))]);
 
-    subplot(1,3,3);
-    imshow(img); 
+    subplot(1,4,3);
+    imshow(img);
+    
+    % The feature map
+    subplot(1,4,4);
+    fmap = reshape(X(:,4), size(img));
+    imshow(fmap);
 
 end

### 2012-12-07 src/plot/plot_intensity_hists.m

@@ -2,6 +2,7 @@ function [ h ] = plot_intensity_hists( h, mdp, X )
 % Plot intensity histogram for each class.   
     
     K = mdp.n_clusters;
+    xs = linspace(eps, 1, 100);
     
     for k = 1:K
         figure(h);
@@ -9,13 +10,13 @@ function [ h ] = plot_intensity_hists( h, mdp, X )
         x = X(idxs, 3);
         
         % Top row plots the empirical histograms
-        subplot(2, K, k), hist(x);
+        subplot(3, K, k), hist(x);
         title(['n = ' num2str(sum(idxs)) ...
                ' m = ' num2str(mean(x), 2) ...
                ' sd = ' num2str(std(x, 1), 2)]);
                    
-        subplot(2, K, K + k);
-        xs = linspace(eps, 1, 100);
+        subplot(3, K, K + k);
+        
         % DAMN THE SPECIAL CASING!
         if k == 1
             plot(xs, mdp.cluster_likes{1}.pdfs{2}.pred_like_scalar(xs));            
@@ -25,5 +26,15 @@ function [ h ] = plot_intensity_hists( h, mdp, X )
             dof = mdp.cluster_likes{k}.pred_dof;
             plot(xs, tpdf( (xs - m) / sd, dof ));            
         end
+        
+        subplot(3, K, 2*K + k);
+        
+        if k ~= 1
+            m   = mdp.cluster_likes{k}.post_mean(3);            
+            sd  = sqrt(mdp.cluster_likes{k}.pred_cov(3, 3));
+            dof = mdp.cluster_likes{k}.pred_dof;
+            plot(xs, tpdf( (xs - m) / sd, dof ));            
+        end            
+        
     end    
 end

### 2012-12-07 src/plot/plot_point_overlay.m

@@ -1,4 +1,4 @@
-function [ g ] = plot_point_overlay( img, mdp, X )
+function [ g ] = plot_point_overlay( img, mdp, X, decorate )
 % Plot point assignments and 95% confidence ellipses.
 % img - a grayscale image (array) of the original data
 % mdp - an MDP whose cluster_likes have a .post_mean and .post_cov
@@ -24,31 +24,33 @@ function [ g ] = plot_point_overlay( img, mdp, X )
     g = image(overlay);
     hold on
     
-    for k = 2:mdp.n_clusters
-        % We'll cheat here and plot the error ellipses of the PREDICTIVE
-        % distributions. This is visually accurate: it shows how likely a
-        % point is to be classified in a given cluster.
-        %
-        % A more correct plot, encapsulating the uncertainty of the model
-        % itself, would be ellipses of the posterior mean.
-        
-        cl = mdp.cluster_likes{k};        
-        mean_x = cl.pred_mean(1);
-        mean_y = cl.pred_mean(2);
-                     
-        cc = c(k, :);
-        
-        text(mean_x, mean_y, num2str(k), ...
-             'FontSize', 30, ...
-             'FontWeight', 'bold', ...
-             'BackgroundColor', 'w', ...
-             'Color', cc);
-        
-        ee = error_ellipse(cl.pred_cov(1:2,1:2), cl.pred_mean(1:2), 'conf', 0.95);
-        set(ee, ...            
-            'Color', 'w', ...
-            'LineWidth', 2);        
-    end    
+    if nargin > 3 && decorate
+        for k = 2:mdp.n_clusters
+            % We'll cheat here and plot the error ellipses of the PREDICTIVE
+            % distributions. This is visually accurate: it shows how likely a
+            % point is to be classified in a given cluster.
+            %
+            % A more correct plot, encapsulating the uncertainty of the model
+            % itself, would be ellipses of the posterior mean.
+
+            cl = mdp.cluster_likes{k};        
+            mean_x = cl.pred_mean(1);
+            mean_y = cl.pred_mean(2);
+
+            cc = c(k, :);
+
+            text(mean_x, mean_y, num2str(k), ...
+                 'FontSize', 30, ...
+                 'FontWeight', 'bold', ...
+                 'BackgroundColor', 'w', ...
+                 'Color', cc);
+
+            ee = error_ellipse(cl.pred_cov(1:2,1:2), cl.pred_mean(1:2), 'conf', 0.95);
[…]

### 2012-12-07 src/run_mdp.m

@@ -2,7 +2,7 @@ function [ samples ] = run_mdp( iter_f, state, params, X, img, mdp, niters, burn
 
     loglike = zeros(niters, 1);    
     img_h = figure(1);    
-    set(img_h, 'Units', 'normalized', 'position', [0.1 0.1 0.9 0.4]);
+    set(img_h, 'Units', 'normalized', 'position', [0.1 0.1 1 0.4]);
     clf;
     
     hist_h = figure(2);
@@ -10,7 +10,7 @@ function [ samples ] = run_mdp( iter_f, state, params, X, img, mdp, niters, burn
     clf;    
     
     for n = 1:niters
-        tic
+        tic                
         
         cluster_assigns_old = mdp.cluster_assigns;
         [mdp, state, ll] = iter_f(mdp, state, params);

### 2012-12-07 src/scrape/pre_ann.m

@@ -9,14 +9,12 @@
 F = fullfile(DATA_PREFIX, 'kat11info.h5');
 xyres = 1000;
 zres  = 100;
-
-%% Download info file
 secret;
 KBASE = ['http://openconnecto.me/emca/' KAT11_TOKEN '/'];
+PUBBASE = 'http://openconnecto.me/emca/kasthuri11/';
+
+%% Download info file
 urlwrite([KBASE 'projinfo/'], F);
-
-
-%% Parse the h5 file
 proj_res      = h5read(F, '/PROJECT/RESOLUTION');
 proj_type     = h5read(F, '/PROJECT/TYPE');
 proj_res_str  = num2str(proj_res);
@@ -43,17 +41,59 @@ slice_range   = h5read(F, '/DATASET/SLICERANGE');
 % type kat11_ann.h5
 
 %% Brute force
+
+%% Download bounding boxes
 % This list is discovered empirically. i.e. watching which requests fail.
-topid = 6075
+topid = 6075;
 exclude = [2146 2208 2216 2218];
 
-annoids = 1:topid;
-annoids(exclude) = [];
+annoid = 1:topid;
+annoid = annoid';
+annoid(exclude) = [];
 
-% Chunk our requests
+Nids = length(annoid);
+
+tmp = 'bbox_tmp.h5';
+
+res      = zeros(Nids, 1);
+xyz_off  = zeros(Nids, 3);
+xyz_dim  = zeros(Nids, 3);
+conf     = zeros(Nids, 1);
+kvpairs  = cell(Nids, 1);
+
+for i = 1:Nids
+    a = annoid(i);
+        
+    urlwrite([KBASE '/' num2str(a) '/boundingbox/'], tmp);
+    head = ['/' num2str(a) '/'];
+    
+    res(i)       = h5read(tmp, [head 'RESOLUTION']);
+    xyz_off(i,:) = h5read(tmp, [head 'XYZOFFSET']);
+    xyz_dim(i,:) = h5read(tmp, [head 'XYZDIMENSION']);
+    conf(i)      = h5read(tmp, [head 'METADATA/CONFIDENCE']);
+    kvpairs(i)   = h5read(tmp, [head 'METADATA/KVPAIRS']);
+    disp(num2str(a));
+end
+
+annotation_bboxen = dataset(annoid, res, xyz_off, xyz_dim, conf, kvpairs);
+save('annotation_bboxen.mat', 'annotation_bboxen');
+
+%% Download the cutouts from the server
[…]

### 2012-12-07 src/startup.m

@@ -3,6 +3,7 @@ addpath(fullfile(pwd, 'prob'));
 addpath(fullfile(pwd, 'experiment'));
 addpath(fullfile(pwd, 'plot'));
 addpath(fullfile(pwd, 'util'));
+addpath(fullfile(pwd, 'radonLikeFeatures'));
 
 DATA_PREFIX = fullfile(pwd, '../data/kasthuri11/');
 
@@ -11,7 +12,8 @@ local_startup;
 
 S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
 %A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
-f1 = S1(1000:1199,1000:1199);
+%f1 = S1(1000:1199,1000:1199);
+f1 = S1(1500:1699,1000:1199);
 %a1 = A1(1000:1199,1000:1199);
 PHI1 = feature_map(f1);
 

### 2012-12-08 latex/submission.tex
radon features and background constraint

@@
-\documentclass{article}
-\usepackage{natbib}
+\documentclass[draft]{article}
+\usepackage{fullpage}
+
+\usepackage[draft]{fixme}
 
 \begin{document}
+
+\listoffixmes
 
@@
 \begin{abstract}
-We develop a semi-supervised method to produce a probabilistic segmentation of neurons in electron micrograph images of brain tissue.
-Unlike previous approaches, which generate , BLAH BLAH.
-To do so, we model the distribution of neurons in an image as an infinite Gaussian mixture model.
-Each neuron's pixels are characterized by likelihood function that is a combination of multivariate Gaussian and naive Bayes features, while the total and location of neurons are inferred automatically, sidestepping many model-selection or hypothesis-combination difficulties in current approaches.
-Importantly, our model can incorporate both features derived directly from the data and features derived from the output of a supervised learner.
+We \fxfatal*{we don't really develop; what the best verb?}{develop} a Bayesian nonparametric method to segment neurons in electron micrograph images of brain tissue.
+Our model imposes spatial locality on arbitrary features, automatically determines the number of cells in an image, and can be used in both unsupervised and semi-supervised settings.
+We model the distribution of neurons in an image as an infinite Gaussian mixture model.
+Each neuron's pixels are characterized by likelihood function that is a combination of multivariate Gaussian and naive Bayes features.
 Using data collected by Kasthuri et al., we show that our model can obtain plausible segmentations with no training data, while receiving a significant boost in its presence.
@@
 A new branch of computational neuroscience, \emph{connectomics}, aims to reconstruct the physical location and circuit connectivity of the brain.
-Recent advances in electron microscopy have produced images of 10 nm in the lateral plane and 50 nm in the vertical plane, sufficient detail to capture most of even the smallest neural processes~\citep{Briggman2006}.
-The scale of the data is profound; the human brain is estimated to contain 100 billion neurons connected by up to a quadrillion synapses~\citep{Kasthuri2010}.
+Recent advances in electron microscopy have produced images of 10 nm in the lateral plane and 50 nm in the vertical plane, sufficient detail to capture most of even the smallest neural processes~\cite{Briggman2006}.
+The scale of the data is profound; the human brain is estimated to contain 100 billion neurons connected by up to a quadrillion synapses~\cite{Kasthuri2010}.
 Like the human genome project before, it is clear that automated methods are needed to map the connectome.
@@
 Unfortunately, current baseline speeds of computer-assisted manual tracing would still take 10,000 person-years to construct a mouse cortical column~\cite{Briggman2006}.
-Therefore, in the rest of the paper, we focus on automated techniques.
+Therefore, henceforth we focus on automated techniques.
 
 \section{Related Work}
+\fxfatal{Write topic sentence, summarize your current literature review, and add paragraphs about feature engineering and graph pathfinding methods. Cite Malik's normalized cuts.}
 Automated techniques generally begin by segmenting a 2D slice based on cell membrane boundaries---a highly distinctive feature.
 
-\section{Methods}
-We model an image as an infinite Gaussian mixture model.
-First proposed by Rasmussen~\citep{Rasmussen2000}, this model extends the classical finite Gaussian mixture model with $K$ classes to estimate $K$ automatically from the data.
+\section{Model and Methods}
+
+\fxfatal{Obtain more information about this dataset, particularly stain protocol, and put in the right place.}
+We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \mbox{ nm}^3$ collected by \fxfatal{Unpublished; to cite}{Kasthuri et al.}
+
+\fxfatal{Frank's notes: Think of your initial reaction of "wth are all these features?" What?! Spatial coordinates for GMM?}
+
+We model each pixel as feature vector $\mathbf{x}_n = (x_n, y_n, \phi_{1n}, \ldots, \phi_{mn})$.
+The first two dimensions encode the $x$ and $y$ coordinates of the pixel in order to provide spatial smoothing, while $\phi_{\cdot n}$ represent feature responses.
+Our model can handle arbitrary features, but in this paper, we will put pixel intensity in $\phi_{1\cdot} \in [0,1]$.
+Because cell interiors contain diverse structures, intensities values within a cell vary widely, so relying on $\phi_{1\cdot}$ alone oversegments.
+To reduce oversegmentation, we compute a cel boundary Radon-like feature for $\phi_{1\cdot}$, which homogenizes pixel intensities within a given cell~\cite{Kumar2010}.
+One computes Radon-like functions by taking line integrals of a kernel function along various scanning orientations, broken piecewise at image discontinuities.
+By varying the kernel, different structures and boundaries are enhanced.
+We follow Kumar's approach of averaging the integrals along each scan orientation to obtain a scalar response at each pixel.
+Finally, recall that our background class ($k = 1$) contains all pixels belonging on the cell membrane or in intracellular space.
+We penalize points
+
+%, the cell Background Radon-like feature response in $\phi_{2\cdot}$, and the length of the shortest cluster path from $(
 

### 2012-12-08 src/MDP.m

@@ -15,6 +15,7 @@ classdef MDP < matlab.mixin.Copyable
         N = 0;
         data_dim
         X
+        misc_data        
         prior
         
         % N x 1 vector with range [0, num_clusters]. 0 represents
@@ -29,7 +30,7 @@ classdef MDP < matlab.mixin.Copyable
     end
     
     methods
-        function o = MDP(concentration_or_rhs, X, cluster_assigns, prior)
+        function o = MDP(concentration_or_rhs, X, misc_data, cluster_assigns, prior)
             if nargin == 1 % copy constructor
                 % Assign the arrays
                 rhs = concentration_or_rhs;
@@ -37,6 +38,7 @@ classdef MDP < matlab.mixin.Copyable
                 o.N = rhs.N;
                 o.data_dim = rhs.data_dim;
                 o.X = rhs.X;
+                o.misc_data = misc_data;
                 
                 o.cluster_assigns = rhs.cluster_assigns;
                 o.n_clusters = rhs.n_clusters;
@@ -50,8 +52,9 @@ classdef MDP < matlab.mixin.Copyable
                 % cluster_assigns must be all specified, for now.            
                 [o.N, o.data_dim] = size(X);
                 o.X = X;
+                o.misc_data = misc_data;
                 o.concentration = concentration_or_rhs;                        
-                o.prior = prior;
+                o.prior = copy(prior);
                 
                 nzidxs = cluster_assigns > 0;
                 if sum(nzidxs > 0)                
@@ -63,8 +66,14 @@ classdef MDP < matlab.mixin.Copyable
 
                     for k = 1:o.n_clusters
                         idxs = cluster_assigns == k;
-
+                                                
                         o.cluster_likes{k} = copy(prior);
+                        
+                        % TEST -- did we actually copy?
+                        prior.pdfs{2}.data_n = 999;
+                        assert(o.cluster_likes{k}.pdfs{2}.data_n ~= 999);
+                        prior.pdfs{2}.data_n = 0;
+                        
                         o.cluster_likes{k}.fit(X(idxs,:));                        
                     end
                 end

### 2012-12-08 src/NormalGamma.m

@@ -1,0 +1,194 @@
+classdef NormalGamma < matlab.mixin.Copyable & OnlineDistribution
+%NormalGamma Scalar version of NormalWishart
+%
+% References:
+% [1] Kevin Murphy, Conjugate Bayesian analysis of the Gaussian
+%     distribution, www.cs.ubc.ca/~murphyk/Papers/bayesGauss.pdf
+    
+    properties
+        prior_mean, prior_n, prior_shape, prior_rate;        
+        data_n
+        
+        post_mean, post_n, post_shape, post_rate;
+        pred_mean, pred_dof, 
+        
+        post_mean,      post_chol, post_dof, post_n;
+        
+        pred_mvtparams, pred_chol, pred_dof, pred_n
+    end
+    
+    properties (Dependent)
+        pred_mean, post_cov, pred_cov
+    end            
+    
+    methods
+        function pm = get.pred_mean(o)
+            pm = o.post_mean;
+        end
+        
+        function poc = get.post_cov(o)
+            poc = (o.post_chol' * o.post_chol);
+        end
+        
+        function prc = get.pred_cov(o)
+            prc = (o.pred_chol' * o.pred_chol);
+        end
+        
+        function o = NormalWishart(prior_mean_or_rhs, prior_cov, prior_dof, prior_n)
+            if nargin == 1 % copy constructor
+                % REMEMBER ME!!! Standard trick to copy fields. 
+                fns = properties(prior_mean_or_rhs);
+                for i=1:length(fns)
+                    o.(fns{i}) = rhs.(fns{i});
+                end
+            else                
+                o.prior_mean = prior_mean_or_rhs;
+                o.prior_cov  = prior_cov;
+                o.prior_dof  = prior_dof;
+                o.prior_n    = prior_n;
+
+                o.dim = size(prior_cov, 1);               
+                
+                % IMPORTANT: Make the distribution ready for prediction and
+                % update
+                o.fit([]);
+            end
+        end                
+        
+        % All of the Bayesian update equations take constant time (wrt d).
+        % Online updating really just updates the data covariance.
+        %
+        % Make sure the data covariance computation is dominating the
+        % runtime before going online.
+        %
+        % The real kicker is are likelihood computation, methinks.
+        function fit(o, X)
+% fit(X) computes posterior and predictive parameters. Observations in rows
+            [o.data_n, dim] = size(X);
+            
[…]

### 2012-12-08 src/ProductDistribution.m

@@ -1,4 +1,4 @@
-classdef ProductDistribution < OnlineDistribution
+classdef ProductDistribution < matlab.mixin.Copyable & OnlineDistribution
 % Applies independent probability models to distinct dimensions of X    
     
     properties
@@ -8,31 +8,63 @@ classdef ProductDistribution < OnlineDistribution
         N;
     end
     
+    methods(Access = protected)
+        % Override copyElement method:
+        function cpObj = copyElement(obj)
+            % Make a shallow copy of all four properties
+            cpObj = copyElement@matlab.mixin.Copyable(obj);
+            
+            % Make a deep copy of each Distribution
+            for n = 1:length(obj.pdfs)
+                cpObj.pdfs{n} = copy(obj.pdfs{n});
+            end            
+        end
+    end
+    
     methods
         function o = ProductDistribution(varargin)
             % o = ProductDistribution(startidx, endidx, pdf, ...)
             % Construct a distribution that splits its input X into
             % components
             
-            assert(mod(length(varargin), 3) == 0, 'varargin must be multiple of 3!');
-            o.N = length(varargin) / 3;
-            
-            o.idxranges = cell(o.N, 1);
-            o.pdfs      = cell(o.N, 1);            
-            
-            n = 1;
-            k = 1;
-            while k < length(varargin)
-                o.idxranges{n} = varargin{k}:varargin{k+1};
-                o.pdfs{n}      = varargin{k+2};                
-                k = k + 3;
-                n = n + 1;
+            % Deep copy constructor
+            if nargin == 1
+                rhs = varargin{1};
+                o.N = rhs.N;
+                o.N = rhs.N;
+                o.idxranges = rhs.idxranges;
+                o.pdfs      = cell(o.N, 1);
+                for n = 1:o.N
+                    o.pdfs{n} = copy(rhs.pdfs{n});
+                end
+            else            
+                assert(mod(length(varargin), 3) == 0, 'varargin must be multiple of 3!');
+                o.N = length(varargin) / 3;
+
+                o.idxranges = cell(o.N, 1);
+                o.pdfs      = cell(o.N, 1);            
+
+                n = 1;
+                k = 1;
+                while k < length(varargin)
+                    o.idxranges{n} = varargin{k}:varargin{k+1};
+                    o.pdfs{n}      = varargin{k+2};                
+                    k = k + 3;
+                    n = n + 1;
+                end
             end
         end
         
[…]

### 2012-12-08 src/experiment/test_mdp.m

@@ -1,6 +1,6 @@
 %% Prior!
 clear mdp
-mdp = make_mdp_prior(PHI1);
+mdp = make_mdp_prior(PHI1, misc_data);
 
 %% Run!
 run_mdp(@neal3_iter, [], [], PHI1, f1, mdp, 1000, 0, 1);

### 2012-12-08 src/feature_map.m

@@ -1,4 +1,4 @@
-function [ PHI ] = feature_map( img )
+function [ PHI, misc_data ] = feature_map( img )
     [X, Y] = size(img);
     N = X * Y;
     
@@ -10,11 +10,23 @@ function [ PHI ] = feature_map( img )
     PHI = zeros(N, D);
     
     bg_intx = radonLikeFeatures(img, true, res, @(scanseg) min(scanseg));
-    m_bg_intx = mean(bg_intx,3);
+    m_bg_intx = mean(bg_intx,3);        
     
     % Scale to 0-1
     range = max(m_bg_intx(:)) - min(m_bg_intx(:));
     m_bg_intx = (m_bg_intx - min(m_bg_intx(:))) ./ range;
+    %figure; imshow(m_bg_intx);
+        
+    % argument 2 is use_raw_imo
+    bd_intx = radonLikeFeatures(img, false, res, @(scanseg) mean(scanseg));
+    m_bd_intx = mean(bd_intx,3);
+    
+    range = max(m_bd_intx(:)) - min(m_bd_intx(:));
+    m_bd_intx = (m_bd_intx - min(m_bd_intx(:))) ./ range;
+    %figure; imshow(m_bd_intx);
+    
+    misc_data = struct();    
+    [misc_data.segments, misc_data.slic_inds, misc_data.edge_G] = make_diff_graph(m_bd_intx, 10, 0.1);    
     
     n = 1;
     for y = 1:Y
@@ -27,5 +39,8 @@ function [ PHI ] = feature_map( img )
     % Censor zero intensities (causes some problems for Gamma)
     PHI(:,3) = max(eps, PHI(:,3));
     PHI(:,4) = max(eps, PHI(:,4));
+    
+    % Fifth feature is dynamic
+    PHI(:,5) = zeros(N, 1);
   
 end

### 2012-12-08 src/make_diff_graph.m

@@ -1,0 +1,78 @@
+function [ segments, slic_inds, edge_G ] = make_diff_graph( img, slic_npixels, slic_reg )
+% make_diff_graph - returns sparse adjacency graph with pixel-difference edges
+%
+% img          - [X Y] image matrix
+% slic_npixels - region size passed to vl_slic
+% slic_reg     - regularizer passed to vl_slic
+%
+% segments - [X Y] matrix mapping pixels to superpixel ids
+% edge_G   - sparse [Nslic Nslic] lower triangular graph. Two superpixels
+%            are adjacent if there exist two adjacent underlying pixels.
+
+    [X, Y] = size(img);
+    
+    % vl_slic returns 0-indexed; we want 1-indexed.
+    segments = vl_slic(single(img), slic_npixels, slic_reg) + 1;        
+    
+    assert(min(segments(:)) > 0);
+    Nslic = max(segments(:));            
+    
+    % Allocate some extra to be on the safe side; doesn't really matter
+    ivec = zeros(Nslic^2,1); jvec = zeros(Nslic^2,1); svec = zeros(Nslic^2,1);
+    
+    % Neighborhoods (ignore the edges)
+    sz = size(img);
+    
+    nedge = 1;
+    for n = 1:Nslic
+        inds = uint32(find(segments(:) == n));
+        slic_inds{n} = inds;
+        
+        neighbor_w = zeros(Nslic, 1);
+        neighbor_n = zeros(Nslic, 1);
+        
+        for ind = inds'
+            [r, c] = ind2sub(sz, ind);
+            
+            % Only consider the lower triangular
+            if r > c
+                
+                % Explore the neighborhood
+                for dr = [-1 1]
+                    for dc = [-1 1]
+                        if (r + dr > 1 && r + dr < Y) && (c + dc > 1 && c + dc < X)
+                            other_n = segments(r + dr, c + dc);                    
+
+                            if other_n > 0
+                                neighbor_w(other_n) = neighbor_w(other_n) + ...
+                                                      abs(img(r + dr, c + dc) - img(r, c));
+                                neighbor_n(other_n) = neighbor_n(other_n) + 1;
+                            end
+                        end
+                    end
+                end   
+                            
+                % Now add edges                
+                neighbor_idxs = find(neighbor_w);
+                
+                next_nedge = nedge + length(neighbor_idxs);
[…]

### 2012-12-08 src/make_mdp_prior.m

@@ -1,4 +1,4 @@
-function [ mdp ] = make_mdp_prior( X )
+function [ mdp ] = make_mdp_prior( X, misc_data )
     [N, D] = size(X);
     nX = max(X(:,1));
     nY = max(X(:,2));
@@ -6,16 +6,17 @@ function [ mdp ] = make_mdp_prior( X )
     %% Parameter priors
     prior_mean = [100 100 0.75 0.4]; % mean of the nonzero entries
     
-    prior_cov  = cov(X);
+    prior_cov  = cov(X(:,1:4));
     
     v_intensity = prior_cov(3,3);
     v_bg_radon  = prior_cov(4,4);
     
-    prior_cov(3:4,:) = zeros(2, 4);
-    prior_cov(:,3:4) = zeros(4, 2);    
+    prior_cov(3:end,:) = 0;
+    prior_cov(:,3:end) = 0;    
         
     prior_cov(3,3) = 0.5 * v_intensity;
     prior_cov(4,4) = 0.5 * v_bg_radon;
+    %prior_cov(5,5) = 1;
     
     background = GammaGamma(1, 6, 2);
     
@@ -29,16 +30,25 @@ function [ mdp ] = make_mdp_prior( X )
     cluster_assigns = initialize_one_cluster_and_background(0.99, 0.5, X);
     
     %% Construct objects
-    nw_prior = NormalWishart(prior_mean, prior_cov, prior_dof, prior_n);
-    mdp = MDP(concentration, X, cluster_assigns, nw_prior);
+        
+    feat_prior = NormalWishart(prior_mean, prior_cov, prior_dof, prior_n);
+    dist_prior = NormalWishart(2.5, 1, 1, 1);
     
+    class_prior = ProductDistribution(1, 4, feat_prior, ...
+                                      5, 5, dist_prior);
+    
+    mdp = MDP(concentration, X, misc_data, cluster_assigns, class_prior);
+    
+%    mdp = MDP(concentration, X, edge_G, cluster_assigns, feat_prior);
     
     % Hack: one uniform for (x,y) and another uniform for dimension 4 (the
     % background Radon response) ... ; NO IT'S BORKED!
     mdp.cluster_likes{1} = ProductDistribution(1, 2, Uniform(size(X, 1)), ...
                                                3, 3, background, ...
-                                               4, 4, Uniform(1)); % Because domain is [0,1]
+                                               4, 4, Uniform(1), ... % Domain [0, 1]
[…]

### 2012-12-08 src/neal3_iter.m

@@ -11,17 +11,18 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
     prior.fit([]);
     prior_pred_like = prior.pred_like(mdp.X); % col vector
     
+    x_max = max(mdp.X(:,1));
+    y_max = max(mdp.X(:,2));
+    
     ll = 0;
     
     actual_n = 0;
     
     for n = randperm(mdp.N)
-        actual_n = actual_n + 1;
-        if mod(actual_n, 1000) == 0
-            disp(['actual_n = ' num2str(actual_n)]);
-        end
+
         
-        k_old = mdp.remove_point(n);
+        k_old = mdp.remove_point(n);        
+        
         % Removed a point, so we must refit the posterior
         %
         % If our model is fully online, not a problem.
@@ -31,9 +32,10 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
                     ( mdp.N - 1 + mdp.concentration );
         x_pdf  = zeros(size(z_pdf));        
         % New cluster
-        x_pdf(end) = prior_pred_like(n);        
+        x_pdf(end) = prior_pred_like(n);                
+        
         for k = 1:mdp.n_clusters
-            % Only works for the conjugate case            
+            refit_graph_dist(mdp, x_max, y_max, k_old);
             x_pdf(k) = mdp.cluster_likes{k}.pred_like_scalar(mdp.X(n,:));
         end
         
@@ -42,7 +44,7 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
         % Unnormalized inverse cdf sampling        
         cdf = cumsum(pdf);        
         k_new = find(cdf > cdf(end)*rand(1), 1);
-        mdp.assign(n, k_new);
+        mdp.assign(n, k_new);        
         
         % DESTROY DEAD CLUSTERS!
         if mdp.cluster_counts(k_old) == 0
@@ -50,7 +52,13 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
             mdp.kill_cluster(k_old);
         end
         
-        ll = ll + log(pdf(k_new));                
+        actual_n = actual_n + 1;
+        if mod(actual_n, 1000) == 0
+            disp(['actual_n = ' num2str(actual_n)]);
+            disp(['x_pdf = ' num2str(x_pdf')]);
+        end
+        
+        ll = ll + log(pdf(k_new));
     end
 
 end

### 2012-12-08 src/neal8_iter.m

@@ -1,0 +1,30 @@
+function [ mdp, state, ll ] = neal8_iter( X, mdp, state, params )
+% Gibbs sampling iteration for Algorithm 8 in Neal 2000 [1]
+%
+% state: 
+% References:
+% [1] Radford Neal, Markov Chain Sampling Methods for Dirichlet Process
+%     Mixture Models
+    
+    unrealized_likes = cell(params.m, 1);
+    
+    ll = 0;
+    for n = 1:mdp.N
+        k_old = mdp.remove_point(n);
+        
+        if mdp.cluster_count(k_old) == 0
+            % Singleton: consider the current distribution as a new
+            % candidate
+            m = params.m - 1;
+            unrealized_likes(params.m) = mdp.cluster_likes{k_old};
+        else
+            m = params.m;
+        end
+        
+        % Draw m samples from the prior
+        
+        
+    end    
+
+end
+

### 2012-12-08 src/radonLikeFeatures/getEdgeFeatures.m

@@ -1,0 +1,23 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% getEdgeFeatures.m picks edge enhancing Gaussian Second Derivative 
+%                   Filters
+%
+% Copyright (c) Ritwik Kumar, Harvard University 2010
+%               www.seas.harvard.edu/~rkkumar
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+% KT - This is R(x,y) (equation 4) in the paper. Convolution becomes
+% multiplication in frequency domain.
+
+function [sf] = getEdgeFeatures(im, sup, sc, or)
+
+[imr imc] = size(im);
+imi = fft2(double(im));
+S=makeBarFilters(sup,or,sc);
+nof = size(S,3);
+sf = zeros(imr,imc,nof,'single');
+for i=1:nof
+    sf(:,:,i)= (circshift(ifft2(imi.*fft2(S(:,:,i),imr,imc)),[-fix(sup/2) -fix(sup/2)]));
+end
+
+

### 2012-12-08 src/radonLikeFeatures/makeBarFilters.m

@@ -1,0 +1,51 @@
+function F=makeBarFilters(sup, or, sc)
+
+% modified version of code provided at
+% http://www.robots.ox.ac.uk/~vgg/research/texclass/filters.html
+% see this website for details.
+
+  SUP=sup;                 % Support of the largest filter (must be odd)
+  SCALEX=sqrt(2).^[1:sc];  % Sigma_{x} for the oriented filters
+  NORIENT=or;              % Number of orientations
+
+  NBAR=length(SCALEX)*NORIENT;
+  NF=NBAR;
+  F=zeros(SUP,SUP,NF);
+  hsup=(SUP-1)/2;
+  [x,y]=meshgrid([-hsup:hsup],[hsup:-1:-hsup]);
+  orgpts=[x(:) y(:)]';
+
+  count=1;
+  for scale=1:length(SCALEX),
+    for orient=0:NORIENT-1,
+      angle=pi*orient/NORIENT;  % Not 2pi as filters have symmetry
+      c=cos(angle);s=sin(angle);
+      rotpts=[c -s;s c]*orgpts;
+      F(:,:,count)=makefilter(SCALEX(scale),0,2,rotpts,SUP);
+      count=count+1;
+    end;
+  end; 
+
+return
+
+function f=makefilter(scale,phasex,phasey,pts,sup)
+  gx=gauss1d(3*scale,0,pts(1,:),phasex);
+  gy=gauss1d(scale,0,pts(2,:),phasey);
+  f=normalise(reshape(gx.*gy,sup,sup));
+return
+
+function g=gauss1d(sigma,mean,x,ord)
+% Function to compute gaussian derivatives of order 0 <= ord < 3
+% evaluated at x.
+
+  x=x-mean;num=x.*x;
+  variance=sigma^2;
+  denom=2*variance;  
+  g=exp(-num/denom)/(pi*denom)^0.5;
+  switch ord,
+    case 1, g=-g.*(x/variance);
+    case 2, g=g.*((num-variance)/(variance^2));
+  end;
+return
+
+function f=normalise(f), f=f-mean(f(:)); f=f/sum(abs(f(:))); return

### 2012-12-08 src/radonLikeFeatures/myRadonLikeBackgroundSeg.m

@@ -1,0 +1,103 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% radonLikeFeaturesDemo.m Demonstrates Radon-Like Features for the 
+%             special case of edge enhancement.
+%
+% Copyright (c) Ritwik Kumar, Harvard University 2010
+%               www.seas.harvard.edu/~rkkumar
+%
+% Please cite the following paper if you fine this code useful:
+%  Ritwik Kumar, Amelio V. Reina & Hanspeter Pfister, “Radon-Like 
+%  Features and their Application to Connectomics”, IEEE Computer 
+%  Society Workshop on Mathematical Methods in Biomedical Image 
+%  Analysis (MMBIA), 2010
+%
+% Various steps involved are explained in the comments
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%clear;clc;
+% filename = 'sample.png'; % Input image
+res = 36; % # of directions of scanning
+ang = 360/res; % angular resolution of scanning
+%imo = double(imread(filename)); % read input
+
+
+imo = f2;
+
+%scaling the image to 0 - 255
+imo = imo - min(imo(:));
+imo = imo ./ max(imo(:)) * 255;
+[imro imco] = size(imo);
+
+%edge enhancing features
+sf = getEdgeFeatures(imo,15,3,12);
+imo = (max(sf,[],3));
+
+figure;
+%imshow(imo);
+
+%set the boundary to be 0
+imo(1,:) = 0;imo(imro,:)=0;imo(:,1) = 0;imo(:,imco)=0;
+
+%set the knots
+edo = edge(imo,'canny',0.37);
+
+dumo = ones(imro,imco);% helper matrix to keep track of ROI
+
+%some variable initialization
+dum=1;ed=1;int=1;
+intx = zeros(imro,imco,res,1);
+
+for xx=1:res % repeat for each scan direction
+    fprintf('Scan %d\n',xx);
+    
+    %rotate the image by the scanning angle
+    %this allows us to scan always along the horizontal axis
+    % NOT THE MOST EFFICIENT WAY OF DOING THING
+    
+    % Notation: the indices i,prev:j in the inenr loop below denotes a scan
[…]

### 2012-12-08 src/radonLikeFeatures/radonLikeFeatures.m

@@ -1,0 +1,124 @@
+function [ intx ] = radonLikeFeatures( imo, use_raw_imo, res, extract_func )
+%radonLikeFeatures Generate radon-like features from scan segments
+%
+% intx = radonLikeFeatures(imo, res, extract_fun) given [X Y] image imo,
+% returns [X Y res] feature responses in intx (one for each scan direction)
+% 
+% extract_func is a handle used in 
+%   int(i,prev:j) = extract_func(im(i,prev:j));
+%
+% Adapted from radonLikeFeatureDemo.m
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% radonLikeFeaturesDemo.m Demonstrates Radon-Like Features for the 
+%             special case of edge enhancement.
+%
+% Copyright (c) Ritwik Kumar, Harvard University 2010
+%               www.seas.harvard.edu/~rkkumar
+%
+% Please cite the following paper if you fine this code useful:
+%  Ritwik Kumar, Amelio V. Reina & Hanspeter Pfister, “Radon-Like 
+%  Features and their Application to Connectomics”, IEEE Computer 
+%  Society Workshop on Mathematical Methods in Biomedical Image 
+%  Analysis (MMBIA), 2010
+%
+% Various steps involved are explained in the comments
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%clear;clc;
+% filename = 'sample.png'; % Input image
+%res = 36; % # of directions of scanning
+    ang = 360/res; % angular resolution of scanning
+    %imo = double(imread(filename)); % read input
+
+    %scaling the image to 0 - 255
+    imo = imo - min(imo(:));
+    imo = imo ./ max(imo(:)) * 255;
+    [imro imco] = size(imo);
+
+    %edge enhancing features
+    %sf = getEdgeFeatures(imo,15,3,12);
+    %imo = (max(sf,[],3));
+    
+    % In general, use the enhanced edge features to get knots only    
+    sf = getEdgeFeatures(imo,15,3,12);
+    knot_imo = (max(sf,[],3));
+    
+    if ~use_raw_imo
+        imo = knot_imo;
+    end
+
[…]

### 2012-12-08 src/radonLikeFeatures/radonLikeFeaturesDemo.m

@@ -1,0 +1,90 @@
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% radonLikeFeaturesDemo.m Demonstrates Radon-Like Features for the 
+%             special case of edge enhancement.
+%
+% Copyright (c) Ritwik Kumar, Harvard University 2010
+%               www.seas.harvard.edu/~rkkumar
+%
+% Please cite the following paper if you fine this code useful:
+%  Ritwik Kumar, Amelio V. Reina & Hanspeter Pfister, “Radon-Like 
+%  Features and their Application to Connectomics”, IEEE Computer 
+%  Society Workshop on Mathematical Methods in Biomedical Image 
+%  Analysis (MMBIA), 2010
+%
+% Various steps involved are explained in the comments
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+clear;clc;
+filename = 'sample.png'; % Input image
+res = 36; % # of directions of scanning
+ang = 360/res; % angular resolution of scanning
+imo = double(imread(filename)); % read input
+
+%scaling the image to 0 - 255
+imo = imo - min(imo(:));
+imo = imo ./ max(imo(:)) * 255;
+[imro imco] = size(imo);
+
+%edge enhancing features
+sf = getEdgeFeatures(imo,15,3,12);
+imo = (max(sf,[],3));
+
+%set the boundary to be 0
+imo(1,:) = 0;imo(imro,:)=0;imo(:,1) = 0;imo(:,imco)=0;
+
+%set the knots
+edo = edge(imo,'canny',0.37);
+
+dumo = ones(imro,imco);% helper matrix to keep track of ROI
+
+%some variable initialization
+dum=1;ed=1;int=1;
+intx = zeros(imro,imco,res,1);
+
+for xx=1:res % repeat for each scan direction
+    fprintf('Scan %d\n',xx);
+    
+    %rotate the image by the scanning angle
+    %this allows us to scan always along the horizontal axis
+    % NOT THE MOST EFFICIENT WAY OF DOING THING
+    im = imrotate(imo,ang*(xx-1)); 
+    dum = imrotate(dumo,ang*(xx-1)); %rotate ROI tracker
+    ed = imrotate(edo,ang*(xx-1)); %rotate knot function. 
[…]

### 2012-12-08 src/radonLikeFeatures/runRadon.m

@@ -1,0 +1,33 @@
+%% Configuration
+res = 72;
+img = f1;
+
+%% Cell boundary enhancement
+
+% Do the maximum Gaussian response
+%sf = getEdgeFeatures(img,15,3,12);
+%imo = (max(sf,[],3));
+
+% argument 2 is use_raw_imo
+boundary_intx = radonLikeFeatures(img, false, res, @(scanseg) mean(scanseg));
+figure(1);
+m_boundary_intx = mean(boundary_intx,3);
+imshow(m_boundary_intx, []);
+
+%% Cell background segmentation
+
+bg_intx = radonLikeFeatures(img, true, res, @(scanseg) min(scanseg));
+figure(2);
+imshow(mean(bg_intx,3),[]);
+
+% Plotting rescale
+figure(3);
+rescale_m = mean(bg_intx,3);
+rescale_m = (rescale_m - min(rescale_m(:))) ./ (max(rescale_m(:)) - min(rescale_m(:)));
+imshow(rescale_m);
+
+%% Mitochondrial segmentation
+
+mit_intx = radonLikeFeatures(img, true, res, @(scanseg) mean(scanseg));
+figure(3);
+imshow(mean(mit_intx,3),[]);

### 2012-12-08 src/refit_graph_dist.m

@@ -1,0 +1,35 @@
+function [ mdp ] = refit_graph_dist( mdp, x_max, y_max, k )
+% dist = graph_dist(X, G, s) computes Dijkstra vector for shortest paths
+% from k's cluster mean of feature matrix X.
+    
+    if k > 1
+        s_x = round(mdp.cluster_likes{k}.pdfs{1}.pred_mean(1));
+        s_y = round(mdp.cluster_likes{k}.pdfs{1}.pred_mean(2));
+
+        % Clamp to the image boundary    
+        s_x = max(min(1, s_x), x_max);
+        s_y = max(min(1, s_y), y_max);
+
+        s = sub2ind([x_max y_max], s_x, s_y);                      
+
+        
+        % Compute superpixel distances
+        slic_dist = graphshortestpath(mdp.misc_data.edge_G, mdp.misc_data.segments(s), 'Directed', false)';
+        % Censor Inf
+        slic_dist = min(10000, slic_dist);        
+        
+        dist = zeros(mdp.N, 1);
+        
+        % Copy the slic distances to the full distances
+        for n_slic = 1:length(slic_dist)
+            dist(mdp.misc_data.slic_inds{n_slic}) = slic_dist(n_slic);
+        end
+        
+        mdp.X(:,5) = dist;
+        
+        % Explicit fit
+        kidxs = mdp.cluster_assigns == k;
+        mdp.cluster_likes{k}.pdfs{2}.fit(mdp.X(kidxs,5));
+        
+    end
+end

### 2012-12-08 src/scrape/pre_ann.m

@@ -79,6 +79,7 @@ annotation_bboxen = dataset(annoid, res, xyz_off, xyz_dim, conf, kvpairs);
 save('annotation_bboxen.mat', 'annotation_bboxen');
 
 %% Download the cutouts from the server
+load('annotation_bboxen.mat');
 vox_sz = sum(double(annotation_bboxen(:,'xyz_dim')), 2);
 small_idxs = vox_sz < 1000;
 small_bboxen = annotation_bboxen(small_idxs,:);

### 2012-12-08 src/smc_init.m

@@ -1,0 +1,126 @@
+function [ particles, weights ] = smc_init( mdp, background_like )
+% Runs one SMC pass through X to obtain sensible initialization for
+% subsequent MCMC steps.
+%
+% mdp - an uninitialized MDP object (mdp.cluster_assigns = zeros(...))
+%       Returns a fully populated mdp.cluster_assigns as well as
+%       mdp.cluster_likes.
+% X   - N x D data matrix
+%
+% References
+% [1] Paul Fearnhead, Particle filters for mixture models with an unknown
+%     number of components, Statistics and Computing 14: 11-21, 2004.
+% [2] Paul Fearnhead and Peter Clifford, On-line inference for hidden
+%     Markov models via particle filters, J. R. Statist. Soc. B (2003) pp.
+%     887-899
+% [3] Frank Wood and Michael Black, A nonparametric Bayesian alternative
+%     to spike sorting, Journal of Neuroscience Methods, 2008,
+%     doi:10.1016/j.jneumeth.2008.04.030,
+    
+    X = mdp.X;
+    [N, D] = size(X);           
+    
+    % Preallocate
+    
+    % Maintain a sub particle vector and super particle matrix with
+    % weights.
+    Nsub = 20;
+    MAX_CLUSTERS = 100;
+    
+    particles(1, Nsub) = MDP;
+    weights(1, Nsub) = 0;    
+        
+    % Intialize
+    %
+   
+    perm = randperm(N);
+    
+    % IMPORTANT! We have to start off by giving mdp a background class.    
+    mdp.cluster_likes{1} = background_like;
+    % Now MANUALLY initialize. We can't use mdp.assign, because that would
+    % create a new "normal" cluster.
+    mdp.n_clusters = 1;
+    mdp.cluster_assigns = zeros(N, 1);
+    mdp.cluster_assigns(1) = 1;
+    mdp.cluster_counts(1) = 1;
+    
+    for p = 1:Nsub
+        particles(p) = copy(mdp);
+    end   
+    weights(:)   = 1 / Nsub;    
+
+    % Precompute, as in neal3
+    prior = copy(mdp.prior);
+    prior.fit([]);
+    prior_pred_like = prior.pred_like(X); % col vector
+    
+    super_shape = [MAX_CLUSTERS + 1, Nsub];
+    tic
+    for nn = 2:N
+        
+        if mod(nn, 1000) == 0
+            ktime = toc;
+            speed = 1000 / ktime;
[…]

### 2012-12-08 src/startup.m

@@ -15,5 +15,5 @@ S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
 %f1 = S1(1000:1199,1000:1199);
 f1 = S1(1500:1699,1000:1199);
 %a1 = A1(1000:1199,1000:1199);
-PHI1 = feature_map(f1);
+[PHI, misc_data] = feature_map(f1);
 

### 2012-12-08 src/util/load_ann_slices.m

@@ -1,0 +1,7 @@
+function [ ann_slices ] = load_ann_slices( hdffile, first, last)
+% load_slices loads uint32 identifiers from openconnecto.me
+
+    ann_slices = h5read(hdffile, '/cube', [1 1 first], [Inf Inf last]);
+
+end
+

### 2012-12-08 test/test_gg.m

@@ -1,0 +1,4 @@
+gg = GammaGamma(2, 1, 1);
+
+X = gamrnd(2, 1, 100, 1);
+gg.fit(X);

### 2012-12-18 latex/frank_notes.tex
Auto-Backup Commit of All Projects

\documentclass{article}

\usepackage[square,numbers]{natbib}
\usepackage{amsmath, epsfig}
\usepackage{amsfonts}
\usepackage{subfigure}
\usepackage{graphicx}
\usepackage{amsfonts}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{easybmat}
\usepackage{footmisc}
\usepackage{url}
\usepackage{color}
\usepackage{soul}
\usepackage{multirow}

\begin{document}
\title{Notes}
\author{Frank Wood}
\date{\today}
\maketitle
\section*{October 22, 2012}

\section{Introduction}

We are interested solving the problem of segmenting image stacks.  The particular task we face is the identification and tracing of neural processes in electron micrographs whose planar resolution is around 4 nm and whose slice thickness is around 40 nm.   This task can be phrased as a multi-class classification task where each voxel is labeled with the particular neuron (or other cell type) it belongs to and the number of neurons is unknown.   This task also has been phrased as one of segmenting connected components given an affinity graph.

We approach this problem in a completely different way than that which has appeared in prior art.  Most significantly in terms of departure, we are largely uninterested in complete, space-filling segmentations and instead focus our attention on building a scaffolding that directly mathematically describes the paths taken by and connections between components in the volume.  It is our hypothesis that such a scaffold could be just as useful for further scientific study of the properties of the underlying neural tissue as a space filling segmentation.   For instance, consider a scientific question like ``what fraction of axons that pass the dendrite of this neuron synapse onto it?'' The space-filling segmentation community implicitly assumes that these kinds of questions can be answered via a statistical characterization of labelled connected components (volumes) that touch.  If this is true then a scaffolding model would be useful too because it could be used to  ``steer computational''  towards volumes that should contain features relevant to the statistical characterization of interest (here the synapsing characteristics of the axons passing one particular dendritic tree).

%The mathematical characterization of this scaffold involves asserting that each object of interest corresponds to a parameter vector that is varies over time and space. This parameter vector characterizes an object-specific distribution over feature space (of which real spatial location is a sub vector).  The feature vector computed for each voxel is to be determined.

\section{Notation}

The gray-level voxel intensity at voxel midpoint $x \in \mathbb{R}$, $y  \in \mathbb{R}$ (in the plane), and $z \in \mathbb{R}$ depth (note, in actual implementations $x,y,$ and $z$ are discrete voxel indices and the corresponding real-space locations of the voxels can be derived by assuming fixed-size voxels) is given by
$\mathbf v_i = \{x_i,y_i,z_i, c_i\} \in \mathbb{R}^3\times[0,255]$  The  feature vector computed at each voxel is $\mathbf f_i \in \mathcal{F}$ where $\mathcal{F}$ is a $D$-dimensional feature space and $i$ indexes the collection of voxels.  At the minimum the feature vector $f_i = \{ f_{i,1}, f_{i,2}, f_{i,3}, f_{i,4}, \ldots f_{i,D}\}$ will include $f_{i,1} = v_i(x)$, $f_{i,2} = v_i(y)$, and $f_{i,3} = v_i(c)$.

A single slice at depth $d$ is could be segmented using a simple mixture model.

\begin{align}
\boldsymbol\pi_d & \sim \text{Stick}(\alpha)\\
\Theta_{k} & \sim \text{H}(\Psi) \\
n_i | \boldsymbol\pi_d & \sim \boldsymbol\pi_d \\
f_{i} | \{ n_i = k \} & \sim \text{F}(f_{i} |\Theta_{k})
\end{align}

Here $\Theta_{k}$ is a class specific parameter vector describing the distribution of feature vectors (through the family $\text{F}$) associated with neuron $k$.  The probability that a particular voxel is assigned to a neuron is determined by a combination of the likelihood term and the probability of joining cluster $k$ (given by $k^{\text{th}}$ element of $\pi_{d}$,  $\pi_{d}(k).$  The variable  $n$ indicates ``neuron'' assignment.  Note that the choice of $F$ is open but dependent on the characteristics of the feature vector.  The choice of $H$ will usually be chosen for computational convenience after $F$ is selected.  If $F$ is the product of a number of exponential family distributions then $H$ can be chosen to be conjugate.

In the case of tracing skeletons of connected neural processes this is insufficient as each neurons' processes will, except in a few slices around their nuclear regions, themselves be fragmented and distributed spatially throughout the volume.  In addition the proportion of the total area of particular slices assigned to a particular neuron will also vary along the length of a neuron.

What we would like is a generative model something like the following:  neuron $k$ is, on slice $d$, further fragmented into some number of clusters whose proportions are given by $\boldsymbol\beta_{d,k}$.

\begin{align}
\boldsymbol\pi_d & \sim \text{Stick}(\alpha)\\
\boldsymbol\beta_{d,k} & \sim \text{Stick}(\delta)\\
\Theta_{d, k,\ell} & \sim \text{H}(\Psi) \\
n_i | \boldsymbol\pi_d & \sim \boldsymbol\pi_d \\
b_{i,k} | \{n_i = k \}, \boldsymbol\beta_{d,k}  &  \sim \boldsymbol\beta_{d,k} \\
f_{i} | \{ n_i = k \}, \{ b_i = \ell \} & \sim \text{F}(f_{i} |\Theta_{d, k,\ell})
\end{align}

Such a model would ordinarily be entirely unidentifiable, however, the inter-slice constraints inform the clustering at level $d$ by propagating statistical strength between layers.  In particular we would like to enforce smoothness  constraints like (until the nuclear cell region)

\[\boldsymbol\beta_{d,k} \sim \text{Coag}(\boldsymbol\beta_{d-1,k})\]

and  (after the nuclear cell region)

\[\boldsymbol\beta_{d,k} \sim \text{Frag}(\boldsymbol\beta_{d-1,k})\]

and

\[\Theta_{d, k,\ell} \sim \text{N}(\Theta_{d-1, k,\ell},\text{small})\]

when no fragmentations or coagulations take place.  When coagulations take place then

\[\Theta_{d, k,\ell} \sim \text{N}(\lambda \Theta_{d-1, k,\ell'} + (1-\lambda)\Theta_{d-1, k,\ell*},2\times\text{small})\]

and conversely

\[\Theta_{d, k,\ell'} \sim \text{N}(\Theta_{d-1, k},\frac{1}{2}\times\text{small})\]

for 2-way fragmentations

and something like $\boldsymbol\pi_d \sim \text{GPUDDP}(\boldsymbol\pi_{d-1})$

\begin{align}
\boldsymbol\pi_{} & 
\end{align}

\bibliographystyle{plain}
\bibliography{refs.bib}
\end{document}

### 2012-12-18 src/ext/anisodiffusionsimple.m

@@ -1,0 +1,69 @@
+function anisodiffusionsimple(a)
+a = im2double(a);
+g1 =a;
+g2 =a;
+stepsize1=.2;
+stepsize2=.1;
+nosteps=200;
+verbose=1;
+if verbose
+    figure(verbose);
+    subplot(2,3,1); imshow(g1); title('Original Image'); drawnow;
+end
+h=waitbar(0,'Now is doing Diffusion');
+for i=1:nosteps
+    laplaceg1=snldStep1(g1);
+    g1 = g1 +stepsize1 * laplaceg1;
+    laplaceg2=snldStep2(g2);
+    g2 = g2 +stepsize2 * laplaceg2;
+    waitbar(i/nosteps,h);
+    if verbose
+        figure(verbose);
+        subplot(2,3,2); imshow(uint8(255*(laplaceg1-min(laplaceg1(:)))/(max(laplaceg1(:))-min(laplaceg1(:)))));
+        %subplot(2,2,2); imshow(uint8(255*(laplaceg-min(laplaceg(:)))/(max(laplaceg(:))-min(laplaceg(:)))));%????
+        title('Laplace term')
+        subplot(2,3,3); imshow(g1);
+        title('Anisotropic Diffusion'); drawnow;
+        subplot(2,3,5); imshow(uint8(255*(laplaceg2-min(laplaceg2(:)))/(max(laplaceg2(:))-min(laplaceg2(:)))));
+        %subplot(2,2,2); imshow(uint8(255*(laplaceg-min(laplaceg(:)))/(max(laplaceg(:))-min(laplaceg(:)))));%????
+        title('Laplace term')
+        subplot(2,3,6); imshow(g2);
+        title('Isotropic Diffusion'); drawnow;
+    end
+        waitbar(i/nosteps,h);
+end
+close(h);
+
+function r = snldStep1( L )
+% Discrete numerical scheme of dL/dt for scalar diffusion
+N = size(L, 1);
+M = size(L, 2);
+lamda=.008;
+% Set delta_x,delta_y=2,so we can set delta_x/2=delta_y/2=1
+Lpc = translateImage( L, 1, 0 );
+Lmc = translateImage( L, -1, 0 );
+Lcp = translateImage( L, 0, 1 );
+Lcm = translateImage( L, 0, -1 );
[…]

### 2012-12-18 src/ext/translateImage.m

@@ -1,0 +1,20 @@
+    function r = translateImage( f, di, dj )
+    % Translation of an image
+    N = size( f, 1 );
+    M = size( f, 2 );
+    if di>0
+        iind = [(di+1):N, N*ones(1,di)];
+    elseif di<0
+        iind = [ ones(1,-di), 1:(N+di) ];
+    else
+        iind = 1:N;
+    end
+    if dj>0
+        jind = [(dj+1):M, M*ones(1,dj)];
+    elseif dj<0
+        jind = [ ones(1,-dj), 1:(M+dj) ];
+    else
+        jind = 1:M;
+    end
+    r = f( iind, jind );
+ ...

### 2013-01-30 src/GammaGamma.m
fixed graph feature; precompute distances

@@ -43,12 +43,14 @@ classdef GammaGamma < OnlineDistribution
             o.data_n = length(X);
             o.post_shape = o.prior_shape + o.data_n * o.shape;
             o.post_rate  = o.prior_rate  + sum(X);
+            assert(o.post_rate > 0);
         end
         
         function add_point(o, x)
             o.data_n = o.data_n + 1;
             o.post_shape = o.post_shape + o.shape;
             o.post_rate  = o.post_rate  + x;
+            assert(o.post_rate > 0);
         end
         
         function remove_point(o, x)
@@ -59,26 +61,39 @@ classdef GammaGamma < OnlineDistribution
                 o.data_n = o.data_n - 1;
                 o.post_shape = o.post_shape - o.shape;
                 o.post_rate  = o.post_rate - x;
+                assert(o.post_rate > 0);
             end            
+        end
+        
+
+        
+        % Double check the paper with this
+        function xs = sample_prior(o, n)
+            rate = gamrnd(o.prior_shape, 1 / o.prior_rate);
+            xs = gamrnd(o.prior_shape, 1 / rate, n, 1);
+        end
+        
+        function p = pred_like_scalar(o, X)            
+            % Dubey (1.1)
+            if X == 0 % avoid log(0)
+                log_top_part = o.post_shape * log(o.post_rate);
+                log_bot_part = betaln(o.shape, o.post_shape) + (o.shape + o.post_shape) * log(o.post_rate + X);
+                p = exp(log_top_part - log_bot_part);
+                % Important
+                p = p * 0^(o.shape - 1);
+            else
+                log_top = o.post_shape * log(o.post_rate) + (o.shape - 1) * log(X);
+                log_bot = betaln(o.shape, o.post_shape) + (o.shape + o.post_shape) * log(o.post_rate + X);
+            
+                p = exp(log_top - log_bot);                  
+            end
         end
         
         function p = pred_like(o, X)
         % Predictive likelihood (compound gamma). X can be a vector.
             % Dubey (1.1)
-            log_top = o.post_shape * log(o.post_rate) + (o.shape - 1) * log(X);
-            log_bot = betaln(o.shape, o.post_shape) + (o.shape + o.post_shape) * log(o.post_rate + X);
-            
-            p = exp(log_top - log_bot);
+            p = o.pred_like_scalar(X);            
         end
-        
-        function p = pred_like_scalar(o, X)
-            % Dubey (1.1)
[…]

### 2013-01-30 src/GraphDistDistribution.m

@@ -1,0 +1,123 @@
+classdef GraphDistDistribution < matlab.mixin.Copyable & OnlineDistribution
+% Distribution of the geodesic (graph) distance of a given point to the
+% cluster center 
+% Applies independent probability models to distinct dimensions of X    
+    
+    properties
+        underlying, mdp, misc_data;
+        my_k;
+        last_fitted_kss;
+        
+        slic_dim = 5;
+        % Hack; to softcode later
+        x_max = 200;
+        y_max = 200;
+    end
+    
+    properties (Dependent)
+        % The slic segment OF THE CLUSTER CENTER.
+        k_slic_segment;        
+    end
+    
+    methods(Access = protected)                
+        % Override copyElement method:
+        function cpObj = copyElement(obj)            
+            cpObj = copyElement@matlab.mixin.Copyable(obj);
+            cpObj.mdp = obj.mdp;
+            cpObj.misc_data = obj.misc_data;
+            cpObj.last_fitted_kss = obj.last_fitted_kss;
+            
+%            assert(~isempty(obj.my_k));
+            cpObj.my_k = obj.my_k;
+            
+            % Make a deep copy of underlying
+            cpObj.underlying = copy(obj.underlying);
+        end
+    end
+    
+    methods
+        function kss = get.k_slic_segment(o)
+            assert(o.my_k > 1);
+            
+            s_x = round(o.mdp.cluster_likes{o.my_k}.pdfs{1}.post_mean(1));
+            s_y = round(o.mdp.cluster_likes{o.my_k}.pdfs{1}.post_mean(2));
+
+            % Clamp to the image boundary
+            if s_x < 1 || s_x > o.x_max || s_y < 1 || s_y > o.y_max
+                warning(['crossed image boundary: s_x = ' num2str(s_x) ...
+                         ' s_y = ' num2str(s_y)]);
+            end
+
+            s_x = min(max(1, s_x), o.x_max);
+            s_y = min(max(1, s_y), o.y_max);                
+
+            kss = o.misc_data.segments(s_x,s_y);            
+        end
+        
+        function d = slic_to_dist(o, x)                        
+            if isempty(o.last_fitted_kss) || o.last_fitted_kss ~= o.k_slic_segment
+                %warning('refit triggered');
+                kidxs  = o.mdp.cluster_assigns == o.my_k; 
+                kslics = o.mdp.X(kidxs,o.slic_dim);
+                o.fit(kslics);
+            end
[…]

### 2013-01-30 src/MDP.m

@@ -65,15 +65,8 @@ classdef MDP < matlab.mixin.Copyable
                     o.cluster_counts = accumarray(cluster_assigns(nzidxs), ones(o.N, 1));            
 
                     for k = 1:o.n_clusters
-                        idxs = cluster_assigns == k;
-                                                
-                        o.cluster_likes{k} = copy(prior);
-                        
-                        % TEST -- did we actually copy?
-                        prior.pdfs{2}.data_n = 999;
-                        assert(o.cluster_likes{k}.pdfs{2}.data_n ~= 999);
-                        prior.pdfs{2}.data_n = 0;
-                        
+                        idxs = cluster_assigns == k;                                                
+                        o.cluster_likes{k} = copy(prior);                                                
                         o.cluster_likes{k}.fit(X(idxs,:));                        
                     end
                 end
@@ -112,6 +105,14 @@ classdef MDP < matlab.mixin.Copyable
                 % New
                 dist = copy(o.prior);                
                 o.assign_to_new(n, dist);
+                
+                % HACK: Tell
+                % We can't use isfield, which works only for struct.
+                % (MATLAB sucks)
+                if find(ismember(fieldnames(dist.pdfs{2}), 'my_k')==1) > 0
+                    dist.pdfs{2}.my_k = k;
+                end                
+                
                 dist.fit(o.X(n,:));
             else
                 % Existing
@@ -164,6 +165,15 @@ classdef MDP < matlab.mixin.Copyable
             gtidxs = o.cluster_assigns > k;
             o.cluster_assigns(gtidxs) = o.cluster_assigns(gtidxs) - 1;
             o.n_clusters = o.n_clusters - 1;
+            
+            % Update the my_k's
+            for kk = k:o.n_clusters
+                if kk > 1
+                    assert(o.cluster_likes{kk}.pdfs{2}.my_k - 1 == kk, ...
+                           'GraphDist my_k precondition failed');
+                    o.cluster_likes{kk}.pdfs{2}.my_k = kk;
+                end
+            end
             
             assert(length(o.cluster_counts) == o.n_clusters && ...
                    length(o.cluster_likes) == o.n_clusters, 'bad removal in kill_cluster');

### 2013-01-30 src/experiment/test_mdp.m

@@ -1,10 +1,9 @@
 %% Prior!
 clear mdp
-clear misc_data
-misc_data = struct();
+
 [misc_data.segments, misc_data.slic_inds, misc_data.edge_G] = ...
     make_diff_graph(f1, 10, 0.1);
-misc_data
+
 mdp = make_mdp_prior(PHI, misc_data);
 
 %% Run!

### 2013-01-30 src/feature_map.m

@@ -3,22 +3,23 @@ function [ PHI, misc_data ] = feature_map( img )
     N = X * Y;
     
     % Number of features
-    D = 4;
-    
-    res = 72;
+    D = 5;   
+    randon_angles = 72;
+    canny_thresh = 0.25;
     
     PHI = zeros(N, D);
     
-    bg_intx = radonLikeFeatures(img, true, res, @(scanseg) min(scanseg));
+    bg_intx = radonLikeFeatures(img, true, randon_angles, canny_thresh, @(scanseg) min(scanseg));
     m_bg_intx = mean(bg_intx,3);        
     
     % Scale to 0-1
+    % NO, DON'T DO THIS! 
     range = max(m_bg_intx(:)) - min(m_bg_intx(:));
     m_bg_intx = (m_bg_intx - min(m_bg_intx(:))) ./ range;
     %figure; imshow(m_bg_intx);
         
-    % argument 2 is use_raw_imo
-    bd_intx = radonLikeFeatures(img, false, res, @(scanseg) mean(scanseg));
+    % Argument 2 is use_raw_imo. This turns on Eq (4) in Kumar.
+    bd_intx = radonLikeFeatures(img, false, randon_angles, canny_thresh, @(scanseg) mean(scanseg));
     m_bd_intx = mean(bd_intx,3);
     
     range = max(m_bd_intx(:)) - min(m_bd_intx(:));
@@ -26,21 +27,27 @@ function [ PHI, misc_data ] = feature_map( img )
     %figure; imshow(m_bd_intx);
     
     misc_data = struct();    
+    misc_data.m_bg_intx = m_bg_intx;
+    misc_data.m_bd_intx = m_bd_intx;
+    
+    misc_data.bd_intx = bd_intx;
+    
+    misc_data.img = img;
+    
     [misc_data.segments, misc_data.slic_inds, misc_data.edge_G] = make_diff_graph(m_bd_intx, 10, 0.1);    
+    
+    misc_data.slic_dist = graphallshortestpaths(misc_data.edge_G, 'Directed', false);    
     
     n = 1;
     for y = 1:Y
-        for x = 1:X
-            PHI(n, :) = [x y img(x,y) m_bg_intx(x,y)];
+        for x = 1:X            
+            PHI(n,:) = [x y img(x,y) m_bg_intx(x,y) double(misc_data.segments(x,y))];
             n = n+1;
         end
     end
     
[…]

### 2013-01-30 src/make_diff_graph.m

@@ -12,7 +12,14 @@ function [ segments, slic_inds, edge_G ] = make_diff_graph( img, slic_npixels, s
     [X, Y] = size(img);
     
     % vl_slic returns 0-indexed; we want 1-indexed.
-    segments = vl_slic(single(img), slic_npixels, slic_reg) + 1;        
+    segments = vl_slic(single(img), slic_npixels, slic_reg) + 1;
+    
+%     figure(15);
+%     imagesc(segments);
+    
+    mean_img = showslic(segments, img);
+%     figure(16);
+%     imshow(mean_img);
     
     assert(min(segments(:)) > 0);
     Nslic = max(segments(:));            
@@ -28,43 +35,47 @@ function [ segments, slic_inds, edge_G ] = make_diff_graph( img, slic_npixels, s
         inds = uint32(find(segments(:) == n));
         slic_inds{n} = inds;
         
-        neighbor_w = zeros(Nslic, 1);
-        neighbor_n = zeros(Nslic, 1);
+        neighbor_w = zeros(Nslic, 1);        
         
         for ind = inds'
-            [r, c] = ind2sub(sz, ind);
-            
-            % Only consider the lower triangular
-            if r > c
+            [r, c] = ind2sub(sz, ind);           
                 
-                % Explore the neighborhood
-                for dr = [-1 1]
-                    for dc = [-1 1]
-                        if (r + dr > 1 && r + dr < Y) && (c + dc > 1 && c + dc < X)
-                            other_n = segments(r + dr, c + dc);                    
+            % Explore the neighborhood
+            for dr = [-1 1]
+                for dc = [-1 1]
+                    if (r + dr > 1 && r + dr < Y) && (c + dc > 1 && c + dc < X)
+                        other_n = segments(r + dr, c + dc);                    
 
-                            if other_n > 0
-                                neighbor_w(other_n) = neighbor_w(other_n) + ...
-                                                      abs(img(r + dr, c + dc) - img(r, c));
-                                neighbor_n(other_n) = neighbor_n(other_n) + 1;
-                            end
+                        % CHECK THIS LINE... earlier you asserted other_n >
+                        % 0; should really check if its a different
+                        % segment.
+                        %
+                        % Should really take difference between superpixel
+                        % means.
+                        if other_n ~= n
+                            neighbor_w(other_n) = abs(mean_img(r + dr, c + dc) - mean_img(r, c));
+                                                        
[…]

### 2013-01-30 src/make_mdp_prior.m

@@ -4,7 +4,7 @@ function [ mdp ] = make_mdp_prior( X, misc_data )
     nY = max(X(:,2));
     
     %% Parameter priors
-    prior_mean = [100 100 0.75 0.4]; % mean of the nonzero entries
+    prior_mean = [100 100 0.6 0.3]; % mean of the nonzero entries
     
     prior_cov  = cov(X(:,1:4));
     
@@ -31,30 +31,41 @@ function [ mdp ] = make_mdp_prior( X, misc_data )
     %% Construct objects
         
     feature_prior  = NormalWishart(prior_mean, prior_cov, prior_dof, prior_n);
-    distance_prior = GammaGamma(1, 10, 1000);    
+    distance_underlying = GammaGamma(1, 5, 4);
     
+    
+    % HACK; to patch up later.
     class_prior = ProductDistribution(1, 4, feature_prior, ...
-                                      5, 5, distance_prior);
+                                      5, 5, distance_underlying);
     
     % Hack: low probabilities when plugged into NormalGamma.
     % Interpretation: when starting a new cluster, take its Dijkstra
     % distance to be "far".
+                                          
+    mdp = MDP(concentration, X, misc_data, cluster_assigns, class_prior);    
     
-    % Near = 1e-10, Far = 500, Intermediate = between
-    X(:,5) = 10;
-                                  
-    mdp = MDP(concentration, X, misc_data, cluster_assigns, class_prior);
+    distance_prior = GraphDistDistribution(distance_underlying);
+    distance_post  = GraphDistDistribution(mdp.cluster_likes{2}.pdfs{2});
+        
+    distance_prior.mdp = mdp;
+    distance_post.mdp = mdp;
+    distance_prior.misc_data = misc_data;        
+    distance_post.misc_data = misc_data;
     
-%    mdp = MDP(concentration, X, edge_G, cluster_assigns, feat_prior);
+    distance_post.my_k = 2;
+    
+    % Hack: plug back in the distance underlying. mdp.prior is a
+    % ProductDistribution
+    mdp.prior.pdfs{2} = distance_prior;
+    mdp.cluster_likes{2}.pdfs{2} = distance_post;        
     
     % Hack: one uniform for (x,y) and another uniform for dimension 4 (the
     % background Radon response)
     mdp.cluster_likes{1} = ProductDistribution(1, 2, Uniform(size(X, 1)), ...
                                                3, 3, background, ...
                                                4, 4, Uniform(1), ... % Domain [0, 1]
[…]

### 2013-01-30 src/neal3_iter.m

@@ -8,8 +8,12 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
 
 % Precompute
     prior = copy(mdp.prior);
-    prior.fit([]);
-    prior_pred_like = prior.pred_like(mdp.X); % col vector
+    prior.fit([]);        
+    
+    % Hack: We haven't assigned my_k, so we can't use this
+    prior_pred_like = prior.pdfs{1}.pred_like(mdp.X(:,1:4));
+    % A new class has distance 0 from its cluster center.
+    prior_pred_like = prior_pred_like .* prior.pdfs{2}.underlying.pred_like(0);
     
     x_max = max(mdp.X(:,1));
     y_max = max(mdp.X(:,2));
@@ -18,43 +22,35 @@ function [ mdp, state, ll ] = neal3_iter( mdp, state, params )
     
     actual_n = 0;
     
-    for n = randperm(mdp.N)
-        assert(length(mdp.misc_data.slic_inds{end}) > 1);
-
-        
-        k_old = mdp.remove_point(n);        
-        
-        % Removed a point, so we must refit the posterior
-        %
-        % If our model is fully online, not a problem.
-        %mdp.refit(k_old);
+    for n = randperm(mdp.N)        
+        k_old = mdp.remove_point(n);                        
         
         z_pdf = [ mdp.cluster_counts ; mdp.concentration ] ./ ...
                     ( mdp.N - 1 + mdp.concentration );
-        x_pdf  = zeros(size(z_pdf));        
+        x_pdf  = zeros(size(z_pdf));
+        
+        for k = 1:mdp.n_clusters
+            x_pdf(k) = mdp.cluster_likes{k}.pred_like_scalar(mdp.X(n,:));
+        end
+        
         % New cluster
         x_pdf(end) = prior_pred_like(n);                
         
         % HACK
-        for k = 1:mdp.n_clusters
-            X5_before = mdp.X(:,5);
-            mdp = refit_graph_dist(mdp, x_max, y_max, k);
-            assert(length(mdp.misc_data.slic_inds{end}) > 1);
-            x_pdf(k) = mdp.cluster_likes{k}.pred_like_scalar(mdp.X(n,:));
-            
-            if k > 1 && mdp.misc_data.segments(k) ~= mdp.misc_data.segments(k_old)
-                % Unless k and k_old belong to the same superpixel, make
-                % sure that the graph distance actually changed.            
-                if all(mdp.X(:,5) == X5_before)
[…]

### 2013-01-30 src/old/showslic.m

@@ -1,33 +1,0 @@
-function [ J ] = showslic( segments, I, mapzeroto )
-    
-    J = zeros(size(I), 'single');
-    segments = segments + 1;
-    N = max(segments(:));
-    
-    S = zeros(N, 1);
-    n = zeros(N, 1);
-    
-    [X Y] = size(I);
-    for j = 1:Y
-        for i = 1:X
-            seg = segments(i, j);
-            S(seg) = S(seg) + I(i, j);
-            n(seg) = n(seg) + 1;
-        end
-    end
-    
-    S = S ./ n;
-    
-    % For watershed instead of slic
-    if nargin > 2
-        S(1) = mapzeroto;
-    end
-    
-    for j = 1:Y
-        for i = 1:X                            
-            J(i, j) = S(segments(i, j));
-        end
-    end    
-
-end
-

### 2013-01-30 src/plot/plot_graph_edge_dist.m

@@ -5,17 +5,22 @@ function [ h ] = plot_graph_edge_dist( h, k, mdp, x_max, y_max )
     % This will show results from the most recent k
     figure(h);
         
-    subplot(1, mdp.n_clusters, k);
-
+    subplot(2, mdp.n_clusters, k);
     dist = mdp.X(:,5);
     dist_img = reshape(dist, x_max, y_max);
 
     % Transform the column major to row major
-    dist_img = dist_img';
+    %
+    % That's a lie... 
+    %
+    %dist_img = dist_img';
     imshow(dist_img, [min(dist) max(dist)]);
 
     hold on;            
     scatter(mdp.cluster_likes{k}.pdfs{1}.post_mean(1), ...
             mdp.cluster_likes{k}.pdfs{1}.post_mean(2), 'filled');
     hold off;                   
+    
+    subplot(2, mdp.n_clusters, mdp.n_clusters + k);
+    hist(dist, 100);    
 end

### 2013-01-30 src/plot/plot_intensity_hists.m

@@ -31,8 +31,9 @@ function [ h ] = plot_intensity_hists( h, mdp, X )
         subplot(3, K, 2*K + k);
         
         if k ~= 1
-            xxs = linspace(eps, 500, 100);
-            plot(xxs, mdp.cluster_likes{k}.pdfs{2}.pred_like_scalar(xxs));
+            %xxs = linspace(eps, 500, 100);
+            %plot(xxs, mdp.cluster_likes{k}.pdfs{2}.underlying.pred_like_scalar(xxs));
+            hist(mdp.cluster_likes{k}.pdfs{2}.empirical_distribution());
         end            
         
     end    

### 2013-01-30 src/radonLikeFeatures/radonLikeFeatures.m

@@ -1,4 +1,4 @@
-function [ intx ] = radonLikeFeatures( imo, use_raw_imo, res, extract_func )
+function [ intx ] = radonLikeFeatures( imo, use_raw_imo, res, canny_thresh, extract_func )
 %radonLikeFeatures Generate radon-like features from scan segments
 %
 % intx = radonLikeFeatures(imo, res, extract_fun) given [X Y] image imo,
@@ -48,11 +48,11 @@ function [ intx ] = radonLikeFeatures( imo, use_raw_imo, res, extract_func )
     end
 
     %set the boundary to be 0 -- NO; SET IT TO BE WHITE!
-    imo(1,:) = 0;imo(imro,:)=0;imo(:,1) = 0;imo(:,imco)=0;
-    knot_imo(1,:) = 0;knot_imo(imro,:)=0;knot_imo(:,1) = 0;knot_imo(:,imco)=0;
+    imo(1,:) = 1;imo(imro,:)=1;imo(:,1) = 1;imo(:,imco)=1;
+    knot_imo(1,:) = 1;knot_imo(imro,:)=1;knot_imo(:,1) = 1;knot_imo(:,imco)=1;
 
     %set the knots
-    edo = edge(knot_imo,'canny',0.37);    
+    edo = edge(knot_imo,'canny',canny_thresh);    
     
     dumo = ones(imro,imco);% helper matrix to keep track of ROI
 

### 2013-01-30 src/refit_graph_dist.m

@@ -7,32 +7,30 @@ function [ mdp ] = refit_graph_dist( mdp, x_max, y_max, k )
         s_x = round(mdp.cluster_likes{k}.pdfs{1}.post_mean(1));
         s_y = round(mdp.cluster_likes{k}.pdfs{1}.post_mean(2));
 
-        % Clamp to the image boundary    
+        % Clamp to the image boundary
+        
+        if s_x < 1 || s_x > x_max || s_y < 1 || s_y > y_max
+            warning(['crossed image boundary: s_x = ' num2str(s_x) ...
+                     ' s_y = ' num2str(s_y)]);
+        end
+        
         s_x = min(max(1, s_x), x_max);
-        s_y = min(max(1, s_y), y_max);
+        s_y = min(max(1, s_y), y_max);                
 
         s = sub2ind([x_max y_max], s_x, s_y);
         
         % Compute superpixel distances
         my_slic = mdp.misc_data.segments(s);
-        slic_dist = graphshortestpath(mdp.misc_data.edge_G, my_slic, 'Directed', false)';
+        [slic_dist, slic_path, ~] = graphshortestpath(mdp.misc_data.edge_G, my_slic, 'Directed', false);
+        assert(all(~isinf(slic_dist(:))));
         
-        % Soften Inf: If pixel s cannot reach pixel i, slic_dist(i) will be
-        % Inf. This will mess up probabilities, but we want at least some
-        % probability of assigning s and i to the same cluster to mix.
+        dist = zeros(mdp.N, 1);                
         
-        % This is twice the prior mean, an arbitrary choice as good as any.        
-        cap = 500;
-        slic_dist = min(cap, slic_dist);        
-        
-        dist = zeros(mdp.N, 1);
-        
-        %assert(length(mdp.misc_data.slic_inds{end}) > 1);
-        
-        % Copy the slic distances to the full distances
-        for n_slic = 1:length(slic_dist)     
-            %assert(length(mdp.misc_data.slic_inds{end}) > 1);            
-            dist(mdp.misc_data.slic_inds{n_slic}) = slic_dist(n_slic);        
+        % Copy the normalized slic distances to the full distances
+        for n_slic = 1:length(slic_dist)    
+            %dist(mdp.misc_data.slic_inds{n_slic}) = slic_dist(n_slic);
+
+            dist(mdp.misc_data.slic_inds{n_slic}) = slic_dist(n_slic) / length(slic_path{n_slic});
         end
         
[…]

### 2013-01-30 src/run_mdp.m

@@ -11,9 +11,9 @@ function [ samples ] = run_mdp( iter_f, state, params, X, img, mdp, niters, burn
     set(hist_h, 'Units', 'normalized', 'position', [0 0.5 1 0.4]);
     clf;  
     
-    dist_h = figure(3);
-    set(dist_h, 'Units', 'normalized', 'position', [0 0.7 1 0.4]);
-    clf;
+%     dist_h = figure(30);
+%     set(dist_h, 'Units', 'normalized', 'position', [0 0.7 1 0.4]);
+%     clf;
         
     for n = 1:niters
         tic                
@@ -27,7 +27,9 @@ function [ samples ] = run_mdp( iter_f, state, params, X, img, mdp, niters, burn
         rate     = mdp.N / itertime;
         
         fprintf(1, 'Iter %d: time = %0.2f; speed = %.2f; K = %d, change = %0.3f; loglike = %06.2f\n', ...
-                n, itertime, rate, mdp.n_clusters, change, ll);        
+                n, itertime, rate, mdp.n_clusters, change, ll);       
+        disp(mdp.cluster_counts)
+        
 
         % Save the log-likelihood
         loglike(n) = ll;                

### 2013-01-30 src/startup.m

@@ -13,7 +13,8 @@ local_startup;
 S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
 %A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
 %f1 = S1(1000:1199,1000:1199);
-f1 = S1(1500:1699,1000:1199);
+%f1 = S1(1200:1399,900:1099);
+f1 = S1(1600:1799,1:200);
 %a1 = A1(1000:1199,1000:1199);
 [PHI, misc_data] = feature_map(f1);
 

### 2013-01-30 src/GammaGamma.m
GammaGamma sampling done

@@ -70,7 +70,7 @@ classdef GammaGamma < OnlineDistribution
         % Double check the paper with this
         function xs = sample_prior(o, n)
             rate = gamrnd(o.prior_shape, 1 / o.prior_rate);
-            xs = gamrnd(o.prior_shape, 1 / rate, n, 1);
+            xs = gamrnd(o.shape, 1 / rate, n, 1);
         end
         
         function p = pred_like_scalar(o, X)            

### 2013-01-30 src/make_mdp_prior.m

@@ -4,7 +4,7 @@ function [ mdp ] = make_mdp_prior( X, misc_data )
     nY = max(X(:,2));
     
     %% Parameter priors
-    prior_mean = [100 100 0.6 0.3]; % mean of the nonzero entries
+    prior_mean = [100 100 0.9 0.4]; % mean of the nonzero entries
     
     prior_cov  = cov(X(:,1:4));
     
@@ -26,7 +26,7 @@ function [ mdp ] = make_mdp_prior( X, misc_data )
     
     %% Initialize cluster assignments
     % Wood: assign them uniformly at random
-    cluster_assigns = initialize_one_cluster_and_background(0.99, 0.5, X);
+    cluster_assigns = initialize_one_cluster_and_background(0.99, 0.7, X);
     
     %% Construct objects
         

### 2013-01-30 src/make_mdp_prior.m
fixed make_mdp_prior for brighter picture

@@ -17,7 +17,7 @@ function [ mdp ] = make_mdp_prior( X, misc_data )
     prior_cov(3,3) = 0.5 * v_intensity;
     prior_cov(4,4) = 0.5 * v_bg_radon;    
     
-    background = GammaGamma(1, 6, 2);
+    background = GammaGamma(1, 10, 5);
     
     prior_dof = 5;
     prior_n   = 0.5;
@@ -31,7 +31,7 @@ function [ mdp ] = make_mdp_prior( X, misc_data )
     %% Construct objects
         
     feature_prior  = NormalWishart(prior_mean, prior_cov, prior_dof, prior_n);
-    distance_underlying = GammaGamma(1, 5, 4);
+    distance_underlying = GammaGamma(1, 20, 10);
     
     
     % HACK; to patch up later.

### 2013-01-31 latex/submission.tex
model in paper is no longer a lie

@@
-\documentclass[draft]{article}
+\documentclass[draft,english]{article}
 \usepackage{fullpage}
+\usepackage{amsmath}
 
 \usepackage[draft]{fixme}
+\usepackage{babel}
+
+\newcommand{\+}[1]{\ensuremath{\boldsymbol{\mathrm{#1}}}}
 
@@
 \title{A Bayesian Nonparametric Approach to Neural Image Segmentation}
-
 \author{Kui Tang and Frank Wood}
@@
 \begin{abstract}
-We \fxfatal*{we don't really develop; what the best verb?}{develop} a Bayesian nonparametric method to segment neurons in electron micrograph images of brain tissue.
+We develop a Bayesian nonparametric method to segment neurons in electron micrograph images of brain tissue.
 Our model imposes spatial locality on arbitrary features, automatically determines the number of cells in an image, and can be used in both unsupervised and semi-supervised settings.
@@
 
-\fxfatal{Obtain more information about this dataset, particularly stain protocol, and put in the right place.}
+\fxfatal{What microscopy method did they use? and put in the right place.}
 We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \mbox{ nm}^3$ collected by \fxfatal{Unpublished; to cite}{Kasthuri et al.}
 
-\fxfatal{Frank's notes: Think of your initial reaction of "wth are all these features?" What?! Spatial coordinates for GMM?}
+An image is composed of pixels, which immediately offer three features: $x$- and $y$- coordinates, and intensity (normalized to $[0,1]$).
+One way to segment cells is to model pixels with a Gaussian mixture model (GMM).
+Each cluster represents one cell, and each pixel's mixture weights represent the probability that of membership in a given cell.
+This approach would work well if (1) the number of cells in an image are known \emph{a priori} (2) pixels intensities are homogeneous within cells; and (3) pixels belonging to a cell occupy a continuous area.
 
-We model each pixel as feature vector $\mathbf{x}_n = (x_n, y_n, \phi_{1n}, \ldots, \phi_{mn})$.
-The first two dimensions encode the $x$ and $y$ coordinates of the pixel in order to provide spatial smoothing, while $\phi_{\cdot n}$ represent feature responses.
-Our model can handle arbitrary features, but in this paper, we will put pixel intensity in $\phi_{1\cdot} \in [0,1]$.
+We present our model in section~ref{sec:model}.
+The rest of the subsections explain how we overcome these difficulties.
+First, the number of cells must be inferred from the image itself.
+To do so, we employ a Mixture Dirichlet process, discussed in section~\ref{sec:mdp}.
+Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3.
+We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}, and constrain cluster assignments with a geodesic distance that penalizes boundary crossings in section~\ref{sec:geodesic}.
+
+\subsection{Model}
+\ref{sec:model}
+In summary, pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances  to every other pixel.
+The first four components, $\+{x}_{n,1:4}$ obey a multivariate normal distribution while the  Putting these components together yields the model in Figure~\ref{fig:model}.
+The generative story is:
+\begin{itemize}
+\item $\+\pi \sim \mbox{DirichletProcess}(\alpha)$
+\item $k | \+\pi \sim \mbox{Discrete}(\+\pi)$
+\item If $k$ is a new (empty) cluster:
+  \begin{itemize}
+  \item $\Lambda_k | T, \nu \sim \mbox{Wishart}(\Lambda | T, \nu)$ (in the notation of \cite{Murphy2003})
+  \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
+  \item $\beta_k \sim \mbox{Gamma}(\beta_k | \alpha_0, \beta_0)$
+  \end{itemize}
+\item $\+x | \+\mu_k , \Lambda_k, \beta_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \mbox{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \mbox{rate} = \beta_k \right)$ (in the notation of section~\ref{sec:geodesic})
+\end{itemize}
+
+\subsection{Mixture Dirichlet Processes}
+\label{sec:mdp}
+We begin with a finite mixture model shown in Figure~\ref{fig:finitemix} and derive the mixture Dirichlet process as a limiting case. This follows the derivation first given by \cite{Rasmussen2000}.
+
+The $N$ observations are divided among $K$ classes.
+We record the class membership of $\+{x}_n$ with a latent indicator $z_n \in \{1, \ldots, K \}$.
+Given a class assignment $z_n = k$, we assume $x_n$ is drawn from a distribution $F(\+{x}|\Theta_k)$ where $\Theta_k$ is a vector of parameters.
+Therefore, a mixture model gives each class its own parameter vector, but assumes that observations are conditionally independent given a class.
+
+The indicators $z_n$ are drawn from the discrete distribution $\+\pi$, a $K$-dimensional vector which is in turn is drawn from a symmetric Dirichlet distribution parameterized by $\alpha$: $$p \left( \+\pi | \alpha, K \right) = \frac{ \Gamma ( \alpha ) }{ \Gamma ( \alpha/K ) ^K} \prod_{k=1}^K \+\pi_k^{\alpha/K - 1}.$$ Because the Dirichlet and discrete distributions are conjugate, we can marginalize $\+\pi$.
+Let $\+{z}_{-n}$ denote the vector of all indicators excluding element $n$ and $N_{k,-n}$ denote the number of points assigned to class $k$ excluding point $n$.
+A standard result \fxfatal{CITATION} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }$$ which makes Gibbs sampling easy.
+
+Now consider $K \rightarrow \infty$.
+First, the term $\alpha / K \rightarrow 0$.
+Given finitely many observations, we have only finitely many components $k$ with $N_{k,-n} > 0$.
+Therefore, the density
+\[
+p(z_{n}=k|\+{z}_{-n},\alpha)=\begin{cases}
+\frac{N_{k,-n}}{N-1+\alpha} & N_{k,-n}>0\\
+\frac{\alpha}{N-1+\alpha} & \mbox{otherwise}
+\end{cases}
+\]
+satisfies the limiting conditions and correctly normalizes to 1 when summed with $k = 1, \ldots, \infty$.
+We now say that the mixing distribution $\+\pi$, which is now infinite-dimensional, is drawn from a \emph{Dirichlet process}.
+Since conjugacy allows us to eliminate $\+\pi$ from computations, Gibbs sampling from an infinite mixture is just as efficient as for a finite mixture case.
+
+\subsection{Radon-like Features}
+\label{sec:radon}
+Ideally, for a mixture model, pixel intensities within a cell should be homogeneous.
+But the diversity of intracellular structures and significant imaging noise yield data far from this ideal.
+As a countermeasure, we augment raw pixel intensities with $\emph{Radon-like feature}$ responses, which have proven useful for connectomics segmentation \cite{Kumar2010}.
+\fxfatal{Make this clearer...
+Even the original paper doesn't have a concise description}.
+A Radon-like feature aggregates a function of image intensity over regions defined by structural constraints, such as edges.
+Each feature is measured from a single direction, which we can represent as a \emph{scan line} $\+{l}(t) : [0, 1] \rightarrow \+{R}^2$.
+We represent the structure information as a set of points $\{ t_1, \ldots, t_m \}$ along line $\+{l}$.
+We follow the practice in \cite{Kumar2010} to evaluate Radon-like features along uniform angular intervals (we find 72 to avoid most discretization artifacts) and take the mean response from all directions.
+Of course, using the entire vector of feature responses with a multivariate likelihood is also possible.
+Moreover, in \cite{Kumar2010} as well as this paper, we will use the filter response $$R(x,y) = \max_{\sigma, \phi} \Delta G(\sigma, \phi) \ast I(x,y) $$ where $\Delta G(\sigma, \phi)$ is the Gaussian second derivative filter at scale $\sigma$ and orientation $\phi$.
+We use $R$ to (1) obtain knots, by passing to a Canny edge mapper, and (2) use as input to extraction functions (discussed below).
+Unlike other methods, we use the noisy edge response only as a guide for additional feature extraction, not as a testimony for the cell segments themselves.
+
+Let $T(I, x, y)$ be an $\emph{extraction function}$ which maps an image and $xy$-coordinates to a scalar.
+Suppose a line $\+{l}$ is given and $(x, y) = \+{l}(t)$ for some $t \in [t_i, t_{i+1}]$.
+Then the Radon-like response for image $I$ at $(x, y)$ in direction $\+{l}$ is given by $$\Psi \left( I,\+{l},t \right) = T \left( I, \+{l}(t) \right).$$ We will use the extraction function for cell background segmentation $$T_{\mbox{bg}}(I, \+{l}(t)) = \min_{t \in [t_i, t_{i+1}]} \left\{ I \left( \+{l}(t) \right) \right\}$$ which simply returns the dimmest pixel between any two knots \cite{Kumar2010}.
+Because knots denote edges, this feature tends to homogenize pixel intensities within a cell.
+
+Finally, we want pixels to not cross cell boundaries.
+The first step in encoding this constraint is to use a boundary-enhancing extraction function \cite{Kumar2010} $$T_{\mbox{bd}} = \frac{\int_t^{t+1} R\left( \+{l}(\tau) \right) d\tau }{\left\Vert \+{l}(t_{i+1}) - \+{l}(t_{i}) \right\Vert_2 }.$$ This function assigns to each pixel the average value of the Gaussian second-derivative response.
+Recall that the GSD tends to enhance boundaries, the most salient of which are captured by the Canny detector as knots.
+Thus, $T_{\mbox{bd}}$ smoothes the spurious GSD responses, avoiding oversegmentation.
+\fxfatal{Tighten up wording here}.
+Further, in connectomics data, cell membranes tend to form thick boundaries.
+But classic edge detectors work best at single-pixel gradients and return two edges for a thick boundary.
+\fxfatal{cite Rafael}.
+In particular, the Canny detector produces two knots on either side of a cell boundary.
+In the GSD response $R$, cell boundaries are consistently bright, so the integral in $T_{\mbox{bd}}$ returns a distinctly brighter response on boundaries, regardless of thickness.
+We use this boundary enhancement information in the next section.
+
+\subsection{Geodesic Distance}
+\label{sec:geodesic}
+Figure~\ref{fig:radon_bd} illustrates that cells need not be elliptical, or even convex.
+All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response.
+Therefore, shape constraints based on Euclidean distances are not justified.
+Instead, we define a \emph{geodesic distance}.
+Construct a graph whose vertices are the pixels and undirected edge for adjacent pixels ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$).
+Each edge has a weight $w(i,j) = \left| b_i - b_j \right|$ where $b_k$ is the boundary Radon-like response at pixel $k$.
+We define $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm.
+In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances.Address practical concerns.
+Further, because individual cells are tiny relative to an entire slice, only local distances are needed, because far points must be outside the current cell.
+All paths crossing a cell boundary must traverse pixels with low, high, and again low boundary feature responses.
+Thus, the shortest path wil also be one with a high weight.
+So if pixels $i$ and $j$ do not belong to the same cell, defined by the region enclosed by a loop of pixels with high boundary feature response, then $d(\+{x}_i, \+{x}_j)$ is high.
+
+Recall that the cluster mean contains entries for the $xy$-coordinates $\+{\mu}_{k,1:2} = (\mu_{k,x}, \mu_{k,y})$.
+Fix a point $n$ and a cluster $k$.
+Then $d(\+{x}_n, \+{\mu}_{k,1:2})$ is exponentially distributed, favoring distances close to zero.
+Any point one the opposite side of a boundary from the cluster's center will be far away from zero, and hence have low probability.
+Geodesic distances to other points are ignored.
+
+We model a two-dimensional image as an infinite semi-Gaussian mixture model.
+This extends the classic Gaussian mixture model (GMM) with (a) a variable number of classes and (b) additional naive-Bayes features associated with each class.
+
+\begin{enumerate}
+\item $\phi_{1\cdot} \in [0,1]$: pixel intensity 
+\item $\phi_{2\cdot} \in [0,1]$: average interior Radon-like feature response
+\item $\phi_{3\cdot} \in [0,1]$: geodesic distance  
+\item 
+\end{enumerate}
  Because cell interiors contain diverse structures, intensities values within a cell vary widely, so relying on $\phi_{1\cdot}$ alone oversegments.

### 2013-01-31 src/feature_map.m

@@ -4,7 +4,7 @@ function [ PHI, misc_data ] = feature_map( img )
     
     % Number of features
     D = 5;   
-    randon_angles = 72;
+    randon_angles = 180;
     canny_thresh = 0.25;
     
     PHI = zeros(N, D);

### 2013-01-31 latex/probabilistic-models-tikz-v2.tex
graphical model set

%% LyX 1.6.8 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[oneside,english,american,pdftex, a4paper, twoside, BCOR12mm, titlepage, openright, headsepline, chapterprefix, mpexclude,  idxtotoc, liststotoc, cleardoublestandard]{scrbook}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\setlength{\parskip}{\medskipamount}
\setlength{\parindent}{0pt}
\usepackage{color}
\usepackage{array}
\usepackage{verbatim}
\usepackage{float}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{esint}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}[chapter]
\floatname{algorithm}{Algorithm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\floatstyle{ruled}
\newfloat{algorithm}{tbp}{loa}
\floatname{algorithm}{Algorithm}
\usepackage[noend]{algorithmic}
\newcommand{\forbody}[1]{ #1 \ENDFOR}
\newcommand{\ifbody}[1]{ #1  \ENDIF}
\newcommand{\whilebody}[1]{ #1  \ENDWHILE}
\renewcommand{\algorithmicprint}{\textbf{draw}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

\usepackage{euscript}

\DeclareSymbolFont{rsfscript}{OMS}{rsfs}{m}{n}
\DeclareSymbolFontAlphabet{\mathrsfs}{rsfscript}

% PDF formatting instructions for A4
\pdfpagewidth=210mm % for pdflatex
\pdfpageheight=296mm % for pdflatex

%%%%%%%% begin tikz %%%%%%
\usepackage{tikz,tkz-base}
\usetikzlibrary{shapes,decorations,shadows}
\usetikzlibrary{decorations.pathmorphing}
\usetikzlibrary{decorations.shapes}
\usetikzlibrary{fadings}
\usetikzlibrary{patterns}
\usetikzlibrary{calc}
\usetikzlibrary{decorations.text}
\usetikzlibrary{decorations.footprints}
\usetikzlibrary{decorations.fractals}
\usetikzlibrary{shapes.gates.logic.IEC}
\usetikzlibrary{shapes.gates.logic.US}
\usetikzlibrary{fit,chains}
\usetikzlibrary{positioning}
\usepgflibrary{shapes}
\usetikzlibrary{scopes}
\usetikzlibrary{arrows}
\usetikzlibrary{backgrounds}

\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
minimum size=20pt, font=\fontsize{10}{10}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt},
factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=0pt},
>={triangle 45}}

\pgfdeclarelayer{b}
\pgfdeclarelayer{f}
\pgfsetlayers{b,main,f}

% shapename, fitlist, caption, pos
\newcommand{\plate}[4]{
\begin{pgfonlayer}{b}
\node (invis#1) [draw, color=white, inner sep=2pt,rectangle,fit=#2] {};
\end{pgfonlayer}\begin{pgfonlayer}{f}
\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=2pt,yshift=1pt] {\footnotesize{#3}};
\node (#1) [draw,inner sep=1pt, rectangle,fit=(invis#1) (capt#1),#4] {};
\end{pgfonlayer}
}

\newcommand{\shiftedplate}[5]{
\begin{pgfonlayer}{b}
\node (invis#1) [draw, color=white, inner sep=0 pt,rectangle,fit=#2] {};
\end{pgfonlayer}\begin{pgfonlayer}{f}
\node (capt#1) [#5, xshift=2pt] {\footnotesize{#3}};
\node (#1) [draw,inner sep=2pt, rectangle,fit=(invis#1) (capt#1),#4] {};
\end{pgfonlayer}
}

%shapename, pos, caption, in1, in2, out, captpos
\newcommand{\twofactor}[7]{
\node (#1) [factor] at #2 {};
\node (capt#1) [#7 of #1]{\footnotesize{#3}};
\draw [-] (#4) -- (#1) ;
\draw [-] (#5) -- (#1) ;
\draw [->,thick] (#1) -- (#6);
}

%shapename, pos, caption, in, out, captpos
\newcommand{\factor}[6]{
\node (#1) [factor] at #2 {};
\node (capt#1) [#6 of #1]{\footnotesize{#3}};
\draw [-] (#4) -- (#1) ;
\draw [->,thick] (#1) -- (#5);
}

% name, --, caption, pos
\newcommand{\nofactor}[4]{
\node (#1) [factor, #2]  {};
\node (capt#1) [#4 of #1]{\footnotesize{#3}};
}

%shapename,  fitlist, caption
\newcommand{\namedgate}[3]{
\begin{pgfonlayer}{b}
\node (invisgate#1) [rectangle, draw, color=white,  fit=#2] {};
\end{pgfonlayer}
\node (gatecapt#1) [ above right=0 pt of invisgate#1.north west, xshift=-1pt ] {\footnotesize{#3}};
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgate#1)(gatecapt#1)]{};

}

%shapename,  fitlist, caption
\newcommand{\gate}[3]{
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=#2]{};
}

%shapename,  fitlist1, fitlist2, caption1, caption2
\newcommand{\vertgate}[5]{
\begin{pgfonlayer}{b}
\node (invisgateleft#1) [rectangle, draw, color=white,  fit=#2] {};
\node (invisgateright#1) [rectangle, draw, color=white,  fit=#4] {};
\end{pgfonlayer}
\node (gatecaptleft#1) [ above left=0 pt of invisgateleft#1.north east, xshift=1pt ]{\footnotesize{#3}};
\node (gatecaptright#1) [ above right=0 pt of invisgateright#1.north west, xshift=-1pt ] {\footnotesize{#5}};
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(gatecaptleft#1)(invisgateright#1)(gatecaptright#1)]{};
\draw [-, dashed] (#1.north) -- (#1.south);
}

\newcommand{\vertgateSpec}[5]{
\begin{pgfonlayer}{b}
\node (invisgateleft#1) [rectangle, draw, color=white,  fit=#2] {};
\node (invisgateright#1) [rectangle, draw, color=white,  fit=#4] {};
\end{pgfonlayer}
\node (gatecaptleft#1) [ above left=0 pt of invisgateleft#1.north east, xshift=1pt ]{\footnotesize{#3}};
\node (gatecaptright#1) [ above right=0 pt of invisgateright#1.north west, xshift=-1pt ] {\footnotesize{#5}};
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(gatecaptleft#1)(invisgateright#1)(gatecaptright#1)]{};
\draw [-, dashed] (#1.70) -- (#1.290);
}

\newcommand{\horgate}[5]{
\begin{pgfonlayer}{b}
\node (invisgateleft#1) [rectangle, draw, color=white,  fit=#2] {};
\node (invisgateright#1) [rectangle, draw, color=white,  fit=#4] {};
\end{pgfonlayer}
\node (gatecaptleft#1) [ above right=0 pt of invisgateleft#1.south west, xshift=1pt ]{\footnotesize{#3}};
\node (gatecaptright#1) [ below right=0 pt of invisgateright#1.north west, xshift=-1pt ] {\footnotesize{#5}};
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(gatecaptleft#1)(invisgateright#1)(gatecaptright#1)]{};
\draw [-, dashed] (#1.west) -- (#1.east);
}

\newcommand{\horogate}[5]{
\begin{pgfonlayer}{b}
\node (invisgateleft#1) [rectangle, draw, color=white,  fit=#2] {};
\node (invisgateright#1) [rectangle, draw, color=white,  fit=#4] {};
\end{pgfonlayer}
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(invisgateright#1)]{};
\node (gatecaptleft#1) [ above right=0 pt of #1.west, xshift=0pt ]{\footnotesize{#3}};
\node (gatecaptright#1) [ below right=0 pt of #1.west, xshift=0pt ] {\footnotesize{#5}};

\draw [-, dashed] (#1.west) -- (#1.east);
}

\newcommand{\vertogate}[5]{
\begin{pgfonlayer}{b}
\node (invisgateleft#1) [rectangle, draw, color=white,  fit=#2] {};
\node (invisgateright#1) [rectangle, draw, color=white,  fit=#4] {};
\end{pgfonlayer}
\node (#1) [rectangle,draw,dashed, inner sep=2pt, fit=(invisgateleft#1)(invisgateright#1)]{};
\node (gatecaptleft#1) [ below left=0 pt of #1.north, xshift=0pt ]{\footnotesize{#3}};
\node (gatecaptright#1) [ below right=0 pt of #1.north, xshift=0pt ] {\footnotesize{#5}};

\draw [-, dashed] (#1.north) -- (#1.south);
}

%%%%%%%% end tikz %%%%%%

\makeatother

\usepackage{babel}

\begin{document}

\title{Probabilistic Models in Tikz}

\author{Laura Dietz}

\maketitle
Tikz macros and examples for probabilistic models. I use the directed
factor graph notation in most figures, but you can use the same macros
to draw Bayesian networks.

\selectlanguage{english}%
\global\long\def\items{\EuScript{V}}

\global\long\def\tastes{\CMcal{T}}

\global\long\def\nodes{\EuScript{N}}

\global\long\def\edges{\EuScript{E}}

\global\long\def\content{\mathrsfs{C}}

\global\long\def\vocab{\CMcal{V}}

\global\long\def\docs{\CMcal{D}}

\global\long\def\cites{\CMcal{C}}

\global\long\def\up#1{\uppercase{#1}}

\chapter{Foundations}

%
\begin{figure}[H]
\noindent \begin{centering}
\begin{tabular}{cc}
%
\begin{comment}
p(X|Z,R,Q)\textbackslash{}cdot p(R|Z,Q)\textbackslash{}cdot p(Z|Q)\textbackslash{}cdot
p(Q)
\end{comment}
{}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt] {
&\node (q) [latent] {$Q$};&\\
\node (z) [latent] {$Z$};
&\node (qrzx) [factor,below=3pt] {};&
\node (r) [latent] {$R$};\\
&\node (x) [latent] {$X$};&\\
};

\node (qfac) [factor, above=10pt of q] {};
\draw [-] (q) -- (qfac);

\draw [-] (q) -- (z) node (qz) [midway,factor] {};
\draw [-] (q) -- (r) node (qrz) [midway,factor] {};
\draw [-] (qrz) -- (z);

\draw [-] (r) -- (qrzx) -- (x);
\draw [-] (z) -- (qrzx);
\draw [-] (q) -- (qrzx);

\end{tikzpicture} & %
\begin{comment}
p(X|Z)\textbackslash{}cdot p(Z|Q)\textbackslash{}cdot p(R|Q)\textbackslash{}cdot
p(Q)
\end{comment}
{}

\noindent \centering{}\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt] {
&\node (q) [latent] {$Q$};&\\
\node (z) [latent] {$Z$};
&\node (qrzx) [factor,below=3pt] {};&
\node (r) [latent] {$R$};\\
&\node (x) [latent] {$X$};&\\
};

\node (qfac) [factor, above=10pt of q] {};
\draw [-] (q) -- (qfac);

\draw [-] (q) -- (z) node (qz) [midway,factor] {};
\draw [-] (q) -- (r) node (qrz) [midway,factor] {};

\draw [-] (z) -- (qrzx) -- (x);

\end{tikzpicture}\tabularnewline
a) & b)\tabularnewline
\end{tabular}
\par\end{centering}

\caption[Factor graphs for visualizing factorizations of a joint distribution.]{Factor graphs for visualizing factorizations of a joint distribution
$p(X,Z,R,Q)$. a) General factorization, b) Factorization with conditional
independence assumptions.\label{fig:Factorization}}

\end{figure}

%
\begin{figure}[H]
\noindent \begin{centering}
\begin{tabular}{cc}
\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt] {
&\node (q) [latent] {$Q$};&\\
\node (z) [latent] {$Z$};
&&
\node (r) [latent] {$R$};\\
&\node (x) [latent] {$X$};&\\
};

\draw [->] (q) -- (z) ;
\draw [->] (q) -- (r) ;
\draw [->] (q) -- (z);

\draw [->] (z) -- (r);

\draw [->] (r) -- (x);
\draw [->] (z) -- (x);
\draw [->] (q) -- (x);

\end{tikzpicture} & \noindent \begin{centering}
%
\begin{comment}
p(X|Z)\textbackslash{}cdot p(Z|Q)\textbackslash{}cdot p(R|Q)\textbackslash{}cdot
p(Q)
\end{comment}
{}
\par\end{centering}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt] {
&\node (q) [latent] {$Q$};&\\
\node (z) [latent] {$Z$};
&&
\node (r) [latent] {$R$};\\
&\node (x) [latent] {$X$};&\\
};

\draw [->] (q) -- (z) ;
\draw [->] (q) -- (r) ;

\draw [->] (z)  -- (x);

\end{tikzpicture}\tabularnewline
a) & b)\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Bayesian network visualization for factor graphs in Figure \ref{fig:Factorization}.\label{fig:bayes}}

\end{figure}

%
\begin{figure}[H]
\noindent \begin{centering}
\begin{tabular}{cc}
%
\begin{comment}
p(X|Z,R,Q)\textbackslash{}cdot p(R|Z,Q)\textbackslash{}cdot p(Z|Q)\textbackslash{}cdot
p(Q)
\end{comment}
{}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt] {
&\node (q) [latent] {$Q$};&\\
\node (z) [latent] {$Z$};
&\node (qrzx) [factor,below=3pt] {};&
\node (r) [latent] {$R$};\\
&\node (x) [latent] {$X$};&\\
};

\node (qfac) [factor, above=10pt of q] {};
\draw [->] (qfac) -- (q);

\draw [->] (q) -- (z) node (qz) [midway,factor] {};
\draw [->] (q) -- (r) node (qrz) [midway,factor] {};
\draw [-] (qrz) -- (z);

\draw [->] (r) -- (qrzx) -- (x);
\draw [-] (z) -- (qrzx);
\draw [-] (q) -- (qrzx);

\end{tikzpicture} & %
\begin{comment}
p(X|Z)\textbackslash{}cdot p(Z|Q)\textbackslash{}cdot p(R|Q)\textbackslash{}cdot
p(Q)
\end{comment}
{}

\noindent \centering{}\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=20pt, row sep=10pt] {
&\node (q) [latent] {$Q$};&\\
\node (z) [latent] {$Z$};
&\node (qrzx) [factor,below=3pt] {};&
\node (r) [latent] {$R$};\\
&\node (x) [latent] {$X$};&\\
};

\node (qfac) [factor, above=10pt of q] {};
\draw [->] (qfac) -- (q);

\draw [->] (q) -- (z) node (qz) [midway,factor] {};
\draw [->] (q) -- (r) node (qrz) [midway,factor] {};

\draw [->] (z) -- (qrzx) -- (x);

\end{tikzpicture}\tabularnewline
a) & b)\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Directed factor graph visualization for factor graphs in Figure \ref{fig:Factorization}.\label{fig:dir-factorgraph}}

\end{figure}

%
\begin{figure}[H]
\noindent \begin{centering}
\begin{tabular}{cc}
%
\begin{comment}
gate: Z=z

p(X|Y) p(R|Z)

gate Z!=z

p(X,R|Z)
\end{comment}
{}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=30pt, row sep=10pt] {

&\node (z) [latent] {$Z$};&
\\[20pt,between borders]

\node (xz) [factor] {};
\node (rz) [factor,right = 10 pt of xz] {};
\node (spacexz) [rectangle, draw, transparent,  above = 5pt of xz] {};
&& \node (rzx) [factor] {};
\\

\node (x) [latent] {$X$};
&& \node (r) [latent] {$R$};
\\
};

\draw [->] (z) -- (xz) -- (x);
\draw [->] (z) -- (rz) -- (r);
\draw [->] (z) -- (rzx) -- (x);
\draw [->] (rzx) -- (r);

%shapename,  fitlist1, fitlist2, caption1, caption2
\vertogate{g}{(xz)(rz)(spacexz)}{$Z=z$}{(rzx)}{$Z\neq z$};

\draw [-*,dashed] (z) -- (g);
\end{tikzpicture} & %
\begin{comment}
gate Q=0

p(X|R)

gate Q=1

p(X|Z)
\end{comment}
{}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=30pt, row sep=10pt] {

\node (r) [latent] {$R$};
&\node (q) [latent] {$Q$};
&\node (z) [latent] {$Z$};
\\[20pt,between borders]

\node (xr) [factor] {};
\node (spacexr) [rectangle, draw, transparent,  above = 5pt of xr] {};
&& \node (xz) [factor] {};
\\

&\node (x) [latent] {$X$};
&
\\
};

\draw [->] (r) -- (xr) -- (x);
\draw [->] (z) -- (xz) -- (x);

%shapename,  fitlist1, fitlist2, caption1, caption2
\vertogate{g}{(xr)(spacexr)}{$Q=0$}{(xz)}{$Q=1$};
\draw [-*,dashed] (q) -- (g);
\end{tikzpicture}\tabularnewline
a) & b)\tabularnewline
\end{tabular}
\par\end{centering}

\caption[Representing context-specific independence by gates in factor graphs.]{Representing context-specific independence by gates (dashed boxes)
in factor graphs.\label{fig:factors-and-gates}}

\end{figure}

%
\begin{figure}[H]
\noindent \begin{centering}
\begin{tabular}{c}
%
\begin{comment}
p(theta)

plate: i=1..n

p(X|theta )
\end{comment}
{}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=30pt, row sep=10pt] {

\node (theta) [latent] {$\theta$};
\\

\node (xtheta) [factor] {};
\node (spaceleft) [rectangle, draw, transparent,  left = 25pt of xtheta] {};
\node (spaceright) [rectangle, draw, transparent,  right = 25pt of xtheta] {};
\\

\node (x) [latent] {$X$};
\\
};
\plate{p}{(xtheta)(spaceleft)(spaceright)(x)}{$\forall i=1\ldots n$}{}

\draw [->] (theta) -- (xtheta) -- (x);

\end{tikzpicture}\tabularnewline
\end{tabular}
\par\end{centering}

\caption[Representing repetition of connection templates by plates in factor
graphs.]{Representing repetition of connection templates by plates (solid
boxes) in factor graphs.\label{fig:plates}}

\end{figure}

%
\begin{figure}[H]
\noindent \begin{centering}
\begin{tabular}{c}
%
\begin{comment}
p(theta|alpha, beta)

plate: i=1..n

p(X|theta )
\end{comment}
{}

\begin{tikzpicture}
\node [matrix,matrix anchor=mid, column sep=30pt, row sep=10pt] {

\node (theta) [latent] {$\theta$};
\\

\nofactor{xtheta}{}{Bern}{right=0pt};

\node (spaceleft) [rectangle, draw, transparent,  left = 25pt of xtheta] {};
\node (spaceright) [rectangle, draw, transparent,  right = 25pt of xtheta] {};
\\

\node (x) [obs] {$X$};
\\
};
\plate{p}{(xtheta)(spaceleft)(spaceright)(x)}{$\forall i=1\ldots n$}{}

\draw [->] (theta) -- (xtheta) -- (x);

\nofactor{drawtheta}{above=10pt of theta}{Beta}{right=0pt};
\node (alpha) [const, above left=10pt and 8 pt of drawtheta]  {$\alpha$};
\node (beta) [const, above right=10pt and 8 pt of drawtheta]  {$\beta$};

\draw [->] (alpha) -- (drawtheta) -- (theta);
\draw [-] (beta) -- (drawtheta);

\end{tikzpicture}\tabularnewline
\end{tabular}
\par\end{centering}

\caption{Beta-Bernoulli example with hyperparameters.\label{fig:beta-bernoulli-example-hyperparams}}

\end{figure}

\section{Latent Dirichlet Allocation}

\selectlanguage{american}%
%
\begin{figure}[H]
\selectlanguage{english}%
\noindent \begin{centering}
% Latent Dirichlet allocation
\begin{tikzpicture}
[latent/.style={circle,fill=white,draw=black,inner sep=1pt, minimum size=20pt, font=\fontsize{14}{14}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt}]

\matrix[row sep=4mm,column sep=10mm, matrix anchor=mid] (lda) {
&

\\

&

\\

\node (lambda) [latent] {$\theta$}; 
&

\\

\nofactor{drawtc}{}{Multi}{left=0pt};
&

\\

\node (tc) [latent]  {$T$}; 
&

\\

\nofactor{drawxc}{}{Multi}{above=0pt}; \gate{gatexc}{(drawxc) (captdrawxc)}{$t$};
&  \node (phi) [latent]  {$\phi$};

\\

\node (xc) [obs]  {$X$}; 
&

\\
};

\nofactor{drawlambda}{above=10pt of lambda}{Dir}{right=0pt};
\nofactor{drawphi}{above=10pt of phi}{Dir}{right=0pt};

\node (alphalambda) [const, above=10pt of drawlambda]  {$\alpha_{\theta}$};
\node (alphaphi) [const, above=10pt of drawphi]  {$\alpha_{\phi}$};

\draw [->] (alphaphi) -- (drawphi) -- (phi);
\draw [->] (alphalambda) -- (drawlambda) -- (lambda);

\draw [->] (lambda) -- (drawtc) -- (tc);
\draw [->] (phi) -- (drawxc) -- (xc);

\draw [-*] (tc) -- (gatexc);

\plate{contentplatec}{(tc)(xc)(drawtc)(drawxc)(captdrawtc)(captdrawxc)(gatexc)}{$\forall 1 \leq i \leq n_d  $}{}
\plate{userplatec}{(contentplatec)(lambda)(drawlambda)(captdrawlambda)}{$\forall d \in \mathcal{D}$}{}

\plate{topicplate}{(phi)(drawphi)(captdrawphi)}{$\forall t \in \tastes$}{}

\end{tikzpicture}

\par\end{centering}

\selectlanguage{american}%
\caption{Latent Dirichlet allocation as directed factor graph.\label{fig:lda-factorgraph}}

\end{figure}

\chapter{Citation Influence Model}

\section{Plain Citation Influence Model}

%
\begin{figure}[H]
\selectlanguage{english}%
\noindent \begin{centering}
% Citation Influence
\begin{tikzpicture}
[latent/.style={circle,fill=white,draw=black,inner sep=1pt, minimum size=20pt, font=\fontsize{14}{14}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt}]

\matrix[row sep=4mm,column sep=10mm, matrix anchor=mid] (citinf) {
&
&[3mm] \node (psi) [latent] {$\gamma$}; 
\\

&
& \nofactor{drawf}{}{Multi}{left=0pt};
\\

\node (lambda) [latent] {$\theta$}; 
&
& \node (f) [latent]  {$C$}; 
\\

\nofactor{drawtc}{}{Multi}{left=0pt};
&
& \nofactor{drawtd}{}{Multi}{above=0pt};\gate{gatetd}{(drawtd) (captdrawtd)}{$\tastes$};
\\

\node (tc) [latent]  {$T'$}; 
&
& \node (td) [latent]  {$T$}; 
\\

\nofactor{drawxc}{}{Multi}{above=0pt}; \gate{gatexc}{(drawxc) (captdrawxc)}{$t$};
&  \node (phi) [latent]  {$\phi$};
& \nofactor{drawxd}{}{Multi}{above=0pt}; \gate{gatexd}{(drawxd) (captdrawxd)}{$t$};
\\

\node (xc) [obs]  {$W'$}; 
&
&\node (xd) [obs]  {$W$}; \\
\\
};

\nofactor{drawpsi}{left=10pt of psi}{Dir}{below=0pt};
\nofactor{drawlambda}{above=3pt of lambda}{Dir}{right=0pt};
\nofactor{drawphi}{above=3pt of phi}{Dir}{right=0pt};

\node (alphapsi) [const, left=18pt of drawpsi] {$\alpha_{\gamma}$}; 
\node (alphalambda) [const, above=10pt of drawlambda]  {$\alpha_{\theta}$};
\node (alphaphi) [const, above=10pt of drawphi]  {$\alpha_{\phi}$};

\draw [->] (alphaphi) -- (drawphi) -- (phi);
\draw [->] (alphalambda) -- (drawlambda) -- (lambda);
\draw [->] (alphapsi) -- (drawpsi) -- (psi);

\draw [->] (psi) -- (drawf) -- (f);
\draw [->] (lambda) -- (drawtc) -- (tc);
\draw [->] (lambda) -- (drawtd) -- (td);
\draw [->] (phi) -- (drawxc) -- (xc);
\draw [->] (phi) -- (drawxd) -- (xd);

\draw [-*] (tc) -- (gatexc);
\draw [-*] (td) -- (gatexd);
\draw [-*] (f) -- (gatetd);

\plate{contentplatec}{(tc)(xc)(drawtc)(drawxc)(captdrawtc)(captdrawxc)(gatexc)}{$\forall w' \in c$}{}
\plate{userplatec}{(contentplatec)(lambda)(drawlambda)(captdrawlambda)}{$\forall c \in \mathcal{C}$}{}
\plate{contentplated}{(f)(td)(xd)(drawtd)(drawxd)(drawf)(captdrawtd)(captdrawxd)(captdrawf)(gatetd)(gatexd)}{$\forall w \in d$}{}
\plate{userplated}{(contentplated)(psi)(drawpsi)(captdrawpsi)}{$\forall d \in \mathcal{D}$}{}

\plate{topicplate}{(phi)(drawphi)(captdrawphi)}{$\forall t \in \tastes$}{}

\end{tikzpicture}

\par\end{centering}

\selectlanguage{american}%
\caption{Citation influence model in directed factor graph notation.\label{fig:Citinf-plain-model}}

\end{figure}

%
\begin{algorithm}[H]
%
\begin{algorithmic}
[1]\selectlanguage{english}%
\FORALL{$t\in\tastes$}

\forbody{\PRINT{$\phi_{t}\sim\mbox{Dir}(\alpha_{\phi})$}}

\FORALL{$c\in\mathcal{C}$ as cited role}

\forbody{\PRINT{$\theta_{c}\sim\mbox{Dir}(\alpha_{\theta})$}

\STATE{// Generate content $w$ according to LDA:}

\FORALL{words $w_{c,i}'\in c$}

\forbody{\PRINT{topic $\up t'_{c,i}\sim\mbox{Multi}(\theta_{c})$}

\PRINT{word $\up w'_{c,i}\sim\mbox{Multi}(\phi_{\up t'_{c,i}})$}}}

\FORALL{$d\in\mathcal{D}$ as citing role}

\forbody{\PRINT{$\gamma_{d}\sim\mbox{Dir}(\alpha_{\gamma})$}

\FORALL{words $w_{d,i}\in d$}

\forbody{\PRINT{citation $\up c_{d,i}\sim\mbox{Multi}(\gamma_{d})$}\label{Flo:dedication-begin}

\PRINT{topic $\up t_{d,i}\sim\mbox{Multi}(\theta_{\up c_{d,i}})$}

\PRINT{word $\up w_{d,i}\sim\mbox{Multi}(\phi_{\up t_{d,i}})$}\label{Flo:dedication-end}}}\selectlanguage{american}

\end{algorithmic}

\caption[Generative process of the plain citation influence model.]{Generative process of the plain citation influence model. For a description
of each variable see Table \ref{tab:Variable-description-plain}.\label{alg:Citinf-plain}}

\end{algorithm}

%
\begin{table}[H]
\caption[Notation for the plain citation influence model.]{Notation for the plain citation influence model. Variables in the
cited plate are denoted with prime.}

\label{tab:Variable-description-plain}

{\small }\begin{tabular}{|>{\centering}p{0.11\linewidth}|>{\raggedright}p{0.79\linewidth}|}
\hline 
{\small Symbol} & {\small Description}\tabularnewline
\hline
\hline 
{\small $\up c$, $c$} & {\small cited publication}\tabularnewline
\hline 
{\small $d$} & {\small citing publication}\tabularnewline
\hline 
{\small $\theta$} & {\small shared topic mixture associated with a cited publication}\tabularnewline
\hline 
{\small $\phi$} & {\small characteristic word distribution for each topic}\tabularnewline
\hline 
{\small $\gamma$} & {\small distribution of citation influences}\tabularnewline
\hline 
{\small $\up w$', $\up w$} & {\small words in cited, citing publications respectively}\tabularnewline
\hline 
{\small $\up t$', $\up t$} & {\small topic assignments of tokens in cited, citing publications
respectively}\tabularnewline
\hline 
{\small $\alpha$} & {\small Dirichlet / beta parameters of the multinomial / Bernoulli
distributions}\tabularnewline
\hline
\end{tabular}
\end{table}

\section{Citation Influence Model with Own Topics}

%
\begin{figure}[H]
\selectlanguage{english}%
\noindent \begin{centering}
% Citation Influence
\begin{tikzpicture}
[latent/.style={circle,fill=white,draw=black,inner sep=1pt, minimum size=20pt, font=\fontsize{14}{14}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt}]

\matrix[row sep=4mm,column sep=10mm, matrix anchor=mid] (citinf) {
&
&[3mm] \node (psi) [latent] {$\gamma$};
&[12mm,between origins] \node (balance) [latent] {$\lambda$};
&[12mm,between origins] \node (innov) [latent] {$\psi$};
\\

&
& \nofactor{drawf}{}{Multi}{left=0pt};
&\nofactor{drawbalance}{}{Bern}{left=0pt};
&
\\

\node (lambda) [latent] {$\theta$}; 
&
& \node (f) [latent]  {$C$}; 
&\node (s) [latent]  {$S$}; 
&
\\

\nofactor{drawtc}{}{Multi}{left=0pt};
&
& \nofactor{drawtd}{}{Multi}{above=0pt};\gate{gatetd}{(drawtd) (captdrawtd)}{$\tastes$};
&
&\nofactor{drawtdinnov}{}{Multi}{below=0pt};
\\

\node (tc) [latent]  {$T'$}; 
&
& \node (td) [latent]  {$T$}; 
&
&
\\

\nofactor{drawxc}{}{Multi}{above=0pt}; \gate{gatexc}{(drawxc) (captdrawxc)}{$t$};
&  \node (phi) [latent]  {$\phi$};
& \nofactor{drawxd}{}{Multi}{above=0pt}; \gate{gatexd}{(drawxd) (captdrawxd)}{$t$};
&
&
\\

\node (xc) [obs]  {$W'$}; 
&
&\node (xd) [obs]  {$W$}; 
&
&
\\
\\
};

\vertogate{gates}{(drawtd)(captdrawtd)}{$S=0$}{(drawtdinnov)(captdrawtdinnov)}{$S=1$};

\nofactor{drawlambda}{above=8pt of lambda}{Dir}{right=0pt};
\node (alphalambda) [const, above=10pt of drawlambda]  {$\alpha_{\theta}$};

\nofactor{drawphi}{above=8pt of phi}{Dir}{right=0pt};
\node (alphaphi) [const, above=10pt of drawphi]  {$\alpha_{\phi}$};

\nofactor{drawpsi}{above=8pt of psi}{Dir}{left=0pt};
\node (alphapsi) [const, above=18pt of drawpsi] {$\alpha_{\gamma}$};

\nofactor{drawbalancep}{above=8pt of balance}{Beta}{left=0pt};
\node (alphabalance1) [const, above left=18pt and 8 pt of drawbalancep] {$\alpha_{\lambda_{\theta}}$}; 
\node (alphabalance2) [const, above right=18pt and 8 pt of drawbalancep] {$\alpha_{\lambda_{\psi}}$};

\nofactor{drawinnovp}{above=8pt of innov}{Dir}{left=0pt};
\node (alphainnov) [const, above=18pt of drawinnovp] {$\alpha_{\psi}$};

\draw [->] (alphaphi) -- (drawphi) -- (phi);
\draw [->] (alphalambda) -- (drawlambda) -- (lambda);
\draw [->] (alphapsi) -- (drawpsi) -- (psi);
\draw [->] (alphabalance1) -- (drawbalancep) -- (balance);
\draw [-] (alphabalance2) -- (drawbalancep);
\draw [->] (alphainnov) -- (drawinnovp) -- (innov);

\draw [->] (psi) -- (drawf) -- (f);
\draw [->] (lambda) -- (drawtc) -- (tc);
\draw [->] (lambda) -- (drawtd) -- (td);
\draw [->] (phi) -- (drawxc) -- (xc);
\draw [->] (phi) -- (drawxd) -- (xd);
\draw [->] (balance) -- (drawbalance) -- (s);
\draw [->] (innov) -- (drawtdinnov) -- (td);

\draw [-*] (tc) -- (gatexc);
\draw [-*] (td) -- (gatexd);
\draw [-*] (f) -- (gatetd);
\draw [-*] (s) -- (gates);

\plate{contentplatec}{(tc)(xc)(drawtc)(drawxc)(captdrawtc)(captdrawxc)(gatexc)}{$\forall w' \in c$}{}
\plate{userplatec}{(contentplatec)(lambda)(drawlambda)(captdrawlambda)}{$\forall c \in \mathcal{C}$}{}
\plate{contentplated}{(f)(td)(xd)(drawtd)(drawxd)(drawf)(captdrawtd)(captdrawxd)(captdrawf)(gatetd)(gatexd)(gates)(s)(drawbalance)}{$\forall w \in d$}{}
\plate{userplated}{(contentplated)(psi)(drawpsi)(captdrawpsi)(balance)(drawbalance)(captdrawbalance)(innov)}{$\forall d \in \mathcal{D}$}{}

\plate{topicplate}{(phi)(drawphi)(captdrawphi)}{$\forall t \in \tastes$}{}

\end{tikzpicture}

\par\end{centering}

\selectlanguage{american}%
\caption{Citation influence model with own topics as directed factor graph.\label{fig:Citinf-owntopics-model}}

\end{figure}

%
\begin{algorithm}[H]
%
\begin{algorithmic}
[1]\selectlanguage{english}%
\FORALL{$t\in\tastes$}

\forbody{\PRINT{$\phi_{t}\sim\mbox{Dir}(\alpha_{\phi})$ ranging over the vocabulary
$\vocab$.}}

\FORALL{cited $c\in\mathcal{C}$}

\forbody{\PRINT{topic mixture $\theta_{c}\sim\mbox{Dir}(\alpha_{\theta})$ ranging
over topics $\tastes$.}

\STATE{// Generate content $w$ according to LDA:}

\FORALL{words $w_{c,i}\in c$}

\forbody{\PRINT{topic $\up t'_{c,i}\sim\mbox{Multi}(\theta_{c})$}

\PRINT{word $\up w'_{c,i}\sim\mbox{Multi}(\phi_{\up t'_{c,i}})$}}}

\FORALL{citing $d\in\mathcal{D}$ }

\forbody{\PRINT{$\gamma_{d}\sim\mbox{Dir}(\alpha_{\gamma})$ ranging over $L(d)$.}

\PRINT{$\lambda_{d}\sim\mbox{Beta}(\alpha_{\lambda_{\theta}},\alpha_{\lambda_{\psi}})$}

\PRINT{$\psi_{d}\sim\mbox{Dir}(\alpha_{\psi})$ ranging over topics $\tastes$.}

\FORALL{words $w_{d,i}\in d$}

\forbody{\PRINT{coin toss $\up s_{d,i}\sim\mbox{Bern}(\lambda_{d})$}

\selectlanguage{american}%
\IF{$\up s_{d,i}=0$}

\ifbody{\selectlanguage{english}%
\STATE{// Generate token by inheritance:}

\PRINT{a cited publication $\up c_{d,i}\sim\mbox{Multi}(\gamma_{d})$.}\label{Flo:inherit-begin}

\PRINT{topic $\up t_{d,i}\sim\mbox{Multi}(\theta_{\up c_{d,i}})$ \foreignlanguage{american}{from
the cited document's topic mixture.}}\label{Flo:inherit-end}

\selectlanguage{american}%
\ELSIF{$\up s_{d,i}=1$}

\selectlanguage{english}%
\STATE{// Generate token by innovation:}\label{Flo:innov-begin}

\PRINT{topic $\up t_{d,i}\sim\mbox{Multi}(\psi_{d})$ from the own topic
mixture.}\label{Flo:innov-end}\selectlanguage{american}
}

\selectlanguage{english}%
\PRINT{word $\up w_{d,i}\sim\mbox{Multi}(\phi_{\up t_{d,i}})$}}}\selectlanguage{american}

\end{algorithmic}

\caption{Generative process of the citation influence model with own topics.\label{alg:Citinf-owntopics}}

\end{algorithm}

%
\begin{table}[tb]
\caption{Further notation used in the citation influence model with own topics.}

\label{tab:Variable-description-owntopics}

{\small }\begin{tabular}{|>{\centering}p{0.11\linewidth}|>{\raggedright}p{0.79\linewidth}|}
\hline 
{\small Symbol} & {\small Description}\tabularnewline
\hline
\hline 
{\small $\psi$} & {\small innovation topic mixture of a citing publication}\tabularnewline
\hline 
{\small $\lambda$} & {\small parameter of the coin flip, choosing to draw topics from $\theta$
or $\psi$}\tabularnewline
\hline 
{\small $\up s$} & {\small indicates whether the topic of a citing publication is drawn
from inheritance or innovation}\tabularnewline
\hline
\end{tabular}
\end{table}

\chapter{Shared Taste Model}

\section{Plain Shared Taste Model}

%
\begin{table}[H]
\noindent \begin{centering}
\begin{tabular}{|>{\centering}p{0.2\textwidth}|>{\raggedright}p{0.7\textwidth}|}
\hline 
Symbol & Description\tabularnewline
\hline
\hline 
$\lambda_{\{u,f\}}$ & Shared topic mixture associated with the undirected edge between users
$u$ and $f$.\tabularnewline
\hline 
$T$ & Topic for items.\tabularnewline
\hline 
$\phi$ & Item distribution for each topic.\tabularnewline
\hline 
\selectlanguage{english}%
$\psi_{u}$\selectlanguage{american}
 & Distribution over friends of user $u$ according to their strength
of influence on the user's content.\tabularnewline
\hline 
$F$ & Friend associated with an observed item (if not explained by {}``own
topics'').\tabularnewline
\hline 
$X$ & Observed item in contents of the user.\tabularnewline
\hline 
$\Omega_{u}$ & Background topic mixture of the user (the own topics). Alternatively
a mixture over own items.\tabularnewline
\hline 
$\epsilon$ & Coin parameter deciding on how much content of the user is explained
by the {}``own topics''\tabularnewline
\hline 
$E$ & Boolean indicator, if 1, the observed item is explained by the {}``own
topics'' rather than the shared topics.\tabularnewline
\hline
\end{tabular}
\par\end{centering}

\caption{Notation for shared taste model with own topics.\label{tab:Notation-shared-taste}}

\end{table}

%
\begin{figure}[H]
\selectlanguage{english}%
\noindent \begin{centering}
% Shared Taste Model
\begin{tikzpicture}
[latent/.style={circle,fill=white,draw=black,inner sep=1pt, minimum size=20pt, font=\fontsize{14}{14}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt}]

\matrix[row sep=4mm,column sep=10mm, matrix anchor=mid] (sharedtaste) {

& 
& \node (psi) [latent] {$\psi$}; 
\\

&
& \nofactor{drawf}{}{Multi}{right=0pt};
\\

& 
& \node (f) [latent]  {$F$}; 
\\

\node (lambda) [latent] {$\lambda$};
& 
& \nofactor{drawt}{}{Multi}{above=0pt};\namedgate{gatet}{(drawt) (captdrawt)}{$e=\{u,F\}$};
\\

& 
& \node (t) [latent]  {$T$}; 
\\

\node (phi) [latent]  {$\phi$};
& 
& \nofactor{drawx}{}{Multi}{above=0pt}; \gate{gatex}{(drawx) (captdrawx)}{$t$};
\\

& 
& \node (x) [obs]  {$X$}; \\
};

\nofactor{drawpsi}{left=10pt of psi}{Dir}{below=0pt};
\nofactor{drawlambda}{above=10pt of lambda}{Dir}{right=0pt};
\nofactor{drawphi}{above=10pt of phi}{Dir}{right=0pt};

\node (alphapsi) [const, left=10pt of drawpsi] {$\alpha_{\psi}$}; 
\node (alphalambda) [const, above=10pt of drawlambda]  {$\alpha_{\lambda}$};
\node (alphaphi) [const, above=10pt of drawphi]  {$\alpha_{\phi}$};

\draw [->] (alphaphi) -- (drawphi) -- (phi);
\draw [->] (alphalambda) -- (drawlambda) -- (lambda);
\draw [->] (alphapsi) -- (drawpsi) -- (psi);

\draw [->] (psi) -- (drawf) -- (f);
\draw [->] (lambda) -- (drawt) -- (t);
\draw [->] (phi) -- (drawx) -- (x);

\draw [-*] (t) -- (gatex);
\draw [-*] (f) -- (gatet);

\plate{contentplate}{(f)(t)(x)(drawt)(drawx)(drawf)(captdrawt)(captdrawx)(captdrawf)(gatet)(gatex)}{$\forall x \in \content (u)$}{}
\plate{userplate}{(contentplate)(psi)(drawpsi)(captdrawpsi)}{$\forall u \in \nodes$}{}
\plate{topicplate}{(phi)(drawphi)(captdrawphi)}{$\forall t \in \tastes$}{}
\plate{edgeplate}{(lambda)(drawlambda)(captdrawlambda)}{$\forall e \in \edges$}{}

\end{tikzpicture}

\par\end{centering}

\selectlanguage{american}%
\caption{Shared taste model as directed factor graph.\label{fig:sharedtaste-model-factorgraph}}

\end{figure}

%
\begin{algorithm}[H]
%
\begin{algorithmic}
[1]\selectlanguage{english}%
\FORALL{$t\in\mathcal{T}$}

\forbody{\PRINT{$\phi_{t}\sim\mbox{Dir}(\alpha_{\phi})$}}

\FORALL{$\{u,f\}\in\mathcal{E}$}

\forbody{\PRINT{shared topic mixture $\lambda_{\{f,u\}}=\lambda_{\{u,f\}}\sim\mbox{Dir}(\alpha_{\lambda})$.}\label{Flo:sparse-topic-prior}}

\FORALL{$u\in\mathcal{N}$}

\forbody{\PRINT{friend mixture $\psi_{u}\sim\mbox{Dir}(\alpha_{\psi})$ ranging over
friends of $u$.}

\FORALL{items $x_{u,i}\in\mathcal{C}(u)$}\forbody{\PRINT{friend $F_{u,i}\sim\mbox{Multi}(\psi_{u})$}

\PRINT{topic $T_{u,i}\sim\mbox{Multi}(\lambda_{\{u,F_{u,i}\}})$}

\PRINT{item $X_{u,i}\sim\mbox{Multi}(\phi_{T_{u,i}})$}}}\selectlanguage{american}

\end{algorithmic}

\caption{\selectlanguage{english}%
Generative process of the shared taste model.\label{alg:Shared-Taste-Model}\selectlanguage{american}
}

\end{algorithm}

\section{Shared Taste Model with Own Topics}

%
\begin{algorithm}[H]
%
\begin{algorithmic}
[1]\selectlanguage{english}%
\FORALL{$t\in\mathcal{T}$}

\forbody{\PRINT{$\phi_{t}\sim\mbox{Dir}(\alpha_{\phi})$}}

\FORALL{$\{u,f\}\in\mathcal{E}$}

\forbody{\PRINT{shared topic mixture $\lambda_{\{f,u\}}=\lambda_{\{u,f\}}\sim\mbox{Dir}(\alpha_{\lambda})$.}}

\FORALL{$u\in\mathcal{N}$}

\forbody{\PRINT{friend mixture $\psi_{u}\sim\mbox{Dir}(\alpha_{\psi})$ ranging over
friends of $u$.}

\PRINT{background distribution $\Omega_{u}\sim\mbox{Dir}(\alpha_{\Omega})$}

\PRINT{background coin $\epsilon_{u}\sim\mbox{Beta}(a_{\epsilon},b_{\epsilon})$}

\FORALL{item $x_{u,i}\in\mathcal{C}(u)$}\forbody{\PRINT{coin flip $E_{u,i}\sim\mbox{Bern}(\epsilon_{u})$}

\IF{$E_{u,i}=0$}

\ifbody{\STATE{// Generate from shared topic.}

\PRINT{friend $F_{u,i}\sim\mbox{Multi}(\psi_{u})$}

\PRINT{topic $T_{u,i}\sim\mbox{Multi}(\lambda_{\{u,F_{u,i}\}})$ from share
topics}

\ELSIF{$E_{u,i}=1$}

\STATE{// Generate from background.}

\PRINT{topic $T_{u,i}\sim\mbox{Multi}(\Omega_{u})$ from own topics}}

\PRINT{item $X_{u,i}\sim\mbox{Multi}(\phi_{T_{u,i}})$}}}\selectlanguage{american}

\end{algorithmic}

\caption{\selectlanguage{english}%
Generative process of the shared taste model with own topics.\label{alg:Shared-Taste-Model-own-topics}\selectlanguage{american}
}

\end{algorithm}

%
\begin{figure}[H]
\selectlanguage{english}%
\noindent \begin{centering}
% Shared Taste Model own topics
\begin{tikzpicture}
[latent/.style={circle,fill=white,draw=black,inner sep=1pt, minimum size=20pt, font=\fontsize{14}{14}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt}]

\matrix[row sep=4mm,column sep=10mm, matrix anchor=mid] (sharedtaste) {

& 
& \node (psi) [latent] {$\psi$}; 
&[12mm,between origins] \node (balance) [latent] {$\epsilon$};
&[12mm,between origins] \node (innov) [latent] {$\Omega$};
\\

&
& \nofactor{drawf}{}{Multi}{right=0pt};
& \nofactor{drawe}{}{Bern}{right=0pt};
& 
\\

& 
& \node (f) [latent]  {$F$}; 
& \node (e) [latent]  {$E$}; 
& 
\\

\node (lambda) [latent] {$\lambda$};
& 
& \nofactor{drawt}{}{Multi}{above=0pt};\namedgate{gatet}{(drawt) (captdrawt)}{$e=\{u,F\}$};
&
&\nofactor{drawtinnov}{}{Multi}{below=0pt};
\\

& 
& \node (t) [latent]  {$T$}; 
&
&
\\

\node (phi) [latent]  {$\phi$};
& 
& \nofactor{drawx}{}{Multi}{above=0pt}; \gate{gatex}{(drawx) (captdrawx)}{$t$};
&
&
\\

& 
& \node (x) [obs]  {$X$}; 
&
& 
\\
};

\vertogate{gates}{(gatet)}{$E=0$}{(drawtinnov)(captdrawtinnov)}{$E=1$};

\nofactor{drawpsi}{above=10pt of psi}{Dir}{left=0pt};
\nofactor{drawlambda}{above=10pt of lambda}{Dir}{right=0pt};
\nofactor{drawphi}{above=10pt of phi}{Dir}{right=0pt};
\nofactor{drawbalance}{above=10pt of balance}{Beta}{left=0pt};
\nofactor{drawinnov}{above=10pt of innov}{Dir}{left=0pt};

\node (alphapsi) [const, above=18pt of drawpsi] {$\alpha_{\psi}$}; 
\node (alphalambda) [const, above=18pt of drawlambda]  {$\alpha_{\lambda}$};
\node (alphaphi) [const, above=10pt of drawphi]  {$\alpha_{\phi}$};
\node (abalance) [const, above left=18pt and 8 pt of drawbalance]  {$a_{\epsilon}$};
\node (bbalance) [const, above right=18pt and 8 pt of drawbalance]  {$b_{\epsilon}$};
\node (alphainnov) [const, above=18pt of drawinnov]  {$\alpha_{\Omega}$};

\draw [->] (alphaphi) -- (drawphi) -- (phi);
\draw [->] (alphalambda) -- (drawlambda) -- (lambda);
\draw [->] (alphapsi) -- (drawpsi) -- (psi);
\draw [->] (alphainnov) -- (drawinnov) -- (innov);
\draw [->] (abalance) -- (drawbalance) -- (balance);
\draw [-] (bbalance) -- (drawbalance);

\draw [->] (psi) -- (drawf) -- (f);
\draw [->] (lambda) -- (drawt) -- (t);
\draw [->] (phi) -- (drawx) -- (x);
\draw [->] (balance) -- (drawe) -- (e);
\draw [->] (innov) -- (drawtinnov) -- (t);

\draw [-*] (t) -- (gatex);
\draw [-*] (f) -- (gatet.101);
\draw [-*] (e) -- (gates);

\plate{contentplate}{(gates)(drawe)(f)(t)(x)(drawt)(drawx)(drawf)(captdrawt)(captdrawx)(captdrawf)(gatet)(gatex)}{$\forall x \in \content (u)$}{}
\plate{userplate}{(contentplate)(psi)(drawpsi)(captdrawpsi)(drawbalance)(captdrawbalance)(drawinnov)(captdrawinnov)}{$\forall u \in \nodes$}{}
\plate{topicplate}{(phi)(drawphi)(captdrawphi)}{$\forall t \in \tastes$}{}
\plate{edgeplate}{(lambda)(drawlambda)(captdrawlambda)}{$\forall e \in \edges$}{}

\end{tikzpicture}

\par\end{centering}

\selectlanguage{american}%
\caption{Shared taste model with own topics as directed factor graph.\label{fig:sharedtaste-owntopics-model-factorgraph}}

\end{figure}

\section{Pairwise Link LDA model (Nallapati 08)}

The original pairwise Link LDA model (Nallapati, Ahmed, Xing, Cohen.
Joint latent topic models for text and citations, KDD, 2008) is modified
to correct for rarity of interactions (Airoldi, Blei, Fienberg, Xing.
Mixed membership stochastic blockmodels, JMLR, 2008).

%
\begin{figure}[H]
\selectlanguage{english}%
\noindent \begin{centering}
% Pairwise Link LDA model (Nallapati 08)
\begin{tikzpicture}
[latent/.style={circle,fill=white,draw=black,inner sep=1pt, minimum size=20pt, font=\fontsize{14}{14}\selectfont},
obs/.style={latent,fill=gray!25},
const/.style={rectangle, inner sep=0pt}]

%[10mm,between origins]  
\node [matrix,row sep=4mm,column sep=4mm, matrix anchor=mid]  (blockmodel) {
&[6mm]
&[7mm] 
&[5mm]
& 
&
\\

& \node (theta) [latent] {$\theta$}; 
&
& \node (eta) [latent] {$\eta$}; 
& 
&
\\

&
&
&
& \node (false) [const,xshift=7mm] {false};
& 
\\

& \nofactor{drawtd}{}{Multi}{left=0pt}; 
& 
& 
& 
&
\\

& \node (td) [latent] {$T$};
& \nofactor{drawtu}{}{Multi}{below=0pt}; \node (tu) [latent,right=10pt of drawtu] {$T_{n_1}$};
& 
& 
& 
\\

\node (phi) [latent]  {$\phi$};
& \nofactor{drawxd}{}{Multi}{above=0pt}; \gate{gatexd}{(drawxd) (captdrawxd)}{$t$};
& \nofactor{drawtf}{}{Multi}{below=0pt}; 
\node (tf) [latent, right=10pt of drawtf] {$T_{n_2}$}; 
&  \nofactor{drawedge}{}{Bern}{right=0pt}; 
\node (drawedgeborder)[fit=(drawedge),outer sep=1pt]{};
\gate{gateedge}{(drawedgeborder)(drawedge) (captdrawedge)}{$t_{n_1}\rightarrow t_{n_2}$};
& \nofactor{drawfalse}{xshift=7mm}{=}{right=0pt};
& 
\\

& \node (xd) [obs] {$X$};
&
&
& \node (edgeobs) [transparent]{};
&
\\
};

\node (gateedgeborder) [fit=(gateedge),outer sep=2pt]{};

\vertgateSpec{gateback}{(drawedge)(gateedge)(gateedgeborder)(captdrawedge)}{}{(drawfalse)(captdrawfalse)}{};
\node [below left=0pt of gateback.70]{\footnotesize{$\textrm{false}$}};
\node [below right=0pt of gateback.70]{\footnotesize{$\textrm{true}$}};

\node (e) [latent, above=7pt of gateback.70] {$R$};
\nofactor{drawe}{above=10pt of e}{Bern}{left=0pt};
\node (edge) [obs, ellipse,below=4mm of gateback.south]  {$[(n_1,n_2)\in \edges]$};

\node (rho) [latent, above=35mm of gateback.70] {$\rho$};

\nofactor{draweta}{left=10pt of eta}{Beta}{below=0pt};
\nofactor{drawtheta}{above=10pt of theta}{Dir}{right=0pt};
\nofactor{drawphi}{above=10pt of phi}{Dir}{right=0pt};
\nofactor{drawrho}{above=10pt of rho}{Beta}{right=0pt};

\begin{scope}[node distance=17pt and 8pt]
  \node (alphaeta) [const, above left= of draweta] {$a_{\eta}$}; 
  \node (b-eta) [const, above right= of draweta] {$b_{\eta}$}; 
\end{scope}
\begin{scope}[node distance=10pt and 5pt]
   \node (alpharho) [const, above left= of drawrho] {$a_{\rho}$}; 
  \node (b-rho) [const, above right= of drawrho] {$b_{\rho}$}; 
\end{scope}
\node (alphatheta) [const, above=10pt of drawtheta]  {$\alpha_{\theta}$};
\node (alphaphi) [const, above=10pt of drawphi]  {$\alpha_{\phi}$};
\draw [->] (alphaphi) -- (drawphi) -- (phi);
\draw [->] (alphatheta) -- (drawtheta) -- (theta);
\draw [->] (alphaeta) -- (draweta) -- (eta);
\draw [-] (b-eta) -- (draweta);
\draw [->] (alpharho) -- (drawrho) -- (rho);
\draw [-] (b-rho) -- (drawrho);

\draw [->] (phi) -- (drawxd) -- (xd);
\draw [->] (theta) -- (drawtd) -- (td);
\draw [->] (theta) -- (drawtu) node [above,pos=0.55,sloped] {\footnotesize{$n=n_1$}} -- (tu);
\draw [->] (theta) -- (drawtf) node [below,pos=0.70,sloped] {\footnotesize{$n=n_2$}} -- (tf) ;
\draw [->] (eta) -- (drawedge) -- (edge);
\draw [->] (false) -- (drawfalse) -- (edge);
\draw [->] (rho) -- (drawe) -- (e);

\draw [-*] (td) -- (gatexd);
\draw [-*] (tu) -- (gateedge.west);
\draw [-*] (tf) -- (gateedge.west);
\draw [-*] (e) -- (gateback.70);

\plate{contentplate}{(td)(xd)(drawtd)(drawxd)(captdrawtd)(captdrawxd)(gatexd)}{$\forall x \in \content (n)$}{}
\plate{nodeplate}{(contentplate)(theta)(drawtheta)(captdrawtheta)}{$\forall n \in \nodes$}{}

\plate{edgeplate}{(edge)(gateback)(tu)(drawtu)(captdrawtu)(tf)(drawtf)(captdrawtf)(e)(drawe)}{$\forall (n_1,n_2) \in \nodes \times \nodes$}{}
\shiftedplate{matrixplate}{(eta)(draweta)(captdraweta)}{$\tastes \times \tastes$}{}{below right=0 pt  of invismatrixplate.south west }
\plate{topicplate}{(phi)(drawphi)(captdrawphi)}{$\forall t \in \tastes$}{}

\end{tikzpicture}
\par\end{centering}

\selectlanguage{american}%
\caption{Pairwise Link LDA model as directed factor graph.}

\end{figure}

\newpage{}
\end{document}

### 2013-01-31 latex/submission.tex
model in paper is no longer a lie
graphical model set

@@
 \usepackage{amsmath}
+\usepackage{tikz}
+\usepackage{tkz-base}
+\usetikzlibrary{shapes,decorations,shadows}
+\usetikzlibrary{decorations.pathmorphing}
+\usetikzlibrary{decorations.shapes}
+\usetikzlibrary{fadings}
+\usetikzlibrary{patterns}
+\usetikzlibrary{calc}
+\usetikzlibrary{decorations.text}
+\usetikzlibrary{decorations.footprints}
+\usetikzlibrary{decorations.fractals}
+\usetikzlibrary{shapes.gates.logic.IEC}
+\usetikzlibrary{shapes.gates.logic.US}
+\usetikzlibrary{fit,chains}
+\usetikzlibrary{positioning}
+\usepgflibrary{shapes}
+\usetikzlibrary{scopes}
+\usetikzlibrary{arrows}
+\usetikzlibrary{backgrounds}
+\tikzset{latent/.style={circle,fill=white,draw=black,inner sep=1pt, 
+minimum size=20pt, font=\fontsize{10}{10}\selectfont},
+obs/.style={latent,fill=gray!33},
+const/.style={rectangle, inner sep=1pt},
+factor/.style={rectangle, fill=black,minimum size=5pt, inner sep=1pt},
+>={triangle 45}}
+
+\pgfdeclarelayer{b}
+\pgfdeclarelayer{f}
+\pgfsetlayers{b,main,f}
+
+\newcommand{\plate}[4]{
+\begin{pgfonlayer}{b}
+\node (invis#1) [draw, color=white, inner sep=6pt,rectangle,fit=#2] {};
+\end{pgfonlayer}\begin{pgfonlayer}{f}
+\node (capt#1) [ below left=0 pt of invis#1.south east, xshift=1pt,yshift=1pt] {\footnotesize{#3}};
+\node (#1) [draw,inner sep=1pt, rectangle,fit=(invis#1) (capt#1),#4] {};
+\end{pgfonlayer}
+}
+
+\newcommand{\shiftedplate}[5]{
+\begin{pgfonlayer}{b}
+\node (invis#1) [draw, color=white, inner sep=0 pt,rectangle,fit=#2] {};
+\end{pgfonlayer}\begin{pgfonlayer}{f}
+\node (capt#1) [#5, xshift=2pt] {\footnotesize{#3}};
+\node (#1) [draw,inner sep=2pt, rectangle,fit=(invis#1) (capt#1),#4] {};
+\end{pgfonlayer}
+}
 
@@
 
-\section{Model and Methods}
+\section{Model and Features}
 
@@
 
-We present our model in section~ref{sec:model}.
-The rest of the subsections explain how we overcome these difficulties.
+\begin{figure}
+\centering
+
+\begin{tikzpicture}
+\matrix[row sep=6mm, column sep=8mm, matrix anchor=mid] (mdp) {
+  \node (alpha) [const] {$\alpha$}; & & \\
+  \node (pi) [latent] {$\+\pi$}; & \node (beta) [latent] {$\beta_k$}; & \node (alphabeta) [const] {$\alpha_0, \beta_0$}; \\
+  \node (z) [latent] {$z_n$}; & \node(Lambda) [latent] {$\Lambda_k$}; & \node(Lambdanu) [const] {$\Lambda_0, \nu$}; \\
+  \node (x) [obs] {$\+x_n$}; & \node(mu) [latent] {$\+\mu_k$}; & \node(mukappa) [const] {$\+\mu_0, \kappa$}; \\
+};
+
+\draw [->] (alpha) -- (pi);
+\draw [->] (pi) -- (z);
+\draw [->] (z) -- (x);
+\draw [->] (alphabeta) -- (beta);
+\draw [->] (Lambdanu) -- (Lambda);
+\draw [->] (mukappa) -- (mu);
+\draw [->] (Lambda) -- (mu);
+\draw [->] (beta) -- (x);
+\draw [->] (Lambda) -- (x);
+\draw [->] (mu) -- (x);
+
+\plate{Nplate}{(z)(x)}{N}{}
+\plate{Kplate}{(beta)(Lambda)(mu)}{K}{}
+
+\end{tikzpicture}
+\caption{The graphical model} \label{fig:model}
+\end{figure}
+
+\label{sec:model}
+We present our model in Figure~\ref{fig:model}.
+Below, we explain our feature representation and how we overcome the aforementioned difficulties.
 First, the number of cells must be inferred from the image itself.
@@
 
-\subsection{Model}
-\ref{sec:model}
-In summary, pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances  to every other pixel.
-The first four components, $\+{x}_{n,1:4}$ obey a multivariate normal distribution while the  Putting these components together yields the model in Figure~\ref{fig:model}.
+Pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances to every other pixel.
+If position $\+p \in \+R^2$, when discretized, is the location of pixel $n$ and the discrete $\+q$ is the location of pixel $m$, then we write $d(\+p, \+q) = \+d_n,m$.
+Putting these components together yields the model in Figure~\ref{fig:model}.
 The generative story is:
+
 \begin{itemize}
@@
   \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
-  \item $\beta_k \sim \mbox{Gamma}(\beta_k | \alpha_0, \beta_0)$
+  \item $\beta_k | \alpha_0, \beta_0 \sim \mbox{Gamma}(\beta_k | \alpha_0, \mbox{rate} = \beta_0)$
   \end{itemize}
@@
 Therefore, shape constraints based on Euclidean distances are not justified.
-Instead, we define a \emph{geodesic distance}.
+Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." We quantify this property with a \emph{geodesic distance}.
 Construct a graph whose vertices are the pixels and undirected edge for adjacent pixels ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$).
@@
 
-We model a two-dimensional image as an infinite semi-Gaussian mixture model.
-This extends the classic Gaussian mixture model (GMM) with (a) a variable number of classes and (b) additional naive-Bayes features associated with each class.
-
-\begin{enumerate}
-\item $\phi_{1\cdot} \in [0,1]$: pixel intensity 
-\item $\phi_{2\cdot} \in [0,1]$: average interior Radon-like feature response
-\item $\phi_{3\cdot} \in [0,1]$: geodesic distance  
-\item 
-\end{enumerate}
- Because cell interiors contain diverse structures, intensities values within a cell vary widely, so relying on $\phi_{1\cdot}$ alone oversegments.
-To reduce oversegmentation, we compute a cel boundary Radon-like feature for $\phi_{1\cdot}$, which homogenizes pixel intensities within a given cell~\cite{Kumar2010}.
-One computes Radon-like functions by taking line integrals of a kernel function along various scanning orientations, broken piecewise at image discontinuities.
-By varying the kernel, different structures and boundaries are enhanced.
-We follow Kumar's approach of averaging the integrals along each scan orientation to obtain a scalar response at each pixel.
-Finally, recall that our background class ($k = 1$) contains all pixels belonging on the cell membrane or in intracellular space.
-We penalize points
-
-%, the cell Background Radon-like feature response in $\phi_{2\cdot}$, and the length of the shortest cluster path from $(
+\section{Inference}
+Our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings.
 
 \section{Results}
+Forthcoming.
 

### 2013-01-31 latex/submission.tex
most substance is done

@@
-\documentclass[draft,english]{article}
+\documentclass[english]{article}
 \usepackage{fullpage}
@@
 \usepackage{tkz-base}
+\usepackage{graphicx}
+\usepackage{algorithm}
+\usepackage{algorithmicx}
+\usepackage{algpseudocode}
+\graphicspath{{../fig/}}
+
 \usetikzlibrary{shapes,decorations,shadows}
@@
 
-\title{A Bayesian Nonparametric Approach to Neural Image Segmentation}
+\title{Bayesian Nonparametric Neural Image Segmentation}
 \author{Kui Tang and Frank Wood}
@@
 Our model imposes spatial locality on arbitrary features, automatically determines the number of cells in an image, and can be used in both unsupervised and semi-supervised settings.
-We model the distribution of neurons in an image as an infinite Gaussian mixture model.
-Each neuron's pixels are characterized by likelihood function that is a combination of multivariate Gaussian and naive Bayes features.
+We introduce a novel geodesic distance which adaptively adjust segmentation to respect detected cell boundaries.
 Using data collected by Kasthuri et al., we show that our model can obtain plausible segmentations with no training data, while receiving a significant boost in its presence.
+Finally, the Bayesian nonparametric model offers a natural extension to 3 dimensions, where neural columns of segments \emph{through} slices are modelled as coagulation-fragmentation processes, paving the road for a fully Bayesian solution to the connectomics problem.
+\end{abstract}
 
@@
 The scale of the data is profound; the human brain is estimated to contain 100 billion neurons connected by up to a quadrillion synapses~\cite{Kasthuri2010}.
-Like the human genome project before, it is clear that automated methods are needed to map the connectome.
 
@@
 elegans} in 1986, marking axons, dendrites, and cell bodies in each frame and tracing the labels through frames~\cite{White12111986}.
-More recent work uses computer assistance but fundamentally rely on a skilled human's identification of neural bodies~\cite{Jarrell2012,Bock2011Roberts2011}.
-Several authors have developed efficient semi-automated methods \cite{Roberts2011,Unger2009}.
-Unfortunately, current baseline speeds of computer-assisted manual tracing would still take 10,000 person-years to construct a mouse cortical column~\cite{Briggman2006}.
-Therefore, henceforth we focus on automated techniques.
+Recent work in semi-automated methods are accurate and promising \cite{Roberts2011,Unger2009,Jarrell2012,Bock2011}, but current baseline speeds of computer-assisted manual tracing would still take 10,000 person-years to construct a mouse cortical column~\cite{Briggman2006}, hence the need for fully automatic methods.
+
+EM images present several challenges for classical segmentation algorithm.
+First, structures or interest are small and densely packed \cite{Jain2007}.
+Second, due to the physical cutting process, images may be warped from slice to slice \cite{Koshevoy2006}.
 
 \section{Related Work}
-\fxfatal{Write topic sentence, summarize your current literature review, and add paragraphs about feature engineering and graph pathfinding methods. Cite Malik's normalized cuts.}
-Automated techniques generally begin by segmenting a 2D slice based on cell membrane boundaries---a highly distinctive feature.
+Automated techniques generally begin by finding the cell boundaries---a highly distinctive feature.
+Because cells are usually disconnected, segmentation can be reduced to binary classification of whether voxels are inside or outside a cell.
+Segmentation can be done directlyin voxel space.
+Jain et al.
+\cite{Jain2007} achieved an 8.0\% test error with a 34,000 convolutional neural network, a statistically significant improvement other the 18.95\% error of thresholding, while MRFs and CRFs do no better than thresholding.
+Andres et al.
+obtain similar performance, at 8\% misclassification, using the watershed algorithm over a probability map of a random forest with hand-engineered features \cite{Andres2008}.
+However, the dataset used by both used a special intracellular stain which made segmentation easier.
+The generality of these solutions are thus unclear.
+
+More common is to segment in pixel space and track the identity of segmented bodies in the vertical dimension.
+This mitigates the poor vertical resolution of EM images \cite{Kaynig2010, Briggman2006}.
+Jurrus et al.
+\cite{Jurrus2008} first obtain 2D segmentations with watershed after preprocessing linear diffusion filtering (NB: learn what this is!), and construct a weighted graph of these segments, with edges between adjacent sections with edge costs favoring high cross-correlation of segments low lateral displacement.
+Each neural process is thus a shortest path through this graph, computed by Dijkstra's algorithm.
+This cost function is robust to merging and loss of sections, but because Dijkstra's algorithm finds a single path, it cannot model branching.
+Kaynig et al.
+\cite{Kaynig2010a} presuppose segmentation and represent the image as a graph of regions, like Jurrus.
+Pairwise features are represented as edge weights, and a random forest classifier is trained to output a scalar similarity score for each pair.
+Then a probabilistic model encoding geometric constraints with cross-section correspondences as hidden variables is trained with EM.
+Finally, the maximum-likelihood result is the 3D reconstruction.
+These pairwise correspondences are fed to an agglomerative clustering algorithm, constructing the neural processes from ground-up in the same manner as a human.
+Unlike in Jurrus, processes may merge.
+However, each region is explicitly limited to two correspondences: one above and below, so processes still cannot branch.
+
+Feature engineering significantly shapes the state of the art.
+Venkataraju et al.
+\cite{Venkataraju2009} generated 100 Hessian neighborhood features at each pixel and classified Jurrus's dataset with AdaBoost.
+Lucchi et al.
+\cite{Lucchi2010} develop ray descriptors, which capture shape features without a predefined shape model.
+These features are rotation, scale, and translation invariant, but strongly discriminate the types of objects observed in EM images.
+There features were used in SVMs to classify interior and boundary points of mitochondria, which are considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy.
+Kaynig et al.
+encode perceptual grouping constraints---rules of thumb used by human experts to group neuron structures---as pairwise interaction terms in an energy function solved by graph cut.
+Minimizing the energy function gives a segmentation, and identifying maximum overlapping regions between slices is sufficient to obtain a 3D reconstruction
+
+\fxwarning{Who cites Malik's normalized cuts?}
 
@@
 
-\fxfatal{What microscopy method did they use? and put in the right place.}
-We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \mbox{ nm}^3$ collected by \fxfatal{Unpublished; to cite}{Kasthuri et al.}
+\fxwarning{What microscopy method did they use? and put in the right place.}
+We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \text{ nm}^3$ collected by \fxwarning{Unpublished; to cite}{Kasthuri et al.}
 
@@
   \node (alpha) [const] {$\alpha$}; & & \\
-  \node (pi) [latent] {$\+\pi$}; & \node (beta) [latent] {$\beta_k$}; & \node (alphabeta) [const] {$\alpha_0, \beta_0$}; \\
+  \node (pi) [latent] {$\+\pi$}; & \node (beta) [latent] {$\gamma_k$}; & \node (alphabeta) [const] {$\beta_0, \gamma_0$}; \\
   \node (z) [latent] {$z_n$}; & \node(Lambda) [latent] {$\Lambda_k$}; & \node(Lambdanu) [const] {$\Lambda_0, \nu$}; \\
@@
 \end{tikzpicture}
-\caption{The graphical model} \label{fig:model}
+\caption{The graphical model.
+Unlike a standard Gaussian mixture model, our model couples $\gamma_k$ and $\+\mu_k$ through the $\+x_n$, giving the flexibility to model boundary constraints, but also presents challenges in inference.
+See text for details.}\label{fig:model}
 \end{figure}
@@
 First, the number of cells must be inferred from the image itself.
-To do so, we employ a Mixture Dirichlet process, discussed in section~\ref{sec:mdp}.
+To do so, we employ a mixture Dirichlet process, discussed in section~\ref{sec:mdp}.
 Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3.
-We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}, and constrain cluster assignments with a geodesic distance that penalizes boundary crossings in section~\ref{sec:geodesic}.
+We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}.
+Finally, our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings, discussed in section~\ref{sec:geodesic}.
 
 Pixel $i$ has feature vector $\+{x}_n = (x_n, y_n, r_n, f_n, \+{d}_n)$ where $x_n, y_n$ represent the pixel's spatial coordinates, $r_n \in [0,1]$ is the raw pixel intensity, $f_n \in [0,1]$ is the average response of the interior Radon-like feature, and $\+{d}_n \in \+{R}^N$ is a vector of geodesic distances to every other pixel.
-If position $\+p \in \+R^2$, when discretized, is the location of pixel $n$ and the discrete $\+q$ is the location of pixel $m$, then we write $d(\+p, \+q) = \+d_n,m$.
+If position $\+p \in \+R^2$, when discretized, is the location of pixel $n$ and the discrete $\+q$ is the location of pixel $m$, then we write $d(\+p, \+q) = \+d_{n,m}$.
 Putting these components together yields the model in Figure~\ref{fig:model}.
@@
 \begin{itemize}
-\item $\+\pi \sim \mbox{DirichletProcess}(\alpha)$
-\item $k | \+\pi \sim \mbox{Discrete}(\+\pi)$
+\item $\+\pi \sim \text{DirichletProcess}(\alpha)$
+\item $k | \+\pi \sim \text{Discrete}(\+\pi)$
 \item If $k$ is a new (empty) cluster:
   \begin{itemize}
-  \item $\Lambda_k | T, \nu \sim \mbox{Wishart}(\Lambda | T, \nu)$ (in the notation of \cite{Murphy2003})
+  \item $\Lambda_k | \Lambda_0, \nu \sim \text{Wishart}(\Lambda | \Lambda_0, \nu)$ (in the notation of \cite{Murphy2003})
   \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
-  \item $\beta_k | \alpha_0, \beta_0 \sim \mbox{Gamma}(\beta_k | \alpha_0, \mbox{rate} = \beta_0)$
+  \item $\gamma_k | \beta_0, \gamma_0 \sim \text{Gamma}(\gamma_k | \beta_0, \text{rate} = \gamma_0)$
   \end{itemize}
-\item $\+x | \+\mu_k , \Lambda_k, \beta_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \mbox{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \mbox{rate} = \beta_k \right)$ (in the notation of section~\ref{sec:geodesic})
+\item $\+x | \+\mu_k , \Lambda_k, \gamma_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \text{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \text{rate} = \gamma_k \right) \label{eq:xlike}$ (in the notation of section~\ref{sec:geodesic})
 \end{itemize}
@@
 Let $\+{z}_{-n}$ denote the vector of all indicators excluding element $n$ and $N_{k,-n}$ denote the number of points assigned to class $k$ excluding point $n$.
-A standard result \fxfatal{CITATION} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }$$ which makes Gibbs sampling easy.
+A standard result \cite{Rasmussen2000} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }$$ which makes Gibbs sampling easy.
 
@@
 \frac{N_{k,-n}}{N-1+\alpha} & N_{k,-n}>0\\
-\frac{\alpha}{N-1+\alpha} & \mbox{otherwise}
-\end{cases}
+\frac{\alpha}{N-1+\alpha} & \text{otherwise}
+\end{cases} \label{eq:mdp}
 \]
-satisfies the limiting conditions and correctly normalizes to 1 when summed with $k = 1, \ldots, \infty$.
+satisfies the limiting conditions and correctly normalizes to 1 when summed over $k = 1, \ldots, \infty$.
 We now say that the mixing distribution $\+\pi$, which is now infinite-dimensional, is drawn from a \emph{Dirichlet process}.
@@
 \label{sec:radon}
+
+\begin{figure}
+\centering
+\includegraphics[width=0.32\linewidth, clip]{JAN_im}
+\includegraphics[width=0.32\linewidth, clip]{JAN_radon_bg}
+\includegraphics[width=0.32\linewidth, clip]{JAN_radon_bd}
+\caption{A $600\times600$ image slice, its background average Radon response, and its boundary average Radon response.
+The background response homogenizes pixel intensities within a cell body, while the boundary response highlights cell boundaries.} \label{fig:radon}
+\end{figure}
+
 Ideally, for a mixture model, pixel intensities within a cell should be homogeneous.
 But the diversity of intracellular structures and significant imaging noise yield data far from this ideal.
-As a countermeasure, we augment raw pixel intensities with $\emph{Radon-like feature}$ responses, which have proven useful for connectomics segmentation \cite{Kumar2010}.
-\fxfatal{Make this clearer...
+To correct this, we augment raw pixel intensities with $\emph{Radon-like feature}$ responses, which have proven useful for connectomics segmentation \cite{Kumar2010}.
+\fxwarning{Make this clearer...
 Even the original paper doesn't have a concise description}.
@@
 Suppose a line $\+{l}$ is given and $(x, y) = \+{l}(t)$ for some $t \in [t_i, t_{i+1}]$.
-Then the Radon-like response for image $I$ at $(x, y)$ in direction $\+{l}$ is given by $$\Psi \left( I,\+{l},t \right) = T \left( I, \+{l}(t) \right).$$ We will use the extraction function for cell background segmentation $$T_{\mbox{bg}}(I, \+{l}(t)) = \min_{t \in [t_i, t_{i+1}]} \left\{ I \left( \+{l}(t) \right) \right\}$$ which simply returns the dimmest pixel between any two knots \cite{Kumar2010}.
+Then the Radon-like response for image $I$ at $(x, y)$ in direction $\+{l}$ is given by $$\Psi \left( I,\+{l},t \right) = T \left( I, \+{l}(t) \right).$$ We will use the extraction function proposed in \cite{Kumar2010} for cell background segmentation $$T_{\text{bg}}(I, \+{l}(t)) = \min_{t \in [t_i, t_{i+1}]} \left\{ I \left( \+{l}(t) \right) \right\}$$ which simply returns the dimmest pixel between any two knots.
 Because knots denote edges, this feature tends to homogenize pixel intensities within a cell.
@@
 Finally, we want pixels to not cross cell boundaries.
-The first step in encoding this constraint is to use a boundary-enhancing extraction function \cite{Kumar2010} $$T_{\mbox{bd}} = \frac{\int_t^{t+1} R\left( \+{l}(\tau) \right) d\tau }{\left\Vert \+{l}(t_{i+1}) - \+{l}(t_{i}) \right\Vert_2 }.$$ This function assigns to each pixel the average value of the Gaussian second-derivative response.
+The first step in encoding this constraint is to use a boundary-enhancing extraction function \cite{Kumar2010} $$T_{\text{bd}} = \frac{\int_t^{t+1} R\left( \+{l}(\tau) \right) d\tau }{\left\Vert \+{l}(t_{i+1}) - \+{l}(t_{i}) \right\Vert_2 }.$$ This function assigns to each pixel the average value of the Gaussian second-derivative response.
 Recall that the GSD tends to enhance boundaries, the most salient of which are captured by the Canny detector as knots.
-Thus, $T_{\mbox{bd}}$ smoothes the spurious GSD responses, avoiding oversegmentation.
-\fxfatal{Tighten up wording here}.
-Further, in connectomics data, cell membranes tend to form thick boundaries.
+Thus, $T_{\text{bd}}$ smoothes the spurious GSD responses, avoiding oversegmentation.
+\fxwarning{Tighten up wording here.}Further, in connectomics data, cell membranes tend to form thick boundaries.
 But classic edge detectors work best at single-pixel gradients and return two edges for a thick boundary.
-\fxfatal{cite Rafael}.
 In particular, the Canny detector produces two knots on either side of a cell boundary.
-In the GSD response $R$, cell boundaries are consistently bright, so the integral in $T_{\mbox{bd}}$ returns a distinctly brighter response on boundaries, regardless of thickness.
+In the GSD response $R$, cell boundaries are consistently bright, so the integral in $T_{\text{bd}}$ returns a distinctly brighter response on boundaries, regardless of thickness.
 We use this boundary enhancement information in the next section.
@@
 \label{sec:geodesic}
-Figure~\ref{fig:radon_bd} illustrates that cells need not be elliptical, or even convex.
+\begin{figure}
+\includegraphics[width=0.24\linewidth, clip]{JAN_bd_small}
+\includegraphics[width=0.24\linewidth, clip]{JAN_bd_dist}
+\includegraphics[width=0.24\linewidth, clip]{JAN_bd_thresh}
+\includegraphics[width=0.24\linewidth, clip]{JAN_bd_hist}
+\caption{Boundary Radon-like feature response; red marks the center from which the remaining pictures are measured.
+Grayscale map of geodesic distance (of superpixels).
+Threshold of geodesic distance $< 0.1$.
+Distribution of geodesic distances about the red center.
+Although the separation between interior and exterior is easy effort, BASELINE BLAH}\label{fig:geodesic}
+\end{figure}
+Figures~\ref{fig:radon} and~\ref{fig:geodesic} illustrates that cells need not be elliptical, or even convex.
 All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response.
 Therefore, shape constraints based on Euclidean distances are not justified.
-Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." We quantify this property with a \emph{geodesic distance}.
-Construct a graph whose vertices are the pixels and undirected edge for adjacent pixels ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$).
+Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." Let us quantify this statement.
+Construct an undirected graph with pixels as vertices and pairs of adjacent pixels as edges ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$).
 Each edge has a weight $w(i,j) = \left| b_i - b_j \right|$ where $b_k$ is the boundary Radon-like response at pixel $k$.
-We define $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm.
-In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances.Address practical concerns.
+We define the \emph{geodesic distance} $d(\+{x}_i, \+{x}_j)$ to be the total weight of the shortest path between pixels $i$ and $j$, which is computed with Johnson's algorithm.
+In practice, the $O(N^2)$ storage is prohibitive, so we instead we downsample the image into SLIC superpixels \cite{Achanta2012} and compute superpixel-based distances.
 Further, because individual cells are tiny relative to an entire slice, only local distances are needed, because far points must be outside the current cell.
@@
 \section{Inference}
-Our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings.
+Inference is performed by Gibbs sampling, using Radford Neal's Algorithm 8 for non-conjugate priors \cite{Neal2000}.
+This is because the likelihood of $\+x$, which depends on the function $d$, is not a known conjugate-parametric form.\fxwarning{I'm not sure this is the right explanation.} We collapse $\+\pi$, but we do not collapse the other latent variables.
+Denote by $X$ the set of all $\+x_n$ and by $X_k = \left\{ \+x_n \in X : z_n = k \right\}$ and $N_k = | X_k |$.
+Then we derive the conditional posteriors by finding the Markov blankets of each node and using Bayes' rule.
+But because the likelihood of $\+x_n$ is note a standard parametric form, conditionals involving $X)k$ and $\+\mu_k$ cannot be sampled explicitly; we can only compute their likelihoods can be computed up to proportionality and resort to Metropolis-Hastings sampling.
+\begin{align}\label{eq:conds}
+p(z_n = k | \+\pi, \+x_n, \+\mu_k, \Lambda_k) &= p(z_n = k | \+\pi) \mathcal{N}(\+x_n | \+\mu_k, \Lambda_k) \label{eq:z}\\
+p(\gamma_k | X_{z=k} , \beta_0, \gamma_0) &= \text{Gamma}\left( \gamma_k | \beta_0 + N_k, \gamma_0 + \sum_{\+x \in X_k} \+x \right) \label{eq:gamma}\\
+p(\Lambda_k | X_k, \+\mu_k, \gamma_k, \Lambda_0, \nu) &\propto  p(X_k | \+\mu_k, \Lambda_k, \beta_k) \text{Wishart}(\Lambda_k | \Lambda_0, \nu) \label{eq:lambda}\\
+p(\+\mu_k | X_k, \Lambda_k, \gamma_k, \+\mu_0, \kappa) &\propto p(X_k | \+\mu_k, \Lambda_k, \gamma_k) \mathcal{N}(\+\mu_k | \+\mu_0, (\kappa\Lambda)^{-1}) \label{eq:mu}
+\end{align}
+Note that $\+\pi$ remains collapsed; to evaluate \eqref{eq:z}, see \eqref{eq:mdp}.
+
+\begin{algorithm}
+\caption{Gibbs sample iteration}\label{alg:gibbs}
+\begin{algorithmic}
+\Require The current state of the Markov chain, $\Theta = \left\{ N, K, Z, \left\{ N_k, \+\mu_k, \Lambda_k, \gamma_k \right\}_{k=1}^K  \right\}$
+\Ensure The distribution of the next state $\Theta^{*}$ preserves the Markov chain's invariant distribution.
+\Statex
+
+\For{$n := 1:N$}
+  \If{$N_{z_n} = 1$}
+    \State $\left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right) := \left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right)$ 
+    \State Draw $\+\mu_j, \Lambda_j, \gamma_j$ from their priors (see section~\ref{sec:model}), for $j \in \{K+2, \ldots, K+m\}$
+  \Else
+    \State Draw $\+\mu_j, \Lambda_j, \gamma_j$ from their priors (see section~\ref{sec:model}), for $j \in \{K+1, \ldots, K+m\}$
+  \EndIf
+  \State Draw $z_n$ from 
+  \[
+  p(z_{n}=k|\+{z}_{-n},\+x_n,\+\mu_k,\Lambda_k,\gamma_k,\alpha)=\begin{cases}
+  \frac{N_{k,-n}}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \beta_k) & k < K\\
+  \frac{\alpha / m}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \beta_k) & \text{otherwise}
+  \end{cases} \label{eq:nealalg8}
+  \]
+  \State Delete parameters for classes $k$ with $N_k = 0$ and relabel the $z_n$ to remove gaps.
+\EndFor
+
+\For{$k := 1:K$}
+  \State Draw $\gamma_k$ using \eqref{eq:gamma}
+  \State Metropolis-Hastings sample $\Lambda_k$ using likelihood \eqref{eq:lambda} and proposal $\text{Wishart}(\Lambda^{*} | p \Lambda, \nu)$ for some $p > 1$.
+  \State Metropolis-Hastings sample $\+\mu_k$ using likelihood \eqref{eq:mu} and proposal $\mathcal{N}( \+\mu^{*} | C )$.
+\EndFor
+
+\end{algorithmic}
+\end{algorithm}
+
+\subsection{Proposal Distributions}
+Care is required to select proposal distributions for \eqref{eq:mu}, because the likelihood for $\+x_n$ is discontinuous in $\+\mu_k$, since it relies on the discrete function $d$ which arises from graph distances.
+The lack of reliable gradient information frustrates the smooth exploration of low-probability areas.
+However, discontinuities only occur at superpixel boundaries, and sharp discontinuities only at the transition between cell boundary and interior.
+Therefore, we use a diagonal-covariance Gaussian proposal whose $xy$ coordinate marginal forms a 95\% confidence ellipse that covers the neighboring superpixels.
+SLIC superpixels have roughly uniform size, even if shapes differ.
+Thus, a circle can be guaranteed to cover some neighbors of a superpixel.
+Thus, we propose with the following covariance
+\[
+C = \left(\begin{array}{cccc}
+    \sigma^{2}\\
+     & \sigma^{2}\\
+      &  & \tau^{2}\\
+       &  &  & \tau^{2}
+     \end{array}\right)
+\]
+where $\sigma$ is the diameter of a SLIC superpixel and $\tau = 1/2$, because the third and fourth features are constrained to lie in $[0, 1]$.
+
+To sample $\Lambda_k$, since $\+\mu_k$ is constant, equation \eqref{eq:lambda} is continuous in $\Lambda_k$, so a Wishart proposal suffices.
+However, to ensure exploration of the state space, we adjust the variance of the Wishart proposal to be higher than in the prior in \ref{sec:model}.
 
@@
 \section{Discussion}
-
-\end{abstract}
-\section*{October 3, 2012}
-To appear in: Related Work
-
-Koshevoy et al.
-\cite{Koshevoy2006} use a variant of SIFT \cite{Lowe2004} to register a stack of images into a 3D volume.
-According to Koshevoy, images between images exhibit warping due to the physical slicing, physical structural changes, and a distinct rotation and displacement on each frame.
-From the displacements of individual keypoints, a global transformation vector is calculated, allowing slices to be properly aligned.
-However, the SIFT features do not capture neural processes well, and thus is unsuited for identifying neurons across images.
-
-\section*{October 15, 2012}
-Introduction:
-
-Related Work:
-
-EM images present several challenges for classical segmentation algorithm.
-First, structures or interest are small and densely packed \cite{Jain2007}.
-Second, due to the physical cutting process, images may be warped from slice to slice \cite{Koshevoy2006}.
-
-Fully-automated methods use manual labels as input for supervised learning.
-Segmentation can be done in voxel space, after which the learned segmentation can be directly used as a 3D reconstruction.
-Because cells are usually disconnected, segmentation can be reduced to binary classification of whether voxels are inside or outside a cell.
-Jain et al.
-\cite{Jain2007} achieved an 8.0\% test error with a 34,000 convolutional neural network, a statistically significant improvement other the 18.95\% error of thresholding, while MRFs and CRFs do no better than thresholding.
-Andres et al.
-obtain similar performance, at 8\% misclassification, using the watershed algorithm over a probability map of a random forest with hand-engineered features \cite{Andres2008, Gonzales2008}.
-However, the dataset used by both used a special intracellular stain, making the segmentation job easier.
-
-Alternatively, one can segment in pixel space and track the identity of segmented bodies in the vertical dimension.
-This is motivated by the poor vertical resolution of EM images: the thinnest slices are 40nm think \cite{Kaynig2010b, Briggman2006}.
-Jurrus et al.
-\cite{Jurrus2008} first obtain 2D segmentations with watershed after preprocessing linear diffusion filtering (NB: learn what this is!), and construct a weighted graph of these segments, with edges between adjacent sections with edge costs favoring high cross-correlation of segments low lateral displacement.
-Each neural process is thus a shortest path through this graph, computed by Dijkstra's algorithm.
-This cost function is robust to merging and loss of sections, but because Dijkstra's algorithm finds a single path, it cannot model branching.
-Kaynig et al.
-\cite{Kaynig2010a} presuppose segmentation and represent the image as a graph of regions, like Jurrus.
-Pairwise features are represented as edge weights, and a random forest classifier is trained to output a scalar similarity score for each pair.
-Then a probabilistic model encoding geometric constraints with cross-section correspondences as hidden variables is trained with EM.
-Finally, the maximum-likelihood result is the 3D reconstruction.
-These pairwise correspondences are fed to an agglomerative clustering algorithm, constructing the neural processes from ground-up in the same manner as a human.
-Unlike in Jurrus, processes may merge.
-However, each region is explicitly limited to two correspondences: one above and below, so processes still cannot branch.
-
-Feature engineering has improved 2D segmentation.
-Venkataraju et al.
-\cite{Venkataraju2009} generated 100 Hessian neighborhood features at each pixel and classified Jurrus's dataset with AdaBoost.
-Lucchi et al.
-\cite{Lucchi2010} develop ray descriptors, which capture shape features without a predefined shape model.
-These features are rotation, scale, and translation invariant, but strongly discriminate the types of objects observed in EM images.
-There features were used in SVMs to classify interior and boundary points of mitochondria, which are considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy.
-Kaynig et al.
-encode perceptual grouping constraints---rules of thumb used by human experts to group neuron structures---as pairwise interaction terms in an energy function solved by graph cut.
-Minimizing the energy function gives a segmentation, and identifying maximum overlapping regions between slices is sufficient to obtain a 3D reconstruction
-
-\section{November 2, 2012}
-Initial Approaches
-
-We modelled a single slice of neural tissue with a finite Gaussian mixture model with full covariances.
-The feature vector $(x, y, i)$ consisted of the positional coordinates and pixel intensity.
-Each cluster represented the pixels belonging to a single cell, with a special background cluster to contain cell membrance and interstitial space.
-For this background, we experimented with both a triangular probability density $p(i) = 2 - 2*i$ and a sigmoid density of the form $p(i) = \frac{1}{Z} \left[ 1 - \left( 1 + \exp(-p(i-c)) \right)^-1 \right]$ with $p$ and $c$ denoting the precision (inverse scale) and location (the preimage of the midpoint of the range).
-Inference was performed with Gibbs sampling.
-
-Visually, our segmentation seems reasonable, although cells whose interior intensity varies significantly tend to be oversegmented, and the cell boundary model is insufficient because the boundary intensity varies locally (see slides).
-A more accurate background model must be adaptive.
-
-A more serious problem is that our Gaussian intensity model is insufficienltly compact: the standard deviations are far too high to accurately model the bounded $[0, 1]$ intensity range.
-To rectify this, we experimented with modelling the third dimension as a Beta distribution independently of the first two dimensions, but still dependent on class labels.
-We implemented a Metropolis-Hastings sampler using a lognormal proposal and a prior based on \cite{Bouguila2006}.
-We modified the prior to take the form $$ \left[ 1 - exp(-( (s - 0.5)^2)) (1 - exp(-( (a - 1)^2)))(1 - exp(-( (b - 1).^2))) (1 - exp(-d * ((s - 2).^2 + (m - 0.5).^2))) exp(-r / (s.^2 .* m .* (1 - m)) - (k * s.^2) / 2) \right]$$.
-This avoids distributions with $a, b < 1$, which lead to unbounded U-shaped distributions, as well as $a = b = 1$, the uniform distribution.
-Unfortunately, our experiment with this model reverted to axis-aligned decision boundaries.
-It appears that spatio-intensity correlations are essential.
+Forthcoming.
 

### 2013-01-31 src/make_diff_graph.m

@@ -9,7 +9,7 @@ function [ segments, slic_inds, edge_G ] = make_diff_graph( img, slic_npixels, s
 % edge_G   - sparse [Nslic Nslic] lower triangular graph. Two superpixels
 %            are adjacent if there exist two adjacent underlying pixels.
 
-    [X, Y] = size(img);
+    [R, C] = size(img);
     
     % vl_slic returns 0-indexed; we want 1-indexed.
     segments = vl_slic(single(img), slic_npixels, slic_reg) + 1;
@@ -43,7 +43,7 @@ function [ segments, slic_inds, edge_G ] = make_diff_graph( img, slic_npixels, s
             % Explore the neighborhood
             for dr = [-1 1]
                 for dc = [-1 1]
-                    if (r + dr > 1 && r + dr < Y) && (c + dc > 1 && c + dc < X)
+                    if (r + dr > 1 && r + dr < R) && (c + dc > 1 && c + dc < C)
                         other_n = segments(r + dr, c + dc);                    
 
                         % CHECK THIS LINE... earlier you asserted other_n >

### 2013-01-31 src/startup.m

@@ -14,7 +14,8 @@ S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
 %A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
 %f1 = S1(1000:1199,1000:1199);
 %f1 = S1(1200:1399,900:1099);
-f1 = S1(1600:1799,1:200);
+%f1 = S1(1400:1999,200:799);
+f1 = S1(1600:1899,1:300);
 %a1 = A1(1000:1199,1000:1199);
 [PHI, misc_data] = feature_map(f1);
 

### 2013-02-06 latex/submission.tex
revised paper

@@
 
-\listoffixmes
-
 \title{Bayesian Nonparametric Neural Image Segmentation}
@@
 We develop a Bayesian nonparametric method to segment neurons in electron micrograph images of brain tissue.
-Our model imposes spatial locality on arbitrary features, automatically determines the number of cells in an image, and can be used in both unsupervised and semi-supervised settings.
-We introduce a novel geodesic distance which adaptively adjust segmentation to respect detected cell boundaries.
+Our model imposes spatial locality on arbitrary features, automatically determines the number of segments in an image, and can be used in both unsupervised and semi-supervised settings.
+We introduce a novel geodesic distance feature which adaptively adjust segmentation to respect detected cell boundaries.
 Using data collected by Kasthuri et al., we show that our model can obtain plausible segmentations with no training data, while receiving a significant boost in its presence.
-Finally, the Bayesian nonparametric model offers a natural extension to 3 dimensions, where neural columns of segments \emph{through} slices are modelled as coagulation-fragmentation processes, paving the road for a fully Bayesian solution to the connectomics problem.
+Finally, our Bayesian nonparametric naturally extends to 3 dimensions, where neural columns of segments \emph{through} slices are modelled.
 \end{abstract}
@@
 \section{Related Work}
-Automated techniques generally begin by finding the cell boundaries---a highly distinctive feature.
+Automated techniques generally try to find cell boundaries---a highly distinctive feature.
 Because cells are usually disconnected, segmentation can be reduced to binary classification of whether voxels are inside or outside a cell.
@@
 Jain et al.
-\cite{Jain2007} achieved an 8.0\% test error with a 34,000 convolutional neural network, a statistically significant improvement other the 18.95\% error of thresholding, while MRFs and CRFs do no better than thresholding.
+\cite{Jain2007} achieved an 8.0\% test error with respect to hand-labelled ground truth with a convolutional neural network.
+This beats the 18.95\% error of thresholding; MRFs and CRFs do no better.
 Andres et al.
-obtain similar performance, at 8\% misclassification, using the watershed algorithm over a probability map of a random forest with hand-engineered features \cite{Andres2008}.
-However, the dataset used by both used a special intracellular stain which made segmentation easier.
+obtain similar performance over the same dataset using the watershed algorithm over a probability map of a random forest with hand-engineered features \cite{Andres2008}.
+However, their dataset used a special intracellular stain which made segmentation easier.
 The generality of these solutions are thus unclear.
@@
 Jurrus et al.
-\cite{Jurrus2008} first obtain 2D segmentations with watershed after preprocessing linear diffusion filtering (NB: learn what this is!), and construct a weighted graph of these segments, with edges between adjacent sections with edge costs favoring high cross-correlation of segments low lateral displacement.
-Each neural process is thus a shortest path through this graph, computed by Dijkstra's algorithm.
-This cost function is robust to merging and loss of sections, but because Dijkstra's algorithm finds a single path, it cannot model branching.
+\cite{Jurrus2008} first obtain 2D segmentations with watershed after preprocessing, and construct a graph of segments across layers.
+Edges connect all segments that overlap in the plane, and weights favor segments with high cross-correlation low lateral displacement.
+Each neural process is a shortest path through this graph, computed by Dijkstra's algorithm.
+This cost function is robust to merging and losing of sections.
+But because Dijkstra's algorithm finds a single path, it cannot model branching.
 Kaynig et al.
-\cite{Kaynig2010a} presuppose segmentation and represent the image as a graph of regions, like Jurrus.
-Pairwise features are represented as edge weights, and a random forest classifier is trained to output a scalar similarity score for each pair.
+\cite{Kaynig2010a} presuppose segmentation and also represent the image as a graph of regions.
+Edge weights store pairwise features, and a random forest is trained to output a scalar similarity score for each pair.
 Then a probabilistic model encoding geometric constraints with cross-section correspondences as hidden variables is trained with EM.
-Finally, the maximum-likelihood result is the 3D reconstruction.
+The maximum-likelihood result is the 3D reconstruction.
 These pairwise correspondences are fed to an agglomerative clustering algorithm, constructing the neural processes from ground-up in the same manner as a human.
@@
 These features are rotation, scale, and translation invariant, but strongly discriminate the types of objects observed in EM images.
-There features were used in SVMs to classify interior and boundary points of mitochondria, which are considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy.
-Kaynig et al.
-encode perceptual grouping constraints---rules of thumb used by human experts to group neuron structures---as pairwise interaction terms in an energy function solved by graph cut.
-Minimizing the energy function gives a segmentation, and identifying maximum overlapping regions between slices is sufficient to obtain a 3D reconstruction
+There features were used in SVMs to classify interior and boundary points of mitochondria, considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy.
 
@@
 \section{Model and Features}
-
-\fxwarning{What microscopy method did they use? and put in the right place.}
-We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \text{ nm}^3$ collected by \fxwarning{Unpublished; to cite}{Kasthuri et al.}
-
-An image is composed of pixels, which immediately offer three features: $x$- and $y$- coordinates, and intensity (normalized to $[0,1]$).
-One way to segment cells is to model pixels with a Gaussian mixture model (GMM).
-Each cluster represents one cell, and each pixel's mixture weights represent the probability that of membership in a given cell.
-This approach would work well if (1) the number of cells in an image are known \emph{a priori} (2) pixels intensities are homogeneous within cells; and (3) pixels belonging to a cell occupy a continuous area.
-
+\label{sec:model}
 \begin{figure}
@@
 
-\label{sec:model}
+\fxwarning{What microscopy method did they use? and put in the right place.}
+We analyzed a  volume of mouse visual cortex of resolution $3 \times 3 \times 30 \text{ nm}^3$ collected by \fxwarning{Unpublished; to cite}{Kasthuri et al.}
+
+An image is composed of pixels, which immediately offer three features: $x$- and $y$- coordinates, and intensity (normalized to $[0,1]$).
+One way to segment cells is to model pixels with a Gaussian mixture model (GMM).
+Each cluster represents one cell, and each pixel's mixture weights represent the probability that of membership in a given cell.
+This approach would work well if (1) the number of cells in an image are known \emph{a priori} (2) pixels intensities are homogeneous within cells; and (3) pixels belonging to a cell occupy a continuous area.
+
+Although the confidence regions of Gaussians are ellipsoids in feature space, it is important to recall that decision boundaries need not be ellipsoidal at all.
+This is particularly true when the data are inseparable, i.e.~when each point has a high probability of belonging to several nearby clusters.
+This is the case in our neural image data: there are many cell segments, and a pixel's membership has significant uncertainty.
+The actual decision surface is determined by the relative likelihoods of the point for its various neighboring clusters and can be quite complicated.
+
 We present our model in Figure~\ref{fig:model}.
@@
 First, the number of cells must be inferred from the image itself.
-To do so, we employ a mixture Dirichlet process, discussed in section~\ref{sec:mdp}.
+To do so, we employ a Dirichlet process mixture, discussed in section~\ref{sec:mdp}.
 Further, the visual complexity and noise of electron micrograph images do not yield the clear separations required for assumptions 2 and 3.
-We fix this with Radon-like features to separate cell interiors and boundaries in section~\ref{sec:radon}.
+We fix this with Radon-like features \cite{Kumar2010} to separate cell interiors and boundaries in section~\ref{sec:radon}.
 Finally, our model encodes spatial locality in two novel ways: in the Gaussian covariance which bias class membership to Euclidean-close points, and in the geodesic distances which penalizes boundary crossings, discussed in section~\ref{sec:geodesic}.
@@
 \begin{itemize}
-\item $\+\pi \sim \text{DirichletProcess}(\alpha)$
-\item $k | \+\pi \sim \text{Discrete}(\+\pi)$
-\item If $k$ is a new (empty) cluster:
+\item $\+\pi \sim \text{DP}(\alpha)$
+\item $z | \+\pi \sim \text{Discrete}(\+\pi)$
+\item For $k = 1, 2, \ldots$
 \begin{itemize}
-  \item $\Lambda_k | \Lambda_0, \nu \sim \text{Wishart}(\Lambda | \Lambda_0, \nu)$ (in the notation of \cite{Murphy2003})
+\item $\Lambda_k | \Lambda_0, \nu \sim \text{Wishart}(\Lambda | \Lambda_0, \nu)$
 \item $\+\mu_k | \kappa , \Lambda_k \sim \mathcal{N}\left( \+\mu_k | \+\mu_0 , (\kappa\Lambda)^{-1} \right)$
-  \item $\gamma_k | \beta_0, \gamma_0 \sim \text{Gamma}(\gamma_k | \beta_0, \text{rate} = \gamma_0)$
+\item $\gamma_k | \beta_0, \gamma_0 \sim \text{Gamma}(\gamma_k | \beta_0, \gamma_0)$
 \end{itemize}
-\item $\+x | \+\mu_k , \Lambda_k, \gamma_k \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \text{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \text{rate} = \gamma_k \right) \label{eq:xlike}$ (in the notation of section~\ref{sec:geodesic})
+\item $\+x | z = k, \left\{ \+\mu_k , \Lambda_k, \gamma_k \right\}_{k=1}^{\infty} \propto \mathcal{N}(\+x_{1:4} | \+\mu_k , \Lambda_k) \text{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \gamma_k \right) \label{eq:xlike}$
 \end{itemize}
 
-\subsection{Mixture Dirichlet Processes}
-\label{sec:mdp}
-We begin with a finite mixture model shown in Figure~\ref{fig:finitemix} and derive the mixture Dirichlet process as a limiting case. This follows the derivation first given by \cite{Rasmussen2000}.
-
-The $N$ observations are divided among $K$ classes.
-We record the class membership of $\+{x}_n$ with a latent indicator $z_n \in \{1, \ldots, K \}$.
-Given a class assignment $z_n = k$, we assume $x_n$ is drawn from a distribution $F(\+{x}|\Theta_k)$ where $\Theta_k$ is a vector of parameters.
-Therefore, a mixture model gives each class its own parameter vector, but assumes that observations are conditionally independent given a class.
-
-The indicators $z_n$ are drawn from the discrete distribution $\+\pi$, a $K$-dimensional vector which is in turn is drawn from a symmetric Dirichlet distribution parameterized by $\alpha$: $$p \left( \+\pi | \alpha, K \right) = \frac{ \Gamma ( \alpha ) }{ \Gamma ( \alpha/K ) ^K} \prod_{k=1}^K \+\pi_k^{\alpha/K - 1}.$$ Because the Dirichlet and discrete distributions are conjugate, we can marginalize $\+\pi$.
-Let $\+{z}_{-n}$ denote the vector of all indicators excluding element $n$ and $N_{k,-n}$ denote the number of points assigned to class $k$ excluding point $n$.
-A standard result \cite{Rasmussen2000} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }$$ which makes Gibbs sampling easy.
-
-Now consider $K \rightarrow \infty$.
-First, the term $\alpha / K \rightarrow 0$.
-Given finitely many observations, we have only finitely many components $k$ with $N_{k,-n} > 0$.
-Therefore, the density
-\[
-p(z_{n}=k|\+{z}_{-n},\alpha)=\begin{cases}
-\frac{N_{k,-n}}{N-1+\alpha} & N_{k,-n}>0\\
-\frac{\alpha}{N-1+\alpha} & \text{otherwise}
-\end{cases} \label{eq:mdp}
-\]
-satisfies the limiting conditions and correctly normalizes to 1 when summed over $k = 1, \ldots, \infty$.
-We now say that the mixing distribution $\+\pi$, which is now infinite-dimensional, is drawn from a \emph{Dirichlet process}.
-Since conjugacy allows us to eliminate $\+\pi$ from computations, Gibbs sampling from an infinite mixture is just as efficient as for a finite mixture case.
+We use the rate parameterization of the Gamma distribution: $\text{Gamma}(x | \beta, \gamma) = \frac{\gamma^\beta}{\Gamma(\beta)} x^{\beta-1} \exp{(-\gamma x)}$.
+The exponential distribution is just the Gamma distribution with $\alpha = 1$.
+The degrees of freedom of the Wishart distribution is $\nu$.
+The subscript $1:d$ denotes extracting the first $d$ dimensions of the vector.
 
@@
 To correct this, we augment raw pixel intensities with $\emph{Radon-like feature}$ responses, which have proven useful for connectomics segmentation \cite{Kumar2010}.
-\fxwarning{Make this clearer...
-Even the original paper doesn't have a concise description}.
-A Radon-like feature aggregates a function of image intensity over regions defined by structural constraints, such as edges.
-Each feature is measured from a single direction, which we can represent as a \emph{scan line} $\+{l}(t) : [0, 1] \rightarrow \+{R}^2$.
-We represent the structure information as a set of points $\{ t_1, \ldots, t_m \}$ along line $\+{l}$.
-We follow the practice in \cite{Kumar2010} to evaluate Radon-like features along uniform angular intervals (we find 72 to avoid most discretization artifacts) and take the mean response from all directions.
+A Radon-like feature aggregates a function of image intensity over regions defined by structural constraints.
+Given emph{scan line} $\+{l}(t) : [0, 1] \rightarrow \+{R}^2$ denoting a direction along the image, we integrate an \emph{extraction function} piecewise, with segments defined by \emph{knots}, which are structurally salient points of the image.
+In particular, we determine knots as the poitns classified as edges by using a Canny edge mapper on the filter response $$R(x,y) = \max_{\sigma, \phi} \Delta G(\sigma, \phi) \ast I(x,y) $$ given in \cite{Kumar2010}, where $\Delta G(\sigma, \phi)$ is the Gaussian second derivative filter at scale $\sigma$ and orientation $\phi$.
+This edge response is noisy, but unlike other methods, we use it the noisy edge response only as a guide for additional feature extraction, not as a testimony for the cell segments themselves.
+
+For a given scan line, we write its knots as parametric coordinates along the line: $K_{\+l} = \{ t_1, \ldots, t_m \}$.
+We follow the practice in \cite{Kumar2010} to evaluate Radon-like features along uniform angular intervals (we find 72 sufficient to avoid most discretization artifacts) and take the mean response from all directions.
 Of course, using the entire vector of feature responses with a multivariate likelihood is also possible.
-Moreover, in \cite{Kumar2010} as well as this paper, we will use the filter response $$R(x,y) = \max_{\sigma, \phi} \Delta G(\sigma, \phi) \ast I(x,y) $$ where $\Delta G(\sigma, \phi)$ is the Gaussian second derivative filter at scale $\sigma$ and orientation $\phi$.
-We use $R$ to (1) obtain knots, by passing to a Canny edge mapper, and (2) use as input to extraction functions (discussed below).
-Unlike other methods, we use the noisy edge response only as a guide for additional feature extraction, not as a testimony for the cell segments themselves.
-
-Let $T(I, x, y)$ be an $\emph{extraction function}$ which maps an image and $xy$-coordinates to a scalar.
+Let $T(I, x, y)$ be an extraction function which maps an image and $xy$-coordinates to a scalar.
 Suppose a line $\+{l}$ is given and $(x, y) = \+{l}(t)$ for some $t \in [t_i, t_{i+1}]$.
@@
 Thus, $T_{\text{bd}}$ smoothes the spurious GSD responses, avoiding oversegmentation.
-\fxwarning{Tighten up wording here.}Further, in connectomics data, cell membranes tend to form thick boundaries.
+Further, in connectomics data, cell membranes tend to form thick boundaries.
 But classic edge detectors work best at single-pixel gradients and return two edges for a thick boundary.
@@
 \includegraphics[width=0.24\linewidth, clip]{JAN_bd_hist}
-\caption{Boundary Radon-like feature response; red marks the center from which the remaining pictures are measured.
-Grayscale map of geodesic distance (of superpixels).
-Threshold of geodesic distance $< 0.1$.
-Distribution of geodesic distances about the red center.
-Although the separation between interior and exterior is easy effort, BASELINE BLAH}\label{fig:geodesic}
+\caption{(1) Boundary Radon-like feature response; red marks the center from which the remaining pictures are measured.
+(2) Grayscale map of geodesic distance (of superpixels).
+(3) Threshold of geodesic distance $< 0.1$.
+Distribution of geodesic distances about the red center.\label{fig:geodesic}}
 \end{figure}
-Figures~\ref{fig:radon} and~\ref{fig:geodesic} illustrates that cells need not be elliptical, or even convex.
+Figures~\ref{fig:radon} and~\ref{fig:geodesic} illustrate that cells need not be elliptical, or even convex.
 All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response.
 Therefore, shape constraints based on Euclidean distances are not justified.
-Instead, we want to encode the a priori belief that "the cell body should not spill outside its membrane." Let us quantify this statement.
+Instead, we want to encode the a priori belief that ``the cell body should not spill outside its membrane.'' Let us quantify this statement.
 Construct an undirected graph with pixels as vertices and pairs of adjacent pixels as edges ($(x,y)$ is an edge if $y$ is in the 1-pixel neighborhood of $x$).
@@
 All paths crossing a cell boundary must traverse pixels with low, high, and again low boundary feature responses.
-Thus, the shortest path wil also be one with a high weight.
+Thus, the shortest path will also be one with a high weight.
 So if pixels $i$ and $j$ do not belong to the same cell, defined by the region enclosed by a loop of pixels with high boundary feature response, then $d(\+{x}_i, \+{x}_j)$ is high.
@@
 This is because the likelihood of $\+x$, which depends on the function $d$, is not a known conjugate-parametric form.\fxwarning{I'm not sure this is the right explanation.} We collapse $\+\pi$, but we do not collapse the other latent variables.
-Denote by $X$ the set of all $\+x_n$ and by $X_k = \left\{ \+x_n \in X : z_n = k \right\}$ and $N_k = | X_k |$.
-Then we derive the conditional posteriors by finding the Markov blankets of each node and using Bayes' rule.
-But because the likelihood of $\+x_n$ is note a standard parametric form, conditionals involving $X)k$ and $\+\mu_k$ cannot be sampled explicitly; we can only compute their likelihoods can be computed up to proportionality and resort to Metropolis-Hastings sampling.
+Denote by $X_{-n}$ the set of all $\+x_n$ with point $n$ removed, by $X_k = \left\{ \+x_n \in X : z_n = k \right\}$ and $N_k = | X_k |$.
+Then we derive the conditionals by finding the Markov blankets of each node and using Bayes' rule.
+But because the likelihood of $\+x_n$ is not a standard parametric form, conditionals involving $X_k$ and $\+\mu_k$ cannot be sampled explicitly; we can only compute their likelihoods up to proportionality and resort to Metropolis-Hastings sampling.
 \begin{align}\label{eq:conds}
-p(z_n = k | \+\pi, \+x_n, \+\mu_k, \Lambda_k) &= p(z_n = k | \+\pi) \mathcal{N}(\+x_n | \+\mu_k, \Lambda_k) \label{eq:z}\\
-p(\gamma_k | X_{z=k} , \beta_0, \gamma_0) &= \text{Gamma}\left( \gamma_k | \beta_0 + N_k, \gamma_0 + \sum_{\+x \in X_k} \+x \right) \label{eq:gamma}\\
-p(\Lambda_k | X_k, \+\mu_k, \gamma_k, \Lambda_0, \nu) &\propto  p(X_k | \+\mu_k, \Lambda_k, \beta_k) \text{Wishart}(\Lambda_k | \Lambda_0, \nu) \label{eq:lambda}\\
+p(z_n = k | \+\pi, \+x_n, \+\mu_k, \Lambda_k) &\propto p(z_n = k | \+\pi) \mathcal{N}(\+x_n | \+\mu_k, \Lambda_k) \label{eq:z}\\
+p(\gamma_k | X_k, \+\mu_k, \beta_0, \gamma_0) &= \text{Gamma}\left( \gamma_k | \beta_0 + N_k, \gamma_0 + \sum_{\+x \in X_k} d(\+x, \+\mu_k) \right) \label{eq:gamma}\\
+p(\Lambda_k | X_k, \+\mu_k, \gamma_k, \Lambda_0, \nu) &\propto  p(X_k | \+\mu_k, \Lambda_k, \gamma_k) \text{Wishart}(\Lambda_k | \Lambda_0, \nu) \label{eq:lambda}\\
 p(\+\mu_k | X_k, \Lambda_k, \gamma_k, \+\mu_0, \kappa) &\propto p(X_k | \+\mu_k, \Lambda_k, \gamma_k) \mathcal{N}(\+\mu_k | \+\mu_0, (\kappa\Lambda)^{-1}) \label{eq:mu}
@@
 \begin{algorithm}
-\caption{Gibbs sample iteration}\label{alg:gibbs}
+\caption{Gibbs sample sweep}\label{alg:gibbs}
 \begin{algorithmic}
 \Require The current state of the Markov chain, $\Theta = \left\{ N, K, Z, \left\{ N_k, \+\mu_k, \Lambda_k, \gamma_k \right\}_{k=1}^K  \right\}$
-\Ensure The distribution of the next state $\Theta^{*}$ preserves the Markov chain's invariant distribution.
 \Statex
@@
   \[
-  p(z_{n}=k|\+{z}_{-n},\+x_n,\+\mu_k,\Lambda_k,\gamma_k,\alpha)=\begin{cases}
-  \frac{N_{k,-n}}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \beta_k) & k < K\\
-  \frac{\alpha / m}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \beta_k) & \text{otherwise}
+  p(z_{n}=k|\+{z}_{-n},X_{-n},\alpha,\left\{ \+\mu_k,\Lambda_k,\gamma_k\ \right\}_{k=1}^{\infty}) \propto \begin{cases}
+  \frac{N_{k,-n}}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \gamma_k) & k < K\\
+  \frac{\alpha / m}{N-1+\alpha} p(\+x_n | \+\mu_k, \Lambda_k, \gamma_k) & \text{otherwise}
   \end{cases} \label{eq:nealalg8}
@@
   \State Draw $\gamma_k$ using \eqref{eq:gamma}
-  \State Metropolis-Hastings sample $\Lambda_k$ using likelihood \eqref{eq:lambda} and proposal $\text{Wishart}(\Lambda^{*} | p \Lambda, \nu)$ for some $p > 1$.
+  \State Metropolis-Hastings sample $\Lambda_k$ using likelihood \eqref{eq:lambda} and proposal $\text{Wishart}(\Lambda^{*} | b \Lambda, \nu)$ for some $b > 1$.
   \State Metropolis-Hastings sample $\+\mu_k$ using likelihood \eqref{eq:mu} and proposal $\mathcal{N}( \+\mu^{*} | C )$.
@@
 C = \left(\begin{array}{cccc}
-    \sigma^{2}\\
-     & \sigma^{2}\\
-      &  & \tau^{2}\\
-       &  &  & \tau^{2}
+    \sigma^{2} & & & \text{\huge 0}\\
+     & \sigma^{2} & & \\
+      &  & \tau^{2} &\\
+    \text{\huge 0} &  &  & \tau^{2}
      \end{array}\right)
@@
 
+\appendix
+\section{Dirichlet Process Mixtures}
+\label{sec:mdp}
+We begin with a finite Gaussian mixture model and derive the Dirichlet process mixture as a limiting case. This follows the derivation first given by \cite{Rasmussen2000}.
+
+The $N$ observations are divided among $K$ classes.
+We record the class membership of $\+{x}_n$ with a latent indicator $z_n \in \{1, \ldots, K \}$.
+Given a class assignment $z_n = k$, we assume $x_n$ is drawn from a distribution $F(\+{x}|\Theta_k)$ where $\Theta_k$ is a vector of parameters.
+Therefore, a mixture model gives each class its own parameter vector, but assumes that observations are conditionally independent given a class.
+
+The indicators $z_n$ are drawn from the discrete distribution $\+\pi$, a $K$-dimensional vector which is in turn is drawn from a symmetric Dirichlet distribution parameterized by $\alpha$: $$p \left( \+\pi | \alpha, K \right) = \frac{ \Gamma ( \alpha ) }{ \Gamma ( \alpha/K ) ^K} \prod_{k=1}^K \+\pi_k^{\alpha/K - 1}.$$ Because the Dirichlet and discrete distributions are conjugate, we can marginalize $\+\pi$.
+Let $\+{z}_{-n}$ denote the vector of all indicators excluding element $n$ and $N_{k,-n}$ denote the number of points assigned to class $k$ excluding point $n$.
+A standard result \cite{Rasmussen2000} yields the posterior $$p(z_n = k | \+{z}_{-n})  = \frac{ N_{k,-n} + \alpha/K }{ N - 1 + \alpha }.$$
+
+Now consider $K \rightarrow \infty$.
+First, the term $\alpha / K \rightarrow 0$.
+Given finitely many observations, we have only finitely many components $k$ with $N_{k,-n} > 0$.
+Therefore, the density
+\[
+p(z_{n}=k|\+{z}_{-n},\alpha)=\begin{cases}
+\frac{N_{k,-n}}{N-1+\alpha} & N_{k,-n}>0\\
+\frac{\alpha}{N-1+\alpha} & \text{otherwise}
+\end{cases} \label{eq:mdp}
+\]
+satisfies the limiting conditions and correctly normalizes to 1 when summed over $k = 1, \ldots, \infty$.
+The mixing distribution $\+\pi$, which is now infinite-dimensional, can also be said to be drawn from a \emph{Dirichlet process}.
+It is known that even the infinite-dimensional case conjugacy allows us to eliminate $\+\pi$ from computations, making Gibbs sampling from an infinite mixture is just as efficient as for a finite mixture case.
+
 \bibliographystyle{plain}

### 2013-02-06 latex/submission.tex
fixed refs

@@
 These features are rotation, scale, and translation invariant, but strongly discriminate the types of objects observed in EM images.
-There features were used in SVMs to classify interior and boundary points of mitochondria, considerably harder to classify than other cellular structures \cite{Kaynig2010b}, and a graph cut found the final 2D segmentation with 98\% accuracy.
+Their features were used in SVMs to classify interior and boundary points of mitochondria, considerably harder to classify than other cellular structures, and a graph cut found the final 2D segmentation with 98\% accuracy.
 

### 2013-02-06 src/blackdot_summary.m
added black dot analysis

@@ -1,0 +1,35 @@
+Nslices = 100;
+
+DOTS = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_dots_all.hd5'), 500, Nslices);
+DOTS = DOTS(300:400,300:400,:);
+
+thresh = .1;
+
+% Compute frame differences
+% for n = 1:(Nslices + 1)
+%     DOTS(:,:,n) = DOTS(:,:,n+1) - DOTS(:,:,n);
+% end
+
+straight = DOTS(:);
+
+disp(['Summary statistics']);
+disp(['Mean:' num2str(mean(straight)) ' SD: ' num2str(std(straight)) ...
+      ' Median: ' num2str(median(straight)) ' Mode: ' num2str(mode(straight))]);
+
+figure;
+hist(straight, 100);
+title('Black dot feature responses of pixels');
+
+figure;
+hist(straight <= thresh, 100);
+title(['Black dot feature responses <= ' num2str(thresh)]);
+
+% Find the top 1% of points
+q90 = quantile(straight, .9);
+
+indWhite = find(DOTS > q90);
+[xW yW zW] = ind2sub([100 100 100], indWhite);
+
+figure;
+scatter3(xW, yW, zW, 1);
+

### 2013-02-06 src/startup.m

@@ -10,12 +10,12 @@ DATA_PREFIX = fullfile(pwd, '../data/kasthuri11/');
 % Add local overrides here (i.e. cluster computation)
 local_startup;
 
-S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
-%A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
-%f1 = S1(1000:1199,1000:1199);
-%f1 = S1(1200:1399,900:1099);
-%f1 = S1(1400:1999,200:799);
-f1 = S1(1600:1899,1:300);
-%a1 = A1(1000:1199,1000:1199);
-[PHI, misc_data] = feature_map(f1);
+% S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
+% %A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
+% %f1 = S1(1000:1199,1000:1199);
+% %f1 = S1(1200:1399,900:1099);
+% %f1 = S1(1400:1999,200:799);
+% f1 = S1(1600:1899,1:300);
+% %a1 = A1(1000:1199,1000:1199);
+% [PHI, misc_data] = feature_map(f1);
 

### 2013-03-01 latex/submission.tex
updated author list, some more might come
added DDP and M-H sampler code

@@
 \title{Bayesian Nonparametric Neural Image Segmentation}
-\author{Kui Tang and Frank Wood}
+\author{Kui Tang, $\sigma$(Joshua T.~Vogelstein, Jeff Lichtman, Bobby Kasthuri, Randal Burns, 
+\\ Priya Manavalan) Frank Wood}
 \date{\today}
@@
 Figures~\ref{fig:radon} and~\ref{fig:geodesic} illustrate that cells need not be elliptical, or even convex.
-All that can be said is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response.
+All we can say is that they are contained in a closed loop that is generally gives a distinct Radon-like boundary response.
 Therefore, shape constraints based on Euclidean distances are not justified.
@@
 Fix a point $n$ and a cluster $k$.
-Then $d(\+{x}_n, \+{\mu}_{k,1:2})$ is exponentially distributed, favoring distances close to zero.
+Then $d(\+{x}_{n,1:2}, \+{\mu}_{k,1:2})$ is exponentially distributed, favoring distances close to zero.
 Any point one the opposite side of a boundary from the cluster's center will be far away from zero, and hence have low probability.
@@
 Inference is performed by Gibbs sampling, using Radford Neal's Algorithm 8 for non-conjugate priors \cite{Neal2000}.
-This is because the likelihood of $\+x$, which depends on the function $d$, is not a known conjugate-parametric form.\fxwarning{I'm not sure this is the right explanation.} We collapse $\+\pi$, but we do not collapse the other latent variables.
+This is because the likelihood of $\+x$, which depends on the function $d$, is not a known conjugate-parametric form.\fxwarning{I'm not sure this is the right explanation.} To use this sampling method, we collapse $\+\pi$, but we do not collapse the other latent variables.
 Denote by $X_{-n}$ the set of all $\+x_n$ with point $n$ removed, by $X_k = \left\{ \+x_n \in X : z_n = k \right\}$ and $N_k = | X_k |$.
-Then we derive the conditionals by finding the Markov blankets of each node and using Bayes' rule.
-But because the likelihood of $\+x_n$ is not a standard parametric form, conditionals involving $X_k$ and $\+\mu_k$ cannot be sampled explicitly; we can only compute their likelihoods up to proportionality and resort to Metropolis-Hastings sampling.
+Then we derive the conditionals by finding the Markov blankets of each node and using Bayes' rule:
 \begin{align}\label{eq:conds}
-p(z_n = k | \+\pi, \+x_n, \+\mu_k, \Lambda_k) &\propto p(z_n = k | \+\pi) \mathcal{N}(\+x_n | \+\mu_k, \Lambda_k) \label{eq:z}\\
+p(z_n = k | \+\pi, \+x_n, \+\mu_k, \Lambda_k, \gamma_k) &\propto p(z_n = k | \+\pi) \mathcal{N}(\+x_n | \+\mu_k, \Lambda_k) \text{Exponential}\left( d(\+x_{1:2}, \+\mu_{k,1:2}) | \gamma_k \right) \label{eq:z}\\
 p(\gamma_k | X_k, \+\mu_k, \beta_0, \gamma_0) &= \text{Gamma}\left( \gamma_k | \beta_0 + N_k, \gamma_0 + \sum_{\+x \in X_k} d(\+x, \+\mu_k) \right) \label{eq:gamma}\\
@@
 \end{align}
-Note that $\+\pi$ remains collapsed; to evaluate \eqref{eq:z}, see \eqref{eq:mdp}.
+
+Because the likelihood of $\+x_n$ is not a standard parametric form, conditionals involving $X_k$ and $\+\mu_k$ cannot be sampled explicitly; we can only compute their likelihoods up to proportionality and resort to Metropolis-Hastings sampling.
+Further, $\+\pi$ remains collapsed; to evaluate \eqref{eq:z}, see \eqref{eq:mdp}.
 
@@
   \If{$N_{z_n} = 1$}
-    \State $\left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right) := \left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right)$ 
+    \State $\left( \+\mu_{K+1}, \Lambda_{K+1}, \gamma_{K+1} \right) := \left( \+\mu_{z_n}, \Lambda_{z_n}, \gamma_{z_n} \right)$ 
     \State Draw $\+\mu_j, \Lambda_j, \gamma_j$ from their priors (see section~\ref{sec:model}), for $j \in \{K+2, \ldots, K+m\}$

### 2013-03-01 src/GammaGamma.m

@@ -73,6 +73,11 @@ classdef GammaGamma < OnlineDistribution
             xs = gamrnd(o.shape, 1 / rate, n, 1);
         end
         
+        function xs = sample_posterior(o, n)
+            rate = gamrnd(o.post_shape, 1 / o.post_rate);
+            xs = gamrnd(o.post_shape, 1 / rate, n, 1);
+        end
+        
         function p = pred_like_scalar(o, X)            
             % Dubey (1.1)
             if X == 0 % avoid log(0)

### 2013-03-01 src/GraphDist.m

@@ -1,0 +1,58 @@
+classdef GraphDist
+% Compute geodesic distance. Separated from probability model code.
+    
+    properties
+        misc_data
+        
+        % Hack; to softcode later
+        x_max;
+        y_max;
+    end        
+            
+    methods
+        function o = GraphDist(misc_data)
+            o.misc_data = misc_data;
+            o.x_max = size(misc_data.img, 1);
+            o.y_max = size(misc_data.img, 2);
+        end
+        
+        function ss = slic_segment(o, xy)
+        % ss = get.slic_segment(xy) x(1) is x; x(2) is y coord
+        
+            s_x = round(xy(1));
+            s_y = round(xy(2));
+            
+            % Clamp to the image boundary
+            if s_x < 1 || s_x > o.x_max || s_y < 1 || s_y > o.y_max
+                warning(['crossed image boundary: s_x = ' num2str(s_x) ...
+                         ' s_y = ' num2str(s_y)]);
+            end
+
+            s_x = min(max(1, s_x), o.x_max);
+            s_y = min(max(1, s_y), o.y_max);                
+
+            ss = o.misc_data.segments(s_x,s_y);                        
+        end
+        
+        function d = dist(o, X, mu)
+        % d = dist(X, mu) return the precomputed geodesic distance from the
+        % full-dimensional X and the cluster center mu
+            sx = o.slic_segment(X(1:2));
+            sm = o.slic_segment(mu(1:2));
+            
+            d = o.misc_data.slic_dist(sx, sm);            
+        end
+                        
+        function d = slic_to_dist(o, x)                        
+            if isempty(o.last_fitted_kss) || o.last_fitted_kss ~= o.k_slic_segment
+                %warning('refit triggered');
+                kidxs  = o.mdp.cluster_assigns == o.my_k; 
+                kslics = o.mdp.X(kidxs,o.slic_dim);
+                o.fit(kslics);
+            end
+            
+            d = o.misc_data.slic_dist(o.last_fitted_kss,x);
+        end
+    end    
+end
+

### 2013-03-01 src/MDPLight.m

@@ -1,0 +1,144 @@
+classdef MDPLight < matlab.mixin.Copyable
+%MDP Only Bookkeeping for Mixture of Dirichlet Process distribution.
+% This version doesn't care about likelihoods.
+%
+% Also supports adding arbitrarily many new classes.
+%
+% References
+% [1] Radford Neal, Markov Chain Sampling Methods for Dirichlet Process
+%     Mixture Models
+    
+    properties
+        concentration
+        
+        % Just storage
+        N = 0;
+        data_dim
+        X
+        
+        misc_data        
+        % prior is a struct with fields specified by the particular user of
+        % this class.
+        prior
+        
+        % This is part of the Markov chain (as are the cluster
+        % assignments), but we need to manage the sampled parameters (put
+        % in state) together with the rest of the clusters for
+        % adding/removing/etc.
+        state
+        
+        % N x 1 vector with range [0, num_clusters]. 0 represents
+        % unassigned and 1 represents background.
+        cluster_assigns = [];
+        % Number of currently active clusters
+        n_clusters = 0;
+        % Summary statistics
+        cluster_counts = [];                        
+    end
+    
+    methods
+        function o = MDPLight(concentration_or_rhs, X, prior, state, misc_data, cluster_assigns)
+            if nargin == 1 % copy constructor
+                % Assign the arrays
+                rhs = concentration_or_rhs;
+                o.concentration = rhs.concentration;
+                o.N = rhs.N;
+                o.data_dim = rhs.data_dim;
+                o.X = rhs.X;
+                o.misc_data = rhs.misc_data;
+                o.prior = rhs.prior;
+                o.state = rhs.state;
+                
+                o.cluster_assigns = rhs.cluster_assigns;
+                o.n_clusters = rhs.n_clusters;
+                o.cluster_counts = rhs.cluster_counts;
+                
+                % INITIALIZE THE STATE
+                
+            elseif nargin > 0 % support empty constructor for array construction
+                % cluster_assigns must be all specified, for now.            
+                [o.N, o.data_dim] = size(X);
+                o.X = X;
+                o.misc_data = misc_data;
+                o.concentration = concentration_or_rhs;
+                o.misc_data = misc_data;                
+                o.prior = prior;     
+                o.state = state;
+                
+                nzidxs = cluster_assigns > 0;
+                if sum(nzidxs > 0)                
+                    % Initialize our model with cluster_assigns
[…]

### 2013-03-01 src/NormalWishart.m

@@ -177,17 +177,32 @@ classdef NormalWishart < matlab.mixin.Copyable & OnlineDistribution
             p = fast_mvtpdf_scalar(x, o.pred_mvtparams);
         end
         
+        function [mu C] = sample_prior(o, n)
+            % mu is cluster mean, C is cluster COVARIANCE (not precision)
+            
+            % MATLAB uses upper triangular Cholesky (opposite of most
+            % books)
+            prior_cov = o.prior_chol' * o.prior_chol;
+            C  = iwishrnd(prior_cov, o.prior_dof);
+            mu = mvnrnd(o.prior_mean, C / o.prior_n);
+        end
+        
+        function [mu C] = sample_post(o, n)
+            % mu is cluster mean, C is cluster COVARIANCE (not precision)
+            post_cov = o.post_chol' * o.post_chol;
+            C  = iwishrnd(post_cov, o.post_dof);
+            mu = mvnrnd(o.post_mean, C / o.post_n);
+        end
+        
         function [X] = sample_prior_pred(o, n)
 % X = sample_prior(n) samples one mean and cov and n iid datapoints
-            C  = iwishrnd(o.prior_cov, o.prior_dof);
-            mu = mvnrnd(o.prior_mean, C);
+            [C, mu] = o.sample_prior(n);            
             X  = mvnrnd(mu, C, n);
         end
         
         function [X] = sample_posterior_pred(o, n)
 % X = sample_posterior(n) samples one mean and cov and n iid datapoints
-            C  = iwishrnd(o.post_cov, o.post_dof);
-            mu = mvnrnd(o.post_mean, C);
+            [C, mu] = o.sample_post(n);          
             X  = mvnrnd(mu, C, n);
         end
     end

### 2013-03-01 src/NormalWishartOffline.m

@@ -1,0 +1,146 @@
+classdef NormalWishartOffline < matlab.mixin.Copyable & OnlineDistribution
+%NormalWishart Deterministically fit a Gaussian likelihood to its conjugate prior
+%
+% References:
+% [1] Kevin Murphy, Conjugate Bayesian analysis of the Gaussian
+%     distribution, www.cs.ubc.ca/~murphyk/Papers/bayesGauss.pdf
+% [2] Andrew Gelman et al. Bayesian Data Analysis, Second Edition.
+% [3] Stéphane Ross et al, Bayesian Reinforcement Learning in Continuous POMDPs with
+%     Application to Robot Navigation, IEEE ICRA 2008
+% [4] M. H. DeGroot, Optimal Statistical Decisions. McGraw-Hill, 1970.
+    
+    properties
+        prior_mean, prior_cov, prior_dof, prior_n, dim
+        data_n
+        
+        % No storing covariances; only their cholesky factors.
+        post_mean,      post_chol, post_dof, post_n
+        pred_mvtparams, pred_chol, pred_dof, pred_n
+    end
+    
+    properties (Dependent)
+        pred_mean, post_cov, pred_cov
+    end            
+    
+    methods
+        function pm = get.pred_mean(o)
+            pm = o.post_mean;
+        end
+        
+        function poc = get.post_cov(o)
+            poc = (o.post_chol' * o.post_chol);
+        end
+        
+        function prc = get.pred_cov(o)
+            prc = (o.pred_chol' * o.pred_chol);
+        end
+        
+        function o = NormalWishartOffline(prior_mean_or_rhs, prior_cov, prior_dof, prior_n)
+            if nargin == 1 % copy constructor
+                % REMEMBER ME!!! Standard trick to copy fields. 
+                fns = properties(prior_mean_or_rhs);
+                for i=1:length(fns)
+                    o.(fns{i}) = rhs.(fns{i});
+                end
+            else                
+                o.prior_mean = prior_mean_or_rhs;
+                o.prior_cov  = prior_cov;
+                o.prior_dof  = prior_dof;
+                o.prior_n    = prior_n;
+
+                o.dim = size(prior_cov, 1);               
+                
+                % IMPORTANT: Make the distribution ready for prediction and
+                % update
+                o.fit([]);
+            end
+        end                
+        
+        % All of the Bayesian update equations take constant time (wrt d).
+        % Online updating really just updates the data covariance.
+        %
[…]

### 2013-03-01 src/ProductDistribution.m

@@ -92,7 +92,8 @@ classdef ProductDistribution < matlab.mixin.Copyable & OnlineDistribution
             for i = 1:o.N
                 p = p * o.pdfs{i}.pred_like_scalar(x(o.idxranges{i}));
             end
-        end        
+        end
+        
     end
     
 end

### 2013-03-01 src/blackdot_summary.m

@@ -27,7 +27,7 @@ title(['Black dot feature responses <= ' num2str(thresh)]);
 % Find the top 1% of points
 q95 = quantile(straight, .95);
 
-indWhite = find(DOTS > q90);
+indWhite = find(DOTS > q95);
 [xW yW zW] = ind2sub([100 100 100], indWhite);
 
 figure;

### 2013-03-01 src/ddp/my_log_prob/my_log_niwdir_coefratio.m

@@ -1,0 +1,63 @@
+function log_niwdir_coefratio = my_log_niwdir_coefratio(obs, params)
+
+addpath('~/proj/ddpTracking/src/mcmc/');
+
+
+params_n = update_mvnmn_params(obs, params);
+
+
+n = size(obs, 1);
+
+% for spatial dimensions
+d = 2;
+
+
+
+
+% compute log niw coef ratio
+
+t1 = my_log_mvgamma(params_n{5}/2, d);
+
+t2 = (params{5}/2) * log(det(params{4}));
+
+if params{7} > 0
+	
+	t3 = (d/2) * log(params{7});
+
+else
+
+	t3 = 0;
+
+end
+
+t4 = (n*d/2) * log(pi);
+
+t5 = my_log_mvgamma(params{5}/2, d);
+
+t6 = (params_n{5}/2) * log(det(params_n{4}));
+
+if params{7} > 0
+	
+	t7 = (d/2) * log(params_n{7});
+
+else
+
+	t7 = 0;
+
+end
+
+
+log_niw_coefratio = t1 + t2 + t3 - t4 - t5 - t6 - t7;
+
+
+
+% compute log dir coef ratio
+
+%log_dir_coefratio = my_log_dircoef(params{8}) - my_log_dircoef(params_n{8} / sum(obs(3:end-1)));
+log_dir_coefratio = my_log_dircoef(params{8}) - my_log_dircoef(params_n{8});
+
+
+
+% compute log niwdir coef ratio
+
+log_niwdir_coefratio = log_niw_coefratio - 3; % + log_dir_coefratio;

### 2013-03-01 src/ddp/smc1/smc1_infer.m

@@ -1,0 +1,48 @@
+function state = smc1_infer(state,data,params,numsamples,dirstring)
+% inference for a 1-particle particle filter (sequential monte carlo)
+    
+    %addpath('~/proj/ddpTracking/src/viz/',...
+    %        '~/proj/ddpTracking/src/misc/');
+
+    t_start = size(state{2},1)+1;
+    t_end = max(data(:,end));
+    for t = t_start:t_end
+        for s = 1:numsamples
+            % Sample from proposal
+            state = smc1_sample(state,data,params,t,s,numsamples);
+            display_progress();
+            periodic_draw_viz();
+        end
+        state = clean_state(state);
+        periodic_save()
+    end
+
+    % Auxiliary functions    
+    % -------------------
+    function periodic_draw_viz()
+        if true % mod(t-t_start+1,5)==0 || t==t_start || t==t_end
+            clf
+            viz_smc1_result(data(1:size(state{1},2),[1,2,end]),state,'fil','stddev');
+            %view(360*s/25,25) % rotating view
+            view(45,25);
+            zlim([1, t_end]);
+            drawnow
+        end
+    end
+
+    function display_progress()
+        if s==1
+            fprintf('Completed: t=%d, sample=%d',t,s);
+        elseif s==numsamples
+            fprintf(',%d (all samples complete)\n',s);
+        else
+            fprintf(',%d',s);
+        end
+    end
+
+    function periodic_save()
+        % save workspace periodically
+        if mod(t-t_start+1,5)==0 || t==t_end, save([dirstring, 'NEW_INFER_WS.mat']); end
+    end
+
+end

### 2013-03-01 src/ddp/smc1/smc1_main_usbh.m

@@ -1,0 +1,15 @@
+function smc1_main_usbh()
+% main smc1 script for usbh (ants) data
+
+    addpath('~/proj/ddpTracking/src/misc/', ...
+            '~/proj/ddpTracking/src/viz/');
+
+    load('~/proj/ddpTracking/data/data_usbh_combdist0L_10bins_noresize_1.mat','data');
+    data = rarefy_data(data, 10000);
+
+    % Format: params = {crp, del, #aux, tau0, v0, mu0, k0, q0, _,_,_<-#colorbins?}
+    params = {0.1, 0.7, 10, 3*eye(2), 60, [0,0], 0.05, 10*ones(1,10), 1, 1, 1};
+
+    % Format: state = {assignments, clusterParams, clusterSizes}
+    state = {[],{},zeros(size(data,1),0)};
+    state = smc1_infer(state,data,params,5,'~/proj/ddpTracking/results/ants/');

### 2013-03-01 src/ddp/smc1/smc1_sample.m

@@ -1,0 +1,135 @@
+function state = smc1_sample(state,data,params,t,s,numsamples)
+% sample each (alive) cluster's parameters, each assignment,
+% perform deletion step, update size matrix
+    
+    %addpath('~/proj/ddpTracking/src/my_log_prob/',...
+    %        '~/proj/ddpTracking/src/prob/',...
+    %        '~/proj/ddpTracking/src/mcmc/');
+
+    % data_t holds observations at current time t
+    ind_t = find(data(:,end)==t);
+    if size(ind_t,1)>0 ind_t_range = ind_t(1):ind_t(end); else ind_t_range = []; end
+    data_t = data(ind_t,:);
+
+    % final_obs is index of 'most recent' observation (from either current or a previous time)
+    temp = ind_t;
+    temp_t = t;
+    while length(temp)==0 && temp_t>0
+        fprintf('Zero observations for t = %d.\n',temp_t);
+        temp_t = t-1;
+        temp = find(data(:,end)==temp_t);
+    end
+    if temp_t>0 final_obs = temp(end); else final_obs = 0; end  %%%% will this work with the code below?..i call state{3}(final_obs, k)...
+
+    % K is number of clusters
+    K = size(state{2},2);
+    K_init = K;
+
+    % update cluster params based on assignments &/or transition kernel.
+    %   (for s=1, t>1, initialize to clusters at previous time.)
+    for k = 1:K
+        if t>1 && s==1
+            % case where s=1 and t>1,
+            %   initialize cluster params to previous cluster params
+            state{2}{t,k} = state{2}{t-1,k};    
+        else
+            if state{3}(final_obs,k)>0;
+                % if cluster is alive
+                if t==1
+                    % case where t = 1 (and for any s)
+                    data_k = data_t(find(state{1}(ind_t_range)==k),:); % get all obs (of t=1) assigned to cluster k
+                    params_post = update_mvnmn_params(data_k(:,1:2),data_k(:,3:end-1),params);
+                else
+                    % case where s>1 && t>1
+                    %   get all obs (of current t) assigned to cluster k
+                    data_k = data_t(find(state{1}(ind_t_range)==k),:);    
+                    if size(data_k,1)>0 && length(state{2}{t-1,k})>0
+                        % if obs assigned and there exists prev cluster params
[…]

### 2013-03-01 src/experiment/test_mdp.m

@@ -1,10 +1,9 @@
 %% Prior!
 clear mdp
 
-[misc_data.segments, misc_data.slic_inds, misc_data.edge_G] = ...
-    make_diff_graph(f1, 10, 0.1);
+[misc_data.segments, misc_data.slic_inds, misc_data.edge_G] = make_diff_graph(f1, 10, 0.1);
 
-mdp = make_mdp_prior(PHI, misc_data);
+[mdp, init_state, params]  = make_mdplight_prior(PHI, misc_data);
 
 %% Run!
-run_mdp(@neal3_iter, [], [], PHI, f1, mdp, 1000, 0, 1);
+run_mdp(@neal8_iter, init_state, params, PHI, f1, mdp, 1000, 0, 1);

### 2013-03-01 src/kuiDots/blackdot_summary.m

@@ -1,0 +1,37 @@
+%Nslices = 100;
+
+%DOTS = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_dots_all.hd5'), 500, Nslices);
+%DOTS = DOTS(300:400,300:400,:);
+
+load dots.mat
+
+thresh = .1;
+
+% Compute frame differences
+% for n = 1:(Nslices + 1)
+%     DOTS(:,:,n) = DOTS(:,:,n+1) - DOTS(:,:,n);
+% end
+
+straight = DOTS(:);
+
+disp(['Summary statistics']);
+disp(['Mean:' num2str(mean(straight)) ' SD: ' num2str(std(straight)) ...
+      ' Median: ' num2str(median(straight)) ' Mode: ' num2str(mode(straight))]);
+
+figure;
+hist(straight, 100);
+title('Black dot feature responses of pixels');
+
+figure;
+hist(straight <= thresh, 100);
+title(['Black dot feature responses <= ' num2str(thresh)]);
+
+% Find the top 5% of points
+q95 = quantile(straight, .95bl);
+
+indWhite = find(DOTS > q90);
+[xW yW zW] = ind2sub([100 100 100], indWhite);
+
+figure;
+scatter3(xW, yW, zW, 1);
+

### 2013-03-01 src/kuiDots/main_kuiDots.m

@@ -1,0 +1,12 @@
+
+    %addpath(genpath('./../../ddpcode/DDPMO_src')) % path to ddp code
+    load data.mat % path to data matrix output from makeData.m
+
+    % Format: 
+    %params = {crp, del, #aux,       tau0,  v0,   mu0,   k0,           q0, _, _, _<-#colorbins?}
+    % v0 variance
+    params  = {10,  0.3,   10, 0.5*eye(2), 100, [0,0], 0.05, 10*ones(1,5), 1, 1, 1};
+
+    % Format: state = {assignments, clusterParams, clusterSizes}
+    state = {[],{},zeros(size(data,1),0)};
+    state = smc1_infer(state,data,params,5,'./');

### 2013-03-01 src/make_mdplight_prior.m

@@ -1,0 +1,51 @@
+function [ mdp, state, params ] = make_mdplight_prior( X, misc_data )
+    [N, D] = size(X);
+    nX = max(X(:,1));
+    nY = max(X(:,2));
+    
+    %% Parameter priors
+    prior_mean = [100 100 0.9 0.4]'; % mean of the nonzero entries
+    prior_cov  = cov(X(:,1:4));
+    
+    v_intensity = prior_cov(3,3);
+    v_bg_radon  = prior_cov(4,4);
+    
+    prior_cov(3:end,:) = 0;
+    prior_cov(:,3:end) = 0;    
+        
+    prior_cov(3,3) = 0.5 * v_intensity;
+    prior_cov(4,4) = 0.5 * v_bg_radon;        
+    
+    prior_dof = 5;
+    prior_n   = 2;
+    
+    concentration = 1;        
+        
+    prior = struct('rate_shape', 20, 'rate_rate', 10, ...
+                   'mean', prior_mean, 'n', prior_n, ...
+                   'dof', prior_dof, 'prec_chol', chol(inv(prior_cov)), ...
+                   'bg_shape', 10, 'bg_rate', 5);
+    
+    % cluster_assigns specifies two clusters. Start the Markov chain at
+    % their priors.
+    %
+    % Initialization is really arbitrary...
+    state(1).bg_rate = 10;    
+    state(2).mean = prior_mean;
+    state(2).rate = 10;
+    state(2).prec_chol = chol(inv(prior_cov));  
+    
+    cluster_assigns = initialize_one_cluster_and_background(0.99, 0.7, X);               
+    mdp = MDPLight(concentration, X, prior, state, misc_data, cluster_assigns);   
+    
+    graph_dist = GraphDist(misc_data);
+    
+    params = struct('m', 3, 'b', 3, ...
+                    'precision_slice_sample', false, 'mean_slice_sample', false, ...
+                    'bg_Z', length(X), 'graph_dist', graph_dist, ...
+                    'mu_prop_chol', chol(diag([1/20 1/20 1/5 1/5])));
+
+
+    % Add the background class
+    % TODO: Change or sensitivity analysis!            
+end

### 2013-03-01 src/neal8_iter.m

@@ -1,30 +1,228 @@
-function [ mdp, state, ll ] = neal8_iter( X, mdp, state, params )
+function [ mdp, state, ll ] = neal8_iter( mdp, state, params )
 % Gibbs sampling iteration for Algorithm 8 in Neal 2000 [1]
 %
-% state: 
+% params is a struct with fields:
+%   m - number of trial clusters (try 3)
+%   b - scale for Wishart proposal
+%   precision_slice_sample - slice sample?
+%   mean_slice_sample   - slice sample?
+%   bg_Z - normalizing constant for the background likelihood
+%   graph_dist - a GraphDist loaded with 
+%
+% mdp.state is a n_clusters struct matrix with fields:
+%   mean, prec_chol, rate - sampled posterior parameters; 4x1, 4x4, scalar.
+%   bg_rate
+%   
+%   mdp.state(1).
+%
+% mdp is an MDPLight with a field prior, a struct with fields:
+%   rate_shape - 
+%   rate_rate  -
+%   mean      -
+%   n         -
+%   dof       -
+%   prec_chol -
+%   bg_shape
+%   bg_rate
+%
 % References:
 % [1] Radford Neal, Markov Chain Sampling Methods for Dirichlet Process
 %     Mixture Models
-    
-    unrealized_likes = cell(params.m, 1);
-    
-    ll = 0;
+            
+    ll = 0;    
     for n = 1:mdp.N
         k_old = mdp.remove_point(n);
+        %assert(k_old ~= 1, 'Cannot remove the background class!');
         
-        if mdp.cluster_count(k_old) == 0
+        %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+        % Prepare the auxiliary clusters
+        %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+        if mdp.cluster_counts(k_old) == 0
             % Singleton: consider the current distribution as a new
             % candidate
-            m = params.m - 1;
-            unrealized_likes(params.m) = mdp.cluster_likes{k_old};
+            m_start = mdp.n_clusters + 2;
+            m_end = m_start + params.m - 1;            
         else
-            m = params.m;
+            m_start = mdp.n_clusters + 1;
+            m_end = m_start + params.m - 1;
         end
         
[…]

### 2013-03-01 src/scrape/kasthuri11_dots.sh

@@ -1,0 +1,5 @@
+#!/bin/bash
+. secret.m
+
+/usr/bin/curl -o ../../data/kasthuri11/kasthuri11_dots_all.hd5 $DOTS_URL
+

### 2013-03-01 src/startup.m

@@ -5,17 +5,20 @@ addpath(fullfile(pwd, 'plot'));
 addpath(fullfile(pwd, 'util'));
 addpath(fullfile(pwd, 'radonLikeFeatures'));
 
+% DDP
+addpath(genpath(fullfile(pwd, 'ddp')));
+
 DATA_PREFIX = fullfile(pwd, '../data/kasthuri11/');
 
 % Add local overrides here (i.e. cluster computation)
 local_startup;
 
-% S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
-% %A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
-% %f1 = S1(1000:1199,1000:1199);
-% %f1 = S1(1200:1399,900:1099);
-% %f1 = S1(1400:1999,200:799);
-% f1 = S1(1600:1899,1:300);
-% %a1 = A1(1000:1199,1000:1199);
-% [PHI, misc_data] = feature_map(f1);
+S1 = load_slices(fullfile(DATA_PREFIX, 'kasthuri11_mid.hd5'), 1, 1);
+%A1 = load_ann_slices('../data/kasthuri11/kasthuri11_ann_mid.hd5', 1, 1);
+%f1 = S1(1000:1199,1000:1199);
+%f1 = S1(1200:1399,900:1099);
+%f1 = S1(1400:1999,200:799);
+f1 = S1(1600:1899,1:300);
+%a1 = A1(1000:1199,1000:1199);
+[PHI, misc_data] = feature_map(f1);
 

### 2013-03-01 src/threshCmat.m

@@ -1,0 +1,10 @@
+function [ cmat ] = threshCmat( blackdots, quantile )
+    T = size(blackdots, 3);
+    cmat = cell(T, 1);
+    for t = 1:T
+        [x y] = find(blackdots(:,:,t) > quantile);
+        cmat{t} = [x y]
+    end
+end
+
+
